---
title: "Cell communication analysis"
output: html_notebook
---



# General Setup
Setup chunk

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_knit$get("root.dir")
```

Setup reticulate

```{r}
library(reticulate)
knitr::knit_engines$set(python = reticulate::eng_python)
py_available(initialize = FALSE)
use_python(Sys.which("python"))
py_config()
```

Load libraries

```{r}
library(Seurat)
library(ggplot2)
library(ggridges)
library(dplyr)
library(data.table)
```



# Load and preprocess data
Load data (from all cells)

```{r}
allcells_css = readRDS(file = "data/processed/allcells_css.RDS")
end_cells = readRDS(file = "results/zonation_cond/end_cells_zonation.RDS")
hep_cells = readRDS(file = "results/zonation_cond/hep_cells_zonation.RDS")
```

Subset and process each condition

```{r}
cpdb_path = "results/cell_comm/CellPhoneDB_runs"

cond_cells = list()
for(cond in unique(allcells_css@meta.data$Condition)){
  print(cond)
  cond_cells[[cond]] = allcells_css[,allcells_css@meta.data$Condition==cond]
  
  # remove doublet cells
  cond_cells[[cond]] = cond_cells[[cond]][,cond_cells[[cond]]@meta.data$allcells_major!="Doublets"]
  
  # remove contaminating hepatocytes
  cond_cells[[cond]] = cond_cells[[cond]][,colnames(cond_cells[[cond]]) %in% colnames(hep_cells[[cond]]) | cond_cells[[cond]]@meta.data$allcells_major!="Hepatocytes"]
  
  # remove contaminating endothelial cells
  cond_cells[[cond]] = cond_cells[[cond]][,colnames(cond_cells[[cond]]) %in% colnames(end_cells[[cond]]) | !startsWith(cond_cells[[cond]]@meta.data$allcells_simp,"LSEC")]
  
  # add zonation labels
  meta = data.frame(row.names = rownames(cond_cells[[cond]]@meta.data),
                    "Cell" = rownames(cond_cells[[cond]]@meta.data), 
                    "cell_type" = as.character(cond_cells[[cond]]@meta.data$allcells_simp),
                    stringsAsFactors = F)
  
  hep_zones = hep_cells[[cond]]@meta.data$zonation_int
  levels(hep_zones) = c("(-0.0688,-0.0253]" = "Hepatocytes_Z1",
                        "(-0.0253,0.018]" = "Hepatocytes_Z2",
                        "(0.018,0.0615]" = "Hepatocytes_Z3")
  hep_zones = as.character(hep_zones)
  names(hep_zones) = colnames(hep_cells[[cond]])
  meta[names(hep_zones),"cell_type"] = hep_zones
  
  end_zones = end_cells[[cond]]@meta.data$zonation_int
  levels(end_zones) = c("(-0.0194,6.47]" = "LSEC_Z1",
                        "(6.47,12.9]" = "LSEC_Z2",
                        "(12.9,19.4]" = "LSEC_Z3")
  end_zones = as.character(end_zones)
  names(end_zones) = colnames(end_cells[[cond]])
  meta[names(end_zones),"cell_type"] = end_zones
  
  write.table(meta, file = paste0(cpdb_path, "/", cond, "_meta_names_simp.txt"), 
            sep = "\t", col.names = T, row.names = F, quote = F)
  
  # normalise and save
  cond_cells[[cond]] = suppressWarnings(SCTransform(cond_cells[[cond]], do.correct.umi = T, 
                                                   verbose = F, 
                                                   vars.to.regress=c("unique_name",
                                                                     "nCount_RNA"),
                                                   variable.features.rv.th = 1, seed.use = 1,
                                                   return.only.var.genes = F, 
                                                   variable.features.n = NULL))
  
  dat = cbind(rownames(cond_cells[[cond]]@assays$SCT@data),
              Matrix::as.matrix(cond_cells[[cond]]@assays$SCT@data))
  colnames(dat)[1] = "Gene"
  write.table(dat, file = paste0(cpdb_path, "/", cond, "_exp_norm.txt"), 
              sep = "\t", col.names = T, row.names = F, quote = F)
}
```

## Running CellPhoneDB
Commands for runnin CellPhoneDB will be written here. Results are output to the `local1` disk to avoid I/O issues, but the output folders should then be copied to: `results/cell_comm_healthy/CellPhoneDB_runs`.  
NOTE: for some reason I'm getting a SegFault when trying to run the heatmap_plot command. This was run on the Euler server instead.

```{r}
for(n in names(cond_cells)){
comm1 = paste0("cellphonedb method statistical_analysis results/cell_comm/CellPhoneDB_runs/", n, "_meta_names_simp.txt results/cell_comm/CellPhoneDB_runs/", n, "_exp_norm.txt --threads=8 --output-path=/local1/USERS/tomasgomes/liver/CellPhoneDB_res --project-name ", n, " --counts-data gene_name")
comm2 = paste0("cp -r /local1/USERS/tomasgomes/liver/CellPhoneDB_res/", n, " results/cell_comm/CellPhoneDB_runs/")
comm3 = paste0("cellphonedb plot heatmap_plot --pvalues-path ./results/cell_comm/CellPhoneDB_runs/", n, "/pvalues.txt --output-path ./results/cell_comm/CellPhoneDB_runs/", n, "/ --count-name counts_heat.pdf --count-network-name count_net.txt --interaction-count-name count_inter.txt ./results/cell_comm/CellPhoneDB_runs/", n, "_meta_names_simp.txt")

print(n)
print(comm1)
print(comm2)
print(comm3)
print(".")
}
```


## Examine results
Load results

```{r}
cpdb_path = "results/cell_comm/CellPhoneDB_runs"

sig_means_names_l = list()
dec_l = list()
net_names_l = list()
meta_l = list()
for(n in unique(allcells_css@meta.data$Condition)){
  sig_means_names_l[[n]] = read.table(paste0(cpdb_path, "/", n, "/significant_means.txt"),
                                      header = T, sep = "\t", stringsAsFactors = F)
  dec_l[[n]] = read.table(paste0(cpdb_path, "/", n, "/deconvoluted.txt"),
                          header = T, sep = "\t")
  colnames(dec_l[[n]]) = gsub(".", " ", colnames(dec_l[[n]]), fixed = T)
  colnames(dec_l[[n]])[colnames(dec_l[[n]])=="gd T cells"] = "gd-T cells"
  colnames(dec_l[[n]])[colnames(dec_l[[n]])=="ab T cells"] = "ab-T cells"
  colnames(dec_l[[n]])[colnames(dec_l[[n]])=="Endothelial cells  non LSEC "] = "Endothelial cells (non-LSEC)"
  net_names_l[[n]] = read.table(paste0(cpdb_path, "/", n, "/count_net.txt"),
                                header = T, sep = "\t", stringsAsFactors = F)
  meta_l[[n]] = read.table(paste0(cpdb_path, "/", n, "_meta_names_simp.txt"),
                           header = T, sep = "\t")
}
saveRDS(dec_l, file = "results/cell_comm/deconvoluted_list.RDS")
```

Reformat significant means matrix

```{r}
reform_list = list()
for(n in names(meta_l)){
  simp_reform = reshape2::melt(sig_means_names_l[[n]][,c(1:4,7:9,13:ncol(sig_means_names_l[[n]]))])
  simp_reform = simp_reform[complete.cases(simp_reform),]
  
  lr1 = c()
  lr2 = c()
  for(i in 1:nrow(simp_reform)){
    if(grepl("complex", simp_reform$partner_a[i])){
      lr1 = c(lr1, substr(simp_reform$partner_a[i], 9,100))
    } else{
      lr1 = c(lr1, strsplit(simp_reform$interacting_pair[i], "_")[[1]][1])
    }
    if(grepl("complex", simp_reform$partner_b[i])){
      lr2 = c(lr2, substr(simp_reform$partner_b[i], 9,100))
    } else{
      spltlen = length(strsplit(simp_reform$interacting_pair[i], "_")[[1]])
      lr2 = c(lr2, strsplit(simp_reform$interacting_pair[i], "_")[[1]][spltlen])
    }
  }
  simp_reform$lr1 = lr1
  simp_reform$lr2 = lr2
  
  simp_reform$ct1 = NA
  simp_reform$ct2 = NA
  for(ct in unique(meta_l[[n]]$cell_type)){
    ct_mod = gsub("-", " ", ct, fixed = T)
    ct_mod = gsub("(", " ", ct_mod, fixed = T)
    ct_mod = gsub(")", " ", ct_mod, fixed = T)
    st = grepl(paste0("^",ct_mod), gsub(".", " ", simp_reform$variable, fixed = T), fixed = F)
    en = grepl(paste0(ct_mod,"$"), gsub(".", " ", simp_reform$variable, fixed = T), fixed = F)
    simp_reform$ct1[st] = ct
    simp_reform$ct2[en] = ct
  }
  
  # direction: receptors activate internal signal - that is the direction of the signalling
  ## if both are true, the "net signal" is 0
  simp_reform$dir = ifelse(simp_reform$receptor_a==simp_reform$receptor_b, 0, 
                           ifelse(simp_reform$receptor_a=="True" & 
                                    simp_reform$receptor_b=="False", -1,
                                  ifelse(simp_reform$receptor_a=="False" &
                                           simp_reform$receptor_b=="True", 1, NA)))
  
  simp_reform = simp_reform[,c("id_cp_interaction", "ct1", "ct2", "lr1", "lr2", 
                               "value", "dir")]
  
  for(i in 1:nrow(simp_reform)){
    if(simp_reform[i,"dir"]==-1){
      tmp = simp_reform[i,"ct2"]
      simp_reform[i,"ct2"] = simp_reform[i,"ct1"]
      simp_reform[i,"ct1"] = tmp
      
      tmp = simp_reform[i,"lr2"]
      simp_reform[i,"lr2"] = simp_reform[i,"lr1"]
      simp_reform[i,"lr1"] = tmp
      
      simp_reform[i,"dir"]=1
    }
  }
  
 reform_list[[n]] = simp_reform
}
saveRDS(reform_list, file = "results/cell_comm/reform_list.RDS")
```

Plot interaction counts

```{r}
inter_counts_l = list()
for(n in names(net_names_l)){
  inter_counts = reshape2::dcast(data = net_names_l[[n]], 
                                 formula = SOURCE~TARGET, value.var = "count")
  rownames(inter_counts) = inter_counts[,1]
  inter_counts = inter_counts[,-1]
  
  pdf(paste0("results/cell_comm/", n, "_number_of_interactions.pdf"), useDingbats = F, 
      height = 6, width = 7)
  pheatmap::pheatmap(inter_counts, clustering_method = "ward.D2", main = "All clusters")
  dev.off()
  inter_counts_l[[n]] = inter_counts
}
```

Making two cell type by interaction matrices: one has the ligand mean, another the receptor mean

```{r}
scoring_mats = list()
for(n in names(reform_list)){
  cl_reform = reform_list[[n]]
  # add reversed non-directed interactions, to consider them in both directions
  cl_reform0 = cl_reform[cl_reform$dir==0,]
  cl_reform0 = cl_reform0[,c(1,3,2,5,4,6,7)]
  colnames(cl_reform0) = colnames(cl_reform)
  cl_reform = rbind(cl_reform, cl_reform0)
  
  dec_cl = dec_l[[n]]
  
  mat_lig_cl = data.frame(matrix(NA, nrow = length(unique(cl_reform$id_cp_interaction)), 
                   ncol = length(unique(cl_reform$ct1))))
  mat_rec_cl = data.frame(matrix(NA, nrow = length(unique(cl_reform$id_cp_interaction)), 
                   ncol = length(unique(cl_reform$ct1))))
  colnames(mat_lig_cl) = colnames(mat_rec_cl) = unique(cl_reform$ct1)
  rownames(mat_lig_cl) = rownames(mat_rec_cl) = unique(cl_reform$id_cp_interaction)
  for(i in 1:nrow(cl_reform)){
    g1 = cl_reform[i,"lr1"]
    g2 = cl_reform[i,"lr2"]
    
    c1 = cl_reform[i,"ct1"]
    c2 = cl_reform[i,"ct2"]
    
    int = as.character(cl_reform[i,1])
    
    m1 = mean(dec_cl[dec_cl$id_cp_interaction==int & dec_cl[,1]==g1,c1], na.rm = T)
    if(is.na(m1)) m1 = mean(dec_cl[dec_cl$id_cp_interaction==int & dec_cl[,5]==g1,c1], 
                            na.rm = T)
    mat_lig_cl[int, c1] = m1
    
    m2 = mean(dec_cl[dec_cl$id_cp_interaction==int & dec_cl[,1]==g2,c2], na.rm = T)
    if(is.na(m2)) m2 = mean(dec_cl[dec_cl$id_cp_interaction==int & dec_cl[,5]==g2,c2],
                            na.rm = T)
    if(is.na(m1) | is.na(m2)) print(int)
    mat_rec_cl[int, c2] = m2
  }
  
  # ligand vs receptor correlation
  cor_mat_each_cl = cor(mat_lig_cl, mat_rec_cl, use="pairwise.complete.obs", method = "sp")
  
  scoring_mats[[n]] = list("mat_lig_ct" = mat_lig_cl, "mat_rec_ct" = mat_rec_cl,
                           "cor_mat_each" = cor_mat_each_cl)
  
  # sum(lig, reg) correlation
  mat_lig_cl[is.na(mat_lig_cl)] = 0
  mat_rec_cl[is.na(mat_rec_cl)] = 0
  mat_both_cl = mat_lig_cl+mat_rec_cl
  mat_both_cl[mat_both_cl==0] = NA
  cor_mat_both_cl = cor(mat_both_cl, use="pairwise.complete.obs", method = "sp")
  
  scoring_mats[[n]]$cor_mat_both = cor_mat_both_cl
}
saveRDS(scoring_mats, file = "results/cell_comm/scoring_mats.RDS")
```

Making two cell type by interaction matrices: one has the ligand mean, another the receptor mean (here only directional)

```{r}
scoring_mats_dir = list()
for(n in names(reform_list)){
  cl_reform = reform_list[[n]]
  # add reversed non-directed interactions, to consider them in both directions
  cl_reform0 = cl_reform[cl_reform$dir==0,]
  cl_reform0 = cl_reform0[,c(1,3,2,5,4,6,7)]
  colnames(cl_reform0) = colnames(cl_reform)
  cl_reform = rbind(cl_reform, cl_reform0)
  
  cl_reform = cl_reform[cl_reform$dir==1,]
  
  dec_cl = dec_l[[n]]
  
  mat_lig_cl = data.frame(matrix(NA, nrow = length(unique(cl_reform$id_cp_interaction)), 
                   ncol = length(unique(cl_reform$ct1))))
  mat_rec_cl = data.frame(matrix(NA, nrow = length(unique(cl_reform$id_cp_interaction)), 
                   ncol = length(unique(cl_reform$ct1))))
  colnames(mat_lig_cl) = colnames(mat_rec_cl) = unique(cl_reform$ct1)
  rownames(mat_lig_cl) = rownames(mat_rec_cl) = unique(cl_reform$id_cp_interaction)
  for(i in 1:nrow(cl_reform)){
    g1 = cl_reform[i,"lr1"]
    g2 = cl_reform[i,"lr2"]
    
    c1 = cl_reform[i,"ct1"]
    c2 = cl_reform[i,"ct2"]
    
    int = as.character(cl_reform[i,1])
    
    m1 = mean(dec_cl[dec_cl$id_cp_interaction==int & dec_cl[,1]==g1,c1], na.rm = T)
    if(is.na(m1)) m1 = mean(dec_cl[dec_cl$id_cp_interaction==int & dec_cl[,5]==g1,c1], 
                            na.rm = T)
    mat_lig_cl[int, c1] = m1
    
    m2 = mean(dec_cl[dec_cl$id_cp_interaction==int & dec_cl[,1]==g2,c2], na.rm = T)
    if(is.na(m2)) m2 = mean(dec_cl[dec_cl$id_cp_interaction==int & dec_cl[,5]==g2,c2],
                            na.rm = T)
    if(is.na(m1) | is.na(m2)) print(int)
    mat_rec_cl[int, c2] = m2
  }
  
  # ligand vs receptor correlation
  cor_mat_each_cl = cor(mat_lig_cl, mat_rec_cl, use="pairwise.complete.obs", method = "sp")
  
  scoring_mats_dir[[n]] = list("mat_lig_ct" = mat_lig_cl, "mat_rec_ct" = mat_rec_cl,
                               "cor_mat_each" = cor_mat_each_cl)
  
  # sum(lig, reg) correlation
  mat_lig_cl[is.na(mat_lig_cl)] = 0
  mat_rec_cl[is.na(mat_rec_cl)] = 0
  mat_both_cl = mat_lig_cl+mat_rec_cl
  mat_both_cl[mat_both_cl==0] = NA
  cor_mat_both_cl = cor(mat_both_cl, use="pairwise.complete.obs", method = "sp")
  
  scoring_mats_dir[[n]]$cor_mat_both = cor_mat_both_cl
}
saveRDS(scoring_mats_dir, file = "results/cell_comm/scoring_mats_dir.RDS")
```

Count interactions of each type

```{r}
inter_counts_dir = list()
for(n in names(reform_list)){
  inter_counts_dir[[n]] = list()
  for(i in c(0,1)){
    cl_reform = reform_list[[n]]
    cl_reform = cl_reform[cl_reform$dir==i,]
    
    n_dir_int = matrix(0, nrow = length(unique(cl_reform$ct1)),
                       ncol = length(unique(cl_reform$ct2)))
    rownames(n_dir_int) = colnames(n_dir_int) = unique(cl_reform$ct1)
    for(c1 in unique(cl_reform$ct1)){
      for(c2 in unique(cl_reform$ct1)){
        n_dir_int[c1,c2] = nrow(cl_reform[cl_reform$ct1==c1 & cl_reform$ct2==c2,])
        if(i==0){
          n_dir_int[c1,c2] = n_dir_int[c1,c2]+nrow(cl_reform[cl_reform$ct1==c2 &
                                                               cl_reform$ct2==c1,])
        }
      }
    }
    inter_counts_dir[[n]][[as.character(i)]] = n_dir_int
  }
}
saveRDS(inter_counts_dir, file = "results/cell_comm/inter_counts_dir.RDS")
```

Calculate mutual information for total interactions

```{r, fig.height=6, fig.width=6.2}
mut_info_list = list()
for(n in names(scoring_mats)){
  mi_mat = matrix(0, nrow = ncol(scoring_mats[[n]]$mat_lig_ct), ncol = ncol(scoring_mats[[n]]$mat_rec_ct))
  rownames(mi_mat) = colnames(mi_mat) = colnames(scoring_mats[[n]]$mat_lig_ct)
  for(ct1 in colnames(scoring_mats[[n]]$mat_lig_ct)){
    for(ct2 in colnames(scoring_mats[[n]]$mat_rec_ct)){
      df = cbind(scoring_mats[[n]]$mat_lig_ct[,ct1], scoring_mats[[n]]$mat_rec_ct[,ct2])
      df = df[complete.cases(df),]
      norm_mi = mpmi::cmi(df)$mi[1,2] / sqrt(entropy::entropy(df[,1]) * entropy::entropy(df[,2]))
      mi_mat[ct1, ct2] = norm_mi
    }
  }
  m = data.frame(inter_counts_l[[n]], check.names = F)
  m$r = rownames(m)
  m = reshape2::melt(m)
  mi_mat = reshape2::melt(mi_mat)
  
  mut_info_list[[n]] = merge(mi_mat, m, by = c(1,2))
  colnames(mut_info_list[[n]]) = c("source", "target", "mi", "n")
}
saveRDS(mut_info_list, file = "results/cell_comm/mut_info_list.RDS")
```

Calculate mutual information for directed interactions

```{r, fig.height=6, fig.width=6.2}
mut_info_dir_list = list()
for(n in names(scoring_mats_dir)){
  mi_mat = matrix(0, nrow = ncol(scoring_mats_dir[[n]]$mat_lig_ct), ncol = ncol(scoring_mats_dir[[n]]$mat_rec_ct))
  rownames(mi_mat) = colnames(mi_mat) = colnames(scoring_mats_dir[[n]]$mat_lig_ct)
  for(ct1 in colnames(scoring_mats_dir[[n]]$mat_lig_ct)){
    for(ct2 in colnames(scoring_mats_dir[[n]]$mat_rec_ct)){
      df = cbind(scoring_mats_dir[[n]]$mat_lig_ct[,ct1], scoring_mats_dir[[n]]$mat_rec_ct[,ct2])
      df = df[complete.cases(df),]
      norm_mi = mpmi::cmi(df)$mi[1,2] / sqrt(entropy::entropy(df[,1]) * entropy::entropy(df[,2]))
      mi_mat[ct1, ct2] = norm_mi
    }
  }
  mi_mat[mi_mat<0] = 0 # some values are negative because of the adjustment
  
  mut_info_dir_list[[n]] = mi_mat
  
  m = data.frame(inter_counts_dir[[n]]$`1`, check.names = F)
  m$r = rownames(m)
  m = reshape2::melt(m)
  mi_mat = reshape2::melt(mi_mat)
  
  mut_info_dir_list[[n]] = merge(mi_mat, m, by = c(1,2))
  colnames(mut_info_dir_list[[n]]) = c("source", "target", "mi", "n")
}
saveRDS(mut_info_dir_list, file = "results/cell_comm/mut_info_dir_list.RDS")
```

Counting interactions

```{r}
ninter_all = lapply(inter_counts_l, function(x) data.frame("interactions" = rowSums(x)))
ninter_send = lapply(inter_counts_dir, function(x) data.frame("interactions" = rowSums(x$`1`)))
ninter_recv = lapply(inter_counts_dir, function(x) data.frame("interactions" = colSums(x$`1`)))
```

Heatmaps for total interactions

```{r}
# interaction scores
## healthy
tmp = mut_info_list$healthy
tmp$score = tmp$mi*mut_info_list$healthy$n
mm = reshape2::dcast(tmp, source~target, value.var = "score")
rownames(mm) = mm[,1]
mm = mm[,-1]
pdf("results/cell_comm/healthy_total_interaction_scores.pdf", useDingbats = F, width = 6.2, height = 6)
pheatmap::pheatmap(mm, clustering_method = "ward.D2", color = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "PuBuGn"))(100),
                   annotation_col = ninter_all$healthy)
dev.off()

## regenerating
tmp = mut_info_list$regenerating
tmp$score = tmp$mi*mut_info_list$regenerating$n
mr = reshape2::dcast(tmp, source~target, value.var = "score")
rownames(mr) = mr[,1]
mr = mr[,-1]
pdf("results/cell_comm/regenerating_total_interaction_scores.pdf", useDingbats = F, width = 6.2, height = 6)
pheatmap::pheatmap(mr, clustering_method = "ward.D2", color = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "PuBuGn"))(100),
                   annotation_col = ninter_all$regenerating)
dev.off()

cols = colorRampPalette(rev(RColorBrewer::brewer.pal(n = 7, name = "RdBu")))(100)
## difference between scores
yy = mm-mr
yy[yy>6] = 6
yy[yy<(-10)] = -10
cols_n = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "OrRd"))(100)
lims_cols_h = c(round(min(ninter_all$healthy$interactions)*1/min(ninter_all$regenerating$interactions), 0),
                round(max(ninter_all$healthy$interactions)*100/max(ninter_all$regenerating$interactions), 0))
cols_h = cols_n[(lims_cols_h[1]):(lims_cols_h[2])]
pdf("results/cell_comm/HvR_total_interaction_scores.pdf", useDingbats = F, 
    width = 6.2, height = 6)
pheatmap::pheatmap(yy, color = cols, 
                   breaks = rev(c(seq(max(yy), 0, length.out = 50), 
                                  seq(-0.001, min(yy), length.out = 50))),
                   clustering_method = "ward.D2", 
                   annotation_col = data.frame(row.names = rownames(ninter_all$healthy),
                                               "H" = ninter_all$healthy$interactions, 
                                               "R" = ninter_all$regenerating$interactions),
                   annotation_colors = list("H" = cols_h,
                                            "R" = cols_n))
dev.off()

# number of interactions
pdf("results/cell_comm/healthy_total_interaction_number.pdf", useDingbats = F, 
    width = 6.2, height = 6)
pheatmap::pheatmap(inter_counts_l$healthy, clustering_method = "ward.D2", 
                   color = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "PuBuGn"))(100),
                   annotation_col = ninter_all$healthy)
dev.off()
pdf("results/cell_comm/regenerating_total_interaction_number.pdf", useDingbats = F, 
    width = 6.2, height = 6)
pheatmap::pheatmap(inter_counts_l$regenerating, clustering_method = "ward.D2", 
                   color = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "PuBuGn"))(100),
                   annotation_col = ninter_all$regenerating)
dev.off()

## difference between number of interactins
x = inter_counts_l$healthy-inter_counts_l$regenerating
brks =  seq(-max(abs(x)), max(abs(x)), length.out = 100)
brks2 = brks[(max(x)-brks)>=0]
pdf("results/cell_comm/HvR_total_interaction_number.pdf", useDingbats = F, 
    width = 6.2, height = 6)
pheatmap::pheatmap(x, color = cols, breaks = rev(c(seq(max(x), 0, length.out = 50), seq(-0.001, min(x), length.out = 50))), 
                   clustering_method = "ward.D2", annotation_col = data.frame(row.names = rownames(ninter_all$healthy),
                                                                              "H" = ninter_all$healthy$interactions, 
                                                                              "R" = ninter_all$regenerating$interactions),
                   annotation_colors = list("H" = cols_h,
                                            "R" = cols_n))
dev.off()
```

Heatmaps for directed interactions

```{r}
# interaction scores
## healthy
tmp = mut_info_dir_list$healthy
tmp$score = tmp$mi*mut_info_dir_list$healthy$n
mm = reshape2::dcast(tmp, source~target, value.var = "score")
rownames(mm) = mm[,1]
mm = mm[,-1]
pdf("results/cell_comm/healthy_directed_interaction_scores.pdf", useDingbats = F, width = 6.2, height = 6)
pheatmap::pheatmap(mm, clustering_method = "ward.D2", color = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "PuBuGn"))(100),
                   annotation_row = ninter_send$healthy, annotation_col = ninter_recv$healthy)
dev.off()

## regenerating
tmp = mut_info_dir_list$regenerating
tmp$score = tmp$mi*mut_info_dir_list$regenerating$n
mr = reshape2::dcast(tmp, source~target, value.var = "score")
rownames(mr) = mr[,1]
mr = mr[,-1]
pdf("results/cell_comm/regenerating_directed_interaction_scores.pdf", useDingbats = F, width = 6.2, height = 6)
pheatmap::pheatmap(mr, clustering_method = "ward.D2", color = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "PuBuGn"))(100),
                   annotation_row = ninter_send$regenerating, annotation_col = ninter_recv$regenerating)
dev.off()

cols = colorRampPalette(rev(RColorBrewer::brewer.pal(n = 7, name = "RdBu")))(100)
## difference between scores
yy = mm-mr
yy[yy>3] = 3
yy[yy<(-2)] = -2
cols_n = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "Purples"))(100)
lims_cols_h = c(round(min(ninter_recv$healthy$interactions)*1/min(ninter_recv$regenerating$interactions), 0),
                round(max(ninter_recv$healthy$interactions)*100/max(ninter_recv$regenerating$interactions), 0))
cols_hr = cols_n[(lims_cols_h[1]):(lims_cols_h[2])]
pdf("results/cell_comm/HvR_directed_interaction_scores.pdf", useDingbats = F, width = 6.2, height = 6)
pheatmap::pheatmap(yy, color = cols, breaks = rev(c(seq(max(yy), 0, length.out = 50), 
                                                    seq(-0.001, min(yy), length.out = 50))), 
                   clustering_method = "ward.D2", 
                   annotation_col = data.frame(row.names = rownames(ninter_recv$healthy),
                                               "H" = ninter_recv$healthy$interactions, 
                                               "R" = ninter_recv$regenerating$interactions),
                   annotation_row = data.frame(row.names = rownames(ninter_send$healthy),
                                               "H" = ninter_send$healthy$interactions, 
                                               "R" = ninter_send$regenerating$interactions),
                   annotation_colors = list("H" = cols_hr,
                                            "R" = cols_n))
dev.off()

# number of interactions
pdf("results/cell_comm/healthy_directed_interaction_number.pdf", useDingbats = F, width = 6.2, height = 6)
pheatmap::pheatmap(inter_counts_dir$healthy$`1`, clustering_method = "ward.D2", 
                   color = colorRampPalette(RColorBrewer::brewer.pal(n = 9, 
                                                                     name = "PuBuGn"))(100),
                   annotation_col = ninter_recv$healthy, annotation_row = ninter_send$healthy)
dev.off()
pdf("results/cell_comm/regenerating_directed_interaction_number.pdf", useDingbats = F, 
    width = 6.2, height = 6)
pheatmap::pheatmap(inter_counts_dir$regenerating$`1`, clustering_method = "ward.D2", 
                   color = colorRampPalette(RColorBrewer::brewer.pal(n = 9, 
                                                                     name = "PuBuGn"))(100),
                   annotation_col = ninter_recv$regenerating, annotation_row = ninter_send$regenerating)
dev.off()

## difference between scores
x = inter_counts_dir$healthy$`1`-inter_counts_dir$regenerating$`1`
cols = colorRampPalette(rev(RColorBrewer::brewer.pal(n = 7, name = "RdBu")))(100)
brks =  seq(-max(abs(x)), max(abs(x)), length.out = 100)
brks2 = brks[(max(x)-brks)>=0]
pdf("results/cell_comm/HvR_directed_interaction_number.pdf", useDingbats = F, 
    width = 6.2, height = 6)
pheatmap::pheatmap(x, color = cols, breaks = rev(c(seq(max(x), 0, length.out = 50), 
                                                   seq(-0.001, min(x), length.out = 50))), 
                   clustering_method = "ward.D2",
                   annotation_col = data.frame(row.names = rownames(ninter_recv$healthy),
                                               "H" = ninter_recv$healthy$interactions, 
                                               "R" = ninter_recv$regenerating$interactions),
                   annotation_row = data.frame(row.names = rownames(ninter_send$healthy),
                                               "H" = ninter_send$healthy$interactions, 
                                               "R" = ninter_send$regenerating$interactions),
                   annotation_colors = list("H" = cols_hr,
                                            "R" = cols_n))
dev.off()
```

Compare detected interactions with condition DE genes (include: gene name of complex, if interaction is unique to condition, if elements of interaction are DE in that cell type) 

```{r}
cond_comparison_list = readRDS(file = "results/cond_effect/cond_comparison_list.RDS")

# get unique interactions
ints_regen = setdiff(unique(reform_list$regenerating[,"id_cp_interaction"]),
                     unique(reform_list$healthy[,"id_cp_interaction"]))
ints_healthy = setdiff(unique(reform_list$healthy[,"id_cp_interaction"]),
                       unique(reform_list$regenerating[,"id_cp_interaction"]))

# add genes from complexes
complex_ref = unique(rbind(dec_l$healthy[dec_l$healthy$is_complex=="True",c(1,5)], 
                             dec_l$regenerating[dec_l$regenerating$is_complex=="True",c(1,5)]))
inter_df = list()
for(n in c("healthy", "regenerating")){
  tmp = merge(reform_list[[n]], complex_ref, by.x = 4, by.y = 2, all.x = T)[,c(2,3,4,1,5,8,6,7)]
  inter_df[[n]] = merge(tmp, complex_ref, by.x = 5, by.y = 2, all.x = T)[,c(2,3,4,5,6,1,9,7,8)]
  inter_df[[n]]$gene_name.x[is.na(inter_df[[n]]$gene_name.x)] = inter_df[[n]]$lr1[is.na(inter_df[[n]]$gene_name.x)]
  inter_df[[n]]$gene_name.y[is.na(inter_df[[n]]$gene_name.y)] = inter_df[[n]]$lr2[is.na(inter_df[[n]]$gene_name.y)]
  colnames(inter_df[[n]])[c(5,7)] = c("gn1", "gn2")
}

# add info on DE genes
for(n in names(inter_df)){
  gn1_isDE = c()
  gn2_isDE = c()
  mod = ifelse(n=="healthy", 1, -1)
  for(i in 1:nrow(inter_df[[n]])){
    ct1cond = cond_comparison_list$healthy_v_regenerating$cell_type$celltype==inter_df[[n]]$ct1[i]
    ct2cond = cond_comparison_list$healthy_v_regenerating$cell_type$celltype==inter_df[[n]]$ct2[i]
    gn1cond = cond_comparison_list$healthy_v_regenerating$cell_type$gene==as.character(inter_df[[n]]$gn1[i])
    gn2cond = cond_comparison_list$healthy_v_regenerating$cell_type$gene==as.character(inter_df[[n]]$gn2[i])
    
    if(sum(ct1cond & gn1cond)>0){
      gn1_isDE = c(gn1_isDE, cond_comparison_list$healthy_v_regenerating$cell_type$avg_logFC[ct1cond & gn1cond]*mod>0)
    } else{
      gn1_isDE = c(gn1_isDE, FALSE)
    }
    if(sum(ct2cond & gn2cond)>0){
      gn2_isDE = c(gn2_isDE, cond_comparison_list$healthy_v_regenerating$cell_type$avg_logFC[ct2cond & gn2cond]*mod>0)
    } else{
      gn2_isDE = c(gn2_isDE, FALSE)
    }
  }
  inter_df[[n]]$gn1_isDE = gn1_isDE
  inter_df[[n]]$gn2_isDE = gn2_isDE
}

# add info on unique interactions
inter_df$healthy$isUnique = inter_df$healthy$id_cp_interaction %in% ints_healthy
inter_df$regenerating$isUnique = inter_df$regenerating$id_cp_interaction %in% ints_regen

saveRDS(inter_df, file = "results/cell_comm/interaction_de_unique_HvR.RDS")
```



```{r}
xxx = inter_df$healthy[(inter_df$healthy$gn1_isDE | inter_df$healthy$gn2_isDE) & inter_df$healthy$dir==1,c(1,2,3)]
xxx = unique(xxx)
pheatmap::pheatmap(table(xxx$ct1, xxx$ct2))
xxx = inter_df$regenerating[(inter_df$regenerating$gn1_isDE | inter_df$regenerating$gn2_isDE) & inter_df$regenerating$dir==1,c(1,2,3)]
xxx = unique(xxx)
pheatmap::pheatmap(table(xxx$ct1, xxx$ct2))


xxx = inter_df$healthy[inter_df$healthy$isUnique & inter_df$healthy$dir==1,c(1,2,3)]
xxx = unique(xxx)
pheatmap::pheatmap(table(xxx$ct1, xxx$ct2))
xxx = inter_df$regenerating[inter_df$regenerating$isUnique & inter_df$regenerating$dir==1,c(1,2,3)]
xxx = unique(xxx)
pheatmap::pheatmap(table(xxx$ct1, xxx$ct2))
```





# Pseudotime bins

To save the data, run `run_saveCPDBbinnedData.R`

## Running CellPhoneDB
Commands for runnin CellPhoneDB will be written here. Results are output to the `local1` disk to avoid I/O issues, but the output folders should then be copied to: `results/cell_comm_healthy/CellPhoneDB_runs`.  
NOTE: for some reason I'm getting a SegFault when trying to run the heatmap_plot command. This was run on the Euler server instead.

```{r}
for(n in unique(allcells_css@meta.data$Condition)){
  comm1 = paste0("cellphonedb method statistical_analysis results/cell_comm/CellPhoneDB_runs/", n, 
                 "_meta_names_bins.txt results/cell_comm/CellPhoneDB_runs/", n, 
                 "_exp_norm_bins.txt --threads=8 --output-path=/local1/USERS/tomasgomes/liver/CellPhoneDB_res --project-name ", n, 
                 "_bins --counts-data gene_name")
  comm2 = paste0("cp -r /local1/USERS/tomasgomes/liver/CellPhoneDB_res/", n, 
                 "_bins results/cell_comm/CellPhoneDB_runs/")
  comm3 = paste0("cellphonedb plot heatmap_plot --pvalues-path ./results/cell_comm/CellPhoneDB_runs/", n, 
                 "_bins/pvalues.txt --output-path ./results/cell_comm/CellPhoneDB_runs/", n, 
                 "_bins/ --count-name counts_heat.pdf --count-network-name count_net.txt --interaction-count-name count_inter.txt ./results/cell_comm/CellPhoneDB_runs/", n, 
                 "_meta_names_bins.txt")
  
  print(n)
  print(comm1)
  print(comm2)
  print(comm3)
  print(".")
}
```

## Analysis
Load pseudotime-varying genes

```{r}
reform_list = readRDS(file = "results/cell_comm/reform_list.RDS")
dec_l = readRDS(file = "results/cell_comm/deconvoluted_list.RDS")

# add genes from complexes
complex_ref = unique(rbind(dec_l$healthy[dec_l$healthy$is_complex=="True",c(1,5)], 
                             dec_l$regenerating[dec_l$regenerating$is_complex=="True",c(1,5)]))
inter_df = list()
for(n in c("healthy", "embolised", "regenerating")){
  tmp = merge(reform_list[[n]], complex_ref, by.x = 4, by.y = 2, all.x = T)[,c(2,3,4,1,5,8,6,7)]
  inter_df[[n]] = merge(tmp, complex_ref, by.x = 5, by.y = 2, all.x = T)[,c(2,3,4,5,6,1,9,7,8)]
  inter_df[[n]]$gene_name.x[is.na(inter_df[[n]]$gene_name.x)] = inter_df[[n]]$lr1[is.na(inter_df[[n]]$gene_name.x)]
  inter_df[[n]]$gene_name.y[is.na(inter_df[[n]]$gene_name.y)] = inter_df[[n]]$lr2[is.na(inter_df[[n]]$gene_name.y)]
  colnames(inter_df[[n]])[c(5,7)] = c("gn1", "gn2")
}
```



```{r}
load("results/zonation_cond/hep_binned_traj_qval_fits.RData")
hep_fits = fits_b_list
hep_traj = traj_list
hep_qvals = qvals_b_list

nbins = 100
df_h = data.table("zonation_pt" = hep_traj$healthy)
df_h$zonation_int = cut(df_h$zonation_pt, nbins)

df_e = data.table("zonation_pt" = hep_traj$embolised)
df_e = df_h[df_e, on="zonation_pt", roll=Inf, rollends = T]

df_r = data.table("zonation_pt" = hep_traj$regenerating)
df_r = df_h[df_r, on="zonation_pt", roll=Inf, rollends = T]

mean_bins_h = apply(hep_cells$healthy@assays$SCT@data, 1, 
                    function(x) tapply(x, df_h$zonation_int, mean))
mean_bins_e = apply(hep_cells$embolised@assays$SCT@data, 1, 
                    function(x) tapply(x, df_e$zonation_int, mean))
mean_bins_r = apply(hep_cells$regenerating@assays$SCT@data, 1, 
                    function(x) tapply(x, df_r$zonation_int, mean))

hep_h_max = apply(hep_fits$healthy, 2, function(x) mean(order(x, decreasing = T)[1:10]))
hep_e_max = apply(hep_fits$embolised, 2, function(x) mean(order(x, decreasing = T)[1:10]))
hep_r_max = apply(hep_fits$regenerating, 2, function(x) mean(order(x, decreasing = T)[1:10]))
std <- function(x) sd(x)/sqrt(length(x))
hep_h_sdmax = apply(hep_fits$healthy, 2, function(x) std(order(x, decreasing = T)[1:10]))
hep_e_sdmax = apply(hep_fits$embolised, 2, function(x) std(order(x, decreasing = T)[1:10]))
hep_r_sdmax = apply(hep_fits$regenerating, 2, function(x) std(order(x, decreasing = T)[1:10]))
```



```{r}
load("results/zonation_cond/end_binned_traj_qval_fits.RData")
end_fits = fits_b_list
end_traj = traj_list
end_qvals = qvals_b_list

nbins = 100
df_h = data.table("zonation_pt" = end_traj$healthy)
df_h$zonation_int = cut(df_h$zonation_pt, nbins)

df_e = data.table("zonation_pt" = end_traj$embolised)
df_e = df_h[df_e, on="zonation_pt", roll=Inf, rollends = T]

df_r = data.table("zonation_pt" = end_traj$regenerating)
df_r = df_h[df_r, on="zonation_pt", roll=Inf, rollends = T]

mean_bins_h = apply(end_cells$healthy@assays$SCT@data, 1, 
                    function(x) tapply(x, df_h$zonation_int, mean))
mean_bins_e = apply(end_cells$embolised@assays$SCT@data, 1, 
                    function(x) tapply(x, df_e$zonation_int, mean))
mean_bins_r = apply(end_cells$regenerating@assays$SCT@data, 1, 
                    function(x) tapply(x, df_r$zonation_int, mean))

end_h_max = apply(mean_bins_h, 2, function(x) mean(order(x, decreasing = T)[1:10]))
end_e_max = apply(mean_bins_e, 2, function(x) mean(order(x, decreasing = T)[1:10]))
end_r_max = apply(mean_bins_r, 2, function(x) mean(order(x, decreasing = T)[1:10]))
std <- function(x) sd(x)/sqrt(length(x))
end_h_sdmax = apply(mean_bins_h, 2, function(x) std(order(x, decreasing = T)[1:10]))
end_e_sdmax = apply(mean_bins_e, 2, function(x) std(order(x, decreasing = T)[1:10]))
end_r_sdmax = apply(mean_bins_r, 2, function(x) std(order(x, decreasing = T)[1:10]))
```



```{r}
hep_genes_list = lapply(inter_df, function(x) unique(c(as.character(x$gn1[grepl("Hepa", x$ct1)]), 
                                                       as.character(x$gn2[grepl("Hepa", x$ct2)]))))

hep_var_genes = names(hep_qvals$healthy)[hep_qvals$healthy<=0.05]

xxx = data.frame("cell_type" = c(inter_df$healthy$ct1[grepl("Hepa", inter_df$healthy$ct2) & inter_df$healthy$gn2 %in% hep_genes_list$healthy], inter_df$healthy$ct2[grepl("Hepa", inter_df$healthy$ct1) & inter_df$healthy$gn1 %in% hep_genes_list$healthy]),
           "gene_ct" = c(as.character(inter_df$healthy$gn1[grepl("Hepa", inter_df$healthy$ct2) & inter_df$healthy$gn2 %in% hep_genes_list$healthy]), as.character(inter_df$healthy$gn2[grepl("Hepa", inter_df$healthy$ct1) & inter_df$healthy$gn1 %in% hep_genes_list$healthy])),
           "gene_zon" = c(as.character(inter_df$healthy$gn2[grepl("Hepa", inter_df$healthy$ct2) & inter_df$healthy$gn2 %in% hep_genes_list$healthy]), as.character(inter_df$healthy$gn1[grepl("Hepa", inter_df$healthy$ct1) & inter_df$healthy$gn1 %in% hep_genes_list$healthy])),
           "interact" = c(as.character(inter_df$healthy$id_cp_interaction[grepl("Hepa", inter_df$healthy$ct2) & inter_df$healthy$gn2 %in% hep_genes_list$healthy]), as.character(inter_df$healthy$id_cp_interaction[grepl("Hepa", inter_df$healthy$ct1) & inter_df$healthy$gn1 %in% hep_genes_list$healthy])),
           "dir" = c(as.character(inter_df$healthy$dir[grepl("Hepa", inter_df$healthy$ct2) & inter_df$healthy$gn2 %in% hep_genes_list$healthy]),
                     as.character(inter_df$healthy$dir[grepl("Hepa", inter_df$healthy$ct1) & inter_df$healthy$gn1 %in% hep_genes_list$healthy])))

xxx = merge(xxx, data.frame(hep_h_max), by.x = 3, by.y = 0, all.x = T)
xxx = merge(xxx, data.frame(hep_h_sdmax), by.x = 1, by.y = 0, all.x = T)
xxx = merge(xxx, data.frame("range" = apply(hep_fits$healthy, 2, function(x) diff(range(x)))), 
            by.x = 1, by.y = 0, all.x = T)

plot_df = unique(xxx[!grepl("Hepa", xxx$cell_type) & xxx$hep_h_sdmax<=5 & 
                       xxx$range>=0.1 & xxx$gene_zon %in% hep_var_genes,-c(1,3)])

plt = ggplot(plot_df, aes(x = hep_h_max, colour = cell_type))+
  geom_density()+
  scale_x_continuous(limits = c(0,100))+
  theme_bw()

plot(mean_bins_h[,"VEGFA"], pch = 19, cex = 0.85)
plot(hep_fits$healthy[,"PROS1"], pch = 19, cex = 0.85)
```



```{r}
inter_df = readRDS(file = "results/cell_comm/interaction_de_unique_HvR.RDS")
inter_hr = rbind(inter_df$healthy[,1:9], inter_df$regenerating[,1:9])
inter_hr$cond = c(rep("healthy", nrow(inter_df$healthy)),
                  rep("regenerating", nrow(inter_df$regenerating)))
inter_hr = inter_hr[grepl("LSEC_", inter_hr$ct1) | grepl("Hepa", inter_hr$ct1) |
                      grepl("LSEC_", inter_hr$ct2) | grepl("Hepa", inter_hr$ct2), ]

inter_diff = inter_hr[inter_hr$ct1!=inter_hr$ct2,]
```


```{r, fig.height=5, fig.width=5.2}
hep_genes_h_0 = unique(c(as.character(inter_diff$gn1)[grepl("Hepa", inter_diff$ct1) &
                                                      inter_diff$cond=="healthy"],
                         as.character(inter_diff$gn2)[grepl("Hepa", inter_diff$ct2) &
                                                      inter_diff$cond=="healthy"]))
hep_genes_r_0 = unique(c(as.character(inter_diff$gn1)[grepl("Hepa", inter_diff$ct1) &
                                                      inter_diff$cond=="regenerating"],
                         as.character(inter_diff$gn2)[grepl("Hepa", inter_diff$ct2) &
                                                      inter_diff$cond=="regenerating"]))

nbins = 100
df_h = data.table("zonation_pt" = hep_traj$healthy)
df_h$zonation_int = cut(df_h$zonation_pt, nbins)

df_r = data.table("zonation_pt" = hep_traj$regenerating)
df_r = df_h[df_r, on="zonation_pt", roll=Inf, rollends = T]

mean_bins_h = apply(hep_cells$healthy@assays$SCT@data, 1, 
                    function(x) tapply(x, df_h$zonation_int, mean))
mean_bins_r = apply(hep_cells$regenerating@assays$SCT@data, 1, 
                    function(x) tapply(x, df_r$zonation_int, mean))

hep_h_max = apply(mean_bins_h, 2, function(x) mean(order(x, decreasing = T)[1:10]))
hep_r_max = apply(mean_bins_r, 2, function(x) mean(order(x, decreasing = T)[1:10]))

xxx = data.frame(row.names = hep_genes_h_0, "h" = hep_h_max[hep_genes_h_0], 
                 "r" = hep_r_max[hep_genes_h_0])
xxx = xxx[complete.cases(xxx),]
ggplot(xxx, aes(x = h, y = r))+geom_hex()+theme_bw()+theme(aspect.ratio = 1)
cor(xxx)

par(mfrow = c(2,2))
plot(mean_bins_h[,"COL18A1"], pch = 19, cex = 0.85)
plot(mean_bins_r[,"COL18A1"], pch = 19, cex = 0.85)
plot(hep_fits$healthy[,"COL18A1"], pch = 19, cex = 0.85)
plot(hep_fits$regenerating[,"COL18A1"], pch = 19, cex = 0.85)
```

```{r, fig.height=5, fig.width=5.2}
end_genes_h = unique(c(as.character(inter_diff$gn1)[grepl("LSEC_", inter_diff$ct1) &
                                                        inter_hr$cond=="healthy"],
                         as.character(inter_diff$gn2)[grepl("LSEC_", inter_diff$ct2) &
                                                        inter_diff$cond=="healthy"]))
end_genes_r = unique(c(as.character(inter_diff$gn1)[grepl("LSEC_", inter_diff$ct1) &
                                                      inter_hr$cond=="regenerating"],
                       as.character(inter_diff$gn2)[grepl("LSEC_", inter_diff$ct2) &
                                                      inter_diff$cond=="regenerating"]))

nbins = 100
df_h = data.table("zonation_pt" = end_traj$healthy)
df_h$zonation_int = cut(df_h$zonation_pt, nbins)

df_r = data.table("zonation_pt" = end_traj$regenerating)
df_r = df_h[df_r, on="zonation_pt", roll=Inf, rollends = T]

end_mean_bins_h = apply(end_cells$healthy@assays$SCT@data, 1, 
                    function(x) tapply(x, df_h$zonation_int, mean))
end_mean_bins_r = apply(end_cells$regenerating@assays$SCT@data, 1, 
                    function(x) tapply(x, df_r$zonation_int, mean))

end_h_max = apply(end_mean_bins_h, 2, function(x) mean(order(x, decreasing = T)[1:10]))
end_r_max = apply(end_mean_bins_r, 2, function(x) mean(order(x, decreasing = T)[1:10]))

xxx = data.frame(row.names = end_genes_h, "h" = end_h_max[end_genes_h], 
                 "r" = end_r_max[end_genes_h])
xxx = xxx[complete.cases(xxx),]
ggplot(xxx, aes(x = h, y = r))+geom_hex()+theme_bw()+theme(aspect.ratio = 1)
cor(xxx)

par(mfrow = c(2,2))
plot(end_mean_bins_h[,"PECAM1"], pch = 19, cex = 0.85)
plot(end_mean_bins_r[,"PECAM1"], pch = 19, cex = 0.85)
plot(end_fits$healthy[,"PECAM1"], pch = 19, cex = 0.85)
plot(end_fits$regenerating[,"PECAM1"], pch = 19, cex = 0.85)
```



```{r}
ct1 = inter_diff[grepl("Hepa", inter_diff$ct1),
                 c("gn1", "ct2")]
ct2 = inter_diff[grepl("Hepa", inter_diff$ct2),
                 c("gn2", "ct1")]
ct = unique(rbind(as.matrix(ct1), as.matrix(ct2)))

ctg = merge(xxx, ct, by.x = 0, by.y = 1)

ggplot(ctg, aes(x = h, y = r))+geom_hex()+theme_bw()+theme(aspect.ratio = 1)
ggplot(ctg, aes(x = ct2, y = h, fill = ct2))+
  geom_boxplot()+
  theme_bw()+theme(aspect.ratio = 1)


ct1 = inter_diff[grepl("LSEC_", inter_diff$ct1),
                 c("gn1", "ct2")]
ct2 = inter_diff[grepl("LSEC_", inter_diff$ct2),
                 c("gn2", "ct1")]
ct = unique(rbind(as.matrix(ct1), as.matrix(ct2)))

ctg = merge(xxx, ct, by.x = 0, by.y = 1)

ggplot(ctg, aes(x = h, y = r))+geom_hex()+theme_bw()+theme(aspect.ratio = 1)
ggplot(ctg, aes(x = ct2, y = h, fill = ct2))+
  geom_boxplot()+
  theme_bw()+theme(aspect.ratio = 1)
```






