---
title: "Condition effects and gene signatures"
output: html_notebook
---



# General Setup
Setup chunk

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_knit$get("root.dir")
```

Setup reticulate

```{r}
library(reticulate)
knitr::knit_engines$set(python = reticulate::eng_python)
py_available(initialize = FALSE)
use_python(Sys.which("python"))
py_config()
```

Load libraries

```{r}
library(Seurat)
library(ggplot2)
library(ggrepel)
library(ggridges)
library(dplyr)
library(seqgendiff)
library(clusterProfiler)
library(ReactomePA)
library(org.Hs.eg.db)
```

Other functions

```{r}
# downsample taking the minimum median of UMI counts as a reference
dsSeurat = function(srat, variable, assay = "SCT", numis = "nCount_SCT", seed = 1){
  # input Seurat object and variable to do the downsample based on
  # return Seurat object
  
  set.seed(seed)
  
  # downsample taking the minimum median of UMI counts as a reference
  fact = srat@meta.data[,numis]/min(tapply(srat@meta.data[,numis], srat@meta.data[,variable], median))
  fact = ifelse(fact<=1, 0, log2(fact))
  thinned_counts = seqgendiff::thin_lib(as.matrix(srat@assays[[assay]]@counts), 
                                        thinlog2 = fact, relative = T)
  rownames(thinned_counts$mat) = rownames(srat@assays[[assay]]@data)
  colnames(thinned_counts$mat) = colnames(srat@assays[[assay]]@data)
  srat@assays$SCT@counts = Matrix::Matrix(thinned_counts$mat, sparse = T)
  srat@assays$SCT@data = log1p(srat@assays[[assay]]@counts)
  
  return(srat)
}

rpfilter = function(x){
  return(!grepl("^RPL", x$geneID) & !grepl("^RPS", x$geneID) & 
           !grepl("/RPL", x$geneID, fixed = T) & !grepl("/RPS", x$geneID, fixed = T))
}
```



# Load and preprocess data
Load data (from all cells)

```{r}
allcells_css = readRDS(file = "data/processed/allcells_css.RDS")
end_cells = readRDS(file = "results/zonation_cond/end_cells_zonation_rank.RDS")
hep_cells = readRDS(file = "results/zonation_cond/hep_cells_zonation_rank.RDS")
```

Load chromosome information (important to add to tables)

```{r}
chr_tab = read.csv("../../gene_refs/human/biomart_ens_gn_hgnc_chr.csv", 
                   header = T, stringsAsFactors = F)
```

Add columns where LSEC and Hepatocytes are split by zonation

```{r}
load("results/zonation_cond/hep_binnedDonor_traj_qval_fits_rank.RData")
sig_genes = names(qvals_db_list$healthy)[order(qvals_db_list$healthy, decreasing = F)[1:100]]
bins = cut(hep_cells$embolised$zonation_pt, 100)
m_bs = apply(hep_cells$embolised@assays$SCT@data[sig_genes,], 1, function(x) tapply(x, bins, mean))
pca = prcomp(m_bs, center = T)
plot(pca$x[,1:2], cex = 1:100/100)
sum(pca$x[,2]>0)
sum(pca$x[,2]<0 & pca$x[,1]<0)
sum(pca$x[,2]<0 & pca$x[,1]>0)
```



# Cell Type markers
Calculate cell type markers

```{r}
cell_type_mk = list()
meta = data.frame(row.names = rownames(allcells_css@meta.data),
                  "cell_type" = as.character(allcells_css@meta.data$allcells_simp),
                  stringsAsFactors = F)

meta[grepl("Hepatocytes ", meta$cell_type), "cell_type"] = "Hepatocytes" 

# remove contaminating cells and interaction/doublets
meta = subset(meta, !(cell_type %in% c("Macrophage-T cell interaction",
                                       "Macrophage-Endothelial interaction",
                                       "Lymphoid-Endothelial interaction",
                                       #"LSEC periportal", "LSEC pericentral",
                                       "Kupffer-Endothelial interaction",
                                       "Hepatocytes (portal/midzone)",
                                       "Hepatocytes (portal)","Hepatocytes (midzone)",
                                       "Hepatocytes (central/detox)", "Hepatocytes (central)",
                                       "Hepatocyte-T cell interaction",
                                       "Hepatocyte-Myeloid interaction",
                                       "Hepatocyte-Endothelial interaction",
                                       "Dividing cells")))

allcells_css = AddMetaData(allcells_css, meta)
allcells_css = SetIdent(allcells_css, value = "cell_type")
cell_type_mk[["good"]] = FindAllMarkers(allcells_css[,!is.na(allcells_css@meta.data$cell_type)], 
                                        assay = "SCT", test.use = "wilcox",
                                        pseudocount.use = 0.1, logfc.threshold = 0.2,
                                        max.cells.per.ident = 10000,
                                        min.cells.feature = 10, verbose = T)

# using major labels
plot_df = data.frame(row.names = colnames(allcells_css),
                     "major_ct" = as.character(allcells_css@meta.data$allcells_major),
                     "rem" = as.character(allcells_css@meta.data$allcells_major),
                     stringsAsFactors = F)
plot_df$major_ct[plot_df$major_ct=="T cells" | 
                      plot_df$major_ct=="B cells" | 
                      plot_df$major_ct=="Macrophages" | 
                      plot_df$major_ct=="DCs"] = "Immune"
plot_df$major_ct[plot_df$major_ct=="Endothelial cells"] = "Endothelial"
plot_df$major_ct[plot_df$major_ct=="Stellate cells"] = "Mesenchymal"
plot_df = plot_df[plot_df$major_ct!="Doublets",]
plot_df = plot_df[plot_df$major_ct!="Dividing cells",]

allcells_css = AddMetaData(allcells_css, plot_df)

allcells_css = SetIdent(allcells_css, value = "major_ct")
cell_type_mk[["major"]] = FindAllMarkers(allcells_css[,!is.na(allcells_css@meta.data$major_ct)], 
                                        assay = "SCT", test.use = "wilcox",
                                        pseudocount.use = 0.1, logfc.threshold = 0,
                                        max.cells.per.ident = 10000,
                                        min.cells.feature = 10, verbose = T)

saveRDS(cell_type_mk, file = "results/cond_effect/cell_type_mk.RDS")
```

Intersection of number of genes in top 250

```{r, fig.height=4, fig.width=5}
top_genes = list()
for(n in names(cell_type_mk)){
  top_genes[[n]] = list()
  for(ct in unique(cell_type_mk[[n]]$cluster)){
    top_genes[[n]][[ct]] = cell_type_mk[[n]][cell_type_mk[[n]]$cluster==ct & 
                                               cell_type_mk[[n]]$p_val_adj<=0.05 & 
                                               cell_type_mk[[n]]$avg_logFC>0,]
    top_genes[[n]][[ct]] = top_genes[[n]][[ct]][order(top_genes[[n]][[ct]]$avg_logFC, 
                                                      decreasing = T),"gene"][1:250]
  }
}
names(top_genes) = c("cell_type","major_ct")

saveRDS(top_genes, file = "results/cond_effect/top_genes_mk.RDS")
```

Save allcells_css again with the new annotations

```{r}
saveRDS(allcells_css, file = "data/processed/allcells_css_reannot.RDS")
```



# DE per cell type
## Condition DE per cell type
Calculate DE (only for cell types with more than 10 cells in one condition)

```{r}
comb_cond = combn(unique(allcells_css@meta.data$Condition), 2)
cond_comparison_list = list()
for(i in 1:ncol(comb_cond)){
  comp = paste0(comb_cond[1,i], "_v_", comb_cond[2,i])
  cond_comparison_list[[comp]] = list()
  for(g in c("cell_type", "major_ct")){
    print(g)
    tab_cond = table(allcells_css@meta.data[,g], allcells_css@meta.data$Condition)
    tab_cond = tab_cond[apply(tab_cond, 1, function(x) all(x>10)),]
    for(ct in rownames(tab_cond)){
      cat(ct)
      allcells_css = SetIdent(allcells_css, value = g)
      mk = FindMarkers(allcells_css[,!is.na(allcells_css@meta.data[,g])], 
                       ident.1 = comb_cond[1,i], ident.2 = comb_cond[2,i],
                       group.by = "Condition", subset.ident = ct, pseudocount.use = 0.1, 
                       logfc.threshold = 0, min.cells.feature = 10,
                       assay = "SCT", max.cells.per.ident = 10000)
      mk$gene = rownames(mk)
      mk$group = g
      mk$celltype = ct
      mk$condition = ifelse(mk$avg_logFC<0, comb_cond[2,i], comb_cond[1,i])
      mk = mk[,c("group", "celltype", "gene", "avg_logFC", 
                 "p_val_adj", "condition", "pct.1", "pct.2")]
      
      if(comp %in% names(cond_comparison_list)){
        cond_comparison_list[[comp]][[g]] = rbind(cond_comparison_list[[comp]][[g]], mk)
      } else{
        cond_comparison_list[[comp]][[g]] = mk
      }
    }
    # add chromosome infor
    cond_comparison_list[[comp]][[g]] = merge(cond_comparison_list[[comp]][[g]], 
                                              chr_tab[,c(2,4)], by.x = "gene", by.y = 1, all.x = T)
  }
}
saveRDS(cond_comparison_list, file = "results/cond_effect/cond_comparison_list.RDS")
```

Repeat for heaptocytes, but use downsampling

```{r}
comb_cond = combn(unique(allcells_css@meta.data$Condition), 2)
cond_comparison_hepsub = list()
for(i in 1:ncol(comb_cond)){
  comp = paste0(comb_cond[1,i], "_v_", comb_cond[2,i])
  cond_comparison_hepsub[[comp]] = list()
  for(g in c("cell_type", "major_ct")){
    print(g)
    tab_cond = table(allcells_css@meta.data[,g], allcells_css@meta.data$Condition)
    tab_cond = rownames(tab_cond)[apply(tab_cond, 1, function(x) all(x>10)) & 
                          grepl("Hepato", rownames(tab_cond)) &
                          !grepl("interact", rownames(tab_cond))]
    for(ct in tab_cond){
      cat(ct)
      hep_cells = allcells_css[,allcells_css@meta.data[,g]==ct &
                                 !is.na(allcells_css@meta.data[,g])]
      
      hep_cells = dsSeurat(hep_cells, "Condition")
      
      hep_cells = SetIdent(hep_cells, value = g)
      mk = FindMarkers(hep_cells[,!is.na(hep_cells@meta.data[,g])], 
                       ident.1 = comb_cond[1,i], ident.2 = comb_cond[2,i],
                       group.by = "Condition", subset.ident = ct, pseudocount.use = 0.1, 
                       logfc.threshold = 0, min.cells.feature = 10,
                       assay = "SCT", max.cells.per.ident = 10000)
      mk$gene = rownames(mk)
      mk$group = g
      mk$celltype = ct
      mk$condition = ifelse(mk$avg_logFC<0, comb_cond[2,i], comb_cond[1,i])
      mk = mk[,c("group", "celltype", "gene", "avg_logFC", 
                 "p_val_adj", "condition", "pct.1", "pct.2")]
      
      if(comp %in% names(cond_comparison_hepsub)){
        cond_comparison_hepsub[[comp]][[g]] = rbind(cond_comparison_hepsub[[comp]][[g]], mk)
      } else{
        cond_comparison_hepsub[[comp]][[g]] = mk
      }
    }
    # add chromosome infor
    cond_comparison_hepsub[[comp]][[g]] = merge(cond_comparison_hepsub[[comp]][[g]] , 
                                                chr_tab[,c(2,4)], by.x = "gene", by.y = 1, all.x = T)
  }
}
saveRDS(cond_comparison_hepsub, file = "results/cond_effect/cond_comparison_hepsub.RDS")
```

Add chromosome information

```{r}
chr_tab = read.csv("../../gene_refs/human/biomart_ens_gn_hgnc_chr.csv", 
                   header = T, stringsAsFactors = F)

cond_comp_chr = list()
for(n in names(cond_comparison_list)){
  cond_comp_chr[[n]] = list()
  for(i in names(cond_comparison_list[[n]])){
    cond_comp_chr[[n]][[i]] = merge(cond_comparison_list[[n]][[i]], 
                                    chr_tab[,c(2,4)], by.x = "gene", by.y = 1, all.x = T)
  }
}
saveRDS(cond_comp_chr, file = "results/cond_effect/cond_comp_chr.RDS")

cond_comp_hep_chr = list()
for(n in names(cond_comparison_hepsub)){
  cond_comp_hep_chr[[n]] = list()
  for(i in names(cond_comparison_hepsub[[n]])){
    cond_comp_hep_chr[[n]][[i]] = merge(cond_comparison_hepsub[[n]][[i]], 
                                    chr_tab[,c(2,4)], by.x = "gene", by.y = 1, all.x = T)
  }
}
saveRDS(cond_comp_hep_chr, file = "results/cond_effect/cond_comp_hep_chr.RDS")
```

Clean DE tables - remove Y chromosome genes and markers of different cell types

```{r}
cond_comparison_list_filt = list()
for(g in names(cond_comparison_list[[1]])){
  cond_comparison_list_filt[[g]] = list()
  for(n in names(cond_comparison_list)){
    tg = top_genes[[g]]
    df_cond_de = cond_comparison_list[[n]][[g]]
    for(ct in names(tg)){
      okgenes = df_cond_de[df_cond_de$celltype==ct,"gene"]
      okgenes = okgenes[!(okgenes %in% unlist(tg[-which(names(tg)==ct)]) &
                            !(okgenes %in% tg[[ct]]))]
      df_cond_de = df_cond_de[df_cond_de$celltype!=ct | 
                                (df_cond_de$celltype==ct &
                                   df_cond_de$gene %in% okgenes),]
    }
    df_cond_de = df_cond_de[!grepl("CHR_", df_cond_de$Chromosome.scaffold.name) & 
                              !(df_cond_de$Chromosome.scaffold.name %in% c("Y","KI270734.1")),]
    df_cond_de = df_cond_de[df_cond_de$celltype %in% names(tg),]
    cond_comparison_list_filt[[g]][[n]] = df_cond_de
  }
}

cond_comparison_hepsub_filt = list()
for(g in names(cond_comparison_hepsub[[1]])){
  cond_comparison_hepsub_filt[[g]] = list()
  for(n in names(cond_comparison_hepsub)){
    tg = top_genes[[g]]
    df_cond_de = cond_comparison_hepsub[[n]][[g]]
    for(ct in names(tg)){
      okgenes = df_cond_de[df_cond_de$celltype==ct,"gene"]
      okgenes = okgenes[!(okgenes %in% unlist(tg[-which(names(tg)==ct)]) &
                            !(okgenes %in% tg[[ct]]))]
      df_cond_de = df_cond_de[df_cond_de$celltype!=ct | 
                                (df_cond_de$celltype==ct &
                                   df_cond_de$gene %in% okgenes),]
    }
    df_cond_de = df_cond_de[!grepl("CHR_", df_cond_de$Chromosome.scaffold.name) & 
                              !(df_cond_de$Chromosome.scaffold.name %in% c("Y","KI270734.1")),]
    df_cond_de = df_cond_de[df_cond_de$celltype %in% names(tg),]
    cond_comparison_hepsub_filt[[g]][[n]] = df_cond_de
  }
}

filt_comps = list("cell_type" = cond_comparison_list_filt$cell_type, 
                  "major_ct" = cond_comparison_list_filt$major_ct, 
                  "cell_type_hep" = cond_comparison_hepsub_filt$cell_type,
                  "major_ct_hep" = cond_comparison_hepsub_filt$major_ct)

saveRDS(filt_comps, file = "results/cond_effect/all_filt_cond_comps.RDS")
```


## Donor DE per cell type
Donor markers across the whole data

```{r}
allcells_css = SetIdent(allcells_css, value = "Donor")
donor_markers = FindAllMarkers(allcells_css[,!is.na(allcells_css@meta.data$cell_type)], 
                               assay = "SCT", test.use = "wilcox",
                               pseudocount.use = 0.1, logfc.threshold = 0,
                               max.cells.per.ident = 10000,
                               min.cells.feature = 10, verbose = T)
```

Donor markers per cell type

```{r}
donor_submk_list = list()
allcells_css = SetIdent(allcells_css, value = "Donor")
for(g in c("cell_type", "major_ct", "Condition")){
  print(g)
  donor_submk_list[[g]] = list()
  tab_cond = table(allcells_css@meta.data[,g], allcells_css@meta.data$Donor)
  if(g!="Condition"){
    tab_cond = tab_cond[apply(tab_cond, 1, function(x) all(x>5)),]
  }
  for(ct in rownames(tab_cond)){
    cat(ct)
    mk = FindAllMarkers(allcells_css[,allcells_css@meta.data[,g]==ct], 
                        pseudocount.use = 0.1, 
                        logfc.threshold = 0, min.cells.feature = 5,
                        assay = "SCT", max.cells.per.ident = 10000)
    mk$gene = rownames(mk)
    mk$class = g
    mk$group = ct
    mk = mk[,c("class", "group", "gene", "cluster", "avg_logFC", 
               "p_val_adj", "pct.1", "pct.2")]
    
    donor_submk_list[[g]][[ct]] = mk
    
    # add chromosome infor
    donor_submk_list[[g]][[ct]] = merge(donor_submk_list[[g]][[ct]], chr_tab[,c(2,4)], 
                                        by.x = "gene", by.y = 1, all.x = T)
  }
}

donor_submk_list$all = donor_markers
saveRDS(donor_submk_list, file = "results/cond_effect/donor_comparison_list.RDS")
```



# Explore results
Reload data

```{r}
allcells_css = readRDS(file = "data/processed/allcells_css_reannot.RDS")
filt_comps = readRDS(file = "results/cond_effect/all_filt_cond_comps.RDS")
```

Get GO Term enrichment

```{r}
fc = 0.3
go_enr_list = list()
for(gr in names(filt_comps)){
  filt_cond_comp_chr = filt_comps[[gr]]
  gr_g = ifelse(endsWith(gr, "hep"), paste0(strsplit(gr, "_")[[1]][1:2], collapse="_"), gr)
  allcells_ct = allcells_css[,!is.na(allcells_css@meta.data[,gr_g])]
  go_enr_list[[gr]] = list()
  for(comp in names(filt_cond_comp_chr)){
    testdf = filt_cond_comp_chr[[comp]][filt_cond_comp_chr[[comp]]$p_val_adj<=0.05,]
    
    ncond = strsplit(comp, "_v_")[[1]]
    go_enr_list[[gr]][[comp]] = list()
    for(ct in unique(testdf$celltype)){
      subtestdf = unique(testdf[testdf$celltype==ct, c("avg_logFC","gene")])
      
      # get DE tested genes
      cells = allcells_ct@meta.data[,gr_g]==ct
      genes = Matrix::rowSums(allcells_ct[,cells]@assays$SCT@data>0)
      genes = genes[genes>(0.1*sum(cells))] # might not include all, but very close
      genes = replace(genes, genes>0, 0)
      genes[subtestdf$gene] = subtestdf$avg_logFC
      
      eg_all = bitr(names(genes), fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
      
      # GO enrichment
      for(cond in ncond){
        fsel = if(cond==ncond[[1]]){
          function(x) return(x>fc)
          } else{
            function(x) return(x<(-fc))
          }
        
        if(length(names(genes)[fsel(genes)])>2){
          eg = bitr(names(genes)[fsel(genes)], fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
          
          # enrichment
          reac = enrichPathway(gene=eg$ENTREZID,pvalueCutoff=0.05, readable=T, universe = eg_all$ENTREZID)
          got = enrichGO(gene = eg$ENTREZID, universe = eg_all$ENTREZID, OrgDb = org.Hs.eg.db, readable = T,
                         ont = "BP", pvalueCutoff  = 1, qvalueCutoff  = 1, pAdjustMethod = "BH")
          hf = "/links/groups/treutlein/USERS/tomasgomes/gene_refs/human/h.all.v7.2.entrez.gmt"
          egmt = enricher(gene = eg$ENTREZID, universe = eg_all$ENTREZID, TERM2GENE=read.gmt(hf),
                          pvalueCutoff = 1, qvalueCutoff = 1, pAdjustMethod = "BH",
                          minGSSize = 0, maxGSSize = 1000000)
          if(!is.null(egmt)){
            egmt = egmt@result
            egmt$geneID = lapply(egmt$geneID, 
                                 function(x) unlist(paste0(bitr(unlist(strsplit(x, "/")),
                                                                fromType="ENTREZID", toType="SYMBOL",
                                                                OrgDb="org.Hs.eg.db")[,2],
                                                           collapse = "/")))
          }
          else{
            egmt = data.frame()
          }
          
          all_terms = rbind(as.data.frame(reac), as.data.frame(got), egmt)
          if(nrow(all_terms)>0){
            all_terms$DB = c(rep("Reactome", nrow(as.data.frame(reac))), 
                             rep("GO Term", nrow(as.data.frame(got))),
                             rep("Hallmark", nrow(egmt)))
            #all_terms = all_terms[all_terms$qvalue<=0.05,]
            all_terms$cond = cond
            
            if(is.null(go_enr_list[[gr]][[comp]][[ct]])){
              go_enr_list[[gr]][[comp]][[ct]] = all_terms
            } else{
              go_enr_list[[gr]][[comp]][[ct]] = rbind(go_enr_list[[gr]][[comp]][[ct]], all_terms)
            }
          }
        }
      }
    }
  }
  print(names(go_enr_list))
  print(names(go_enr_list[[gr]]))
  print(names(go_enr_list[[gr]][[comp]]))
}
saveRDS(go_enr_list, file = "results/cond_effect/go_enr_list.RDS")
```

Tables of DE genes

```{r}
for(g in names(filt_comps)){
  for(comp in names(filt_comps[[g]])){
    subdir = file.path("to_send", g)
    dir.create(subdir)
    write.table(filt_comps[[g]][[comp]], 
                file = paste0(subdir, "/DEconditions_", comp, "_", g, ".csv"), 
                col.names = T, row.names = F, quote = F, sep = "\t")
  }
}

for(g in names(go_enr_list)){
  for(comp in names(go_enr_list[[g]])){
    for(ct in names(go_enr_list[[g]][[comp]])){
      subdir = file.path("to_send/", paste0(g, "/GO_", ct))
      dir.create(subdir)
      write.table(go_enr_list[[g]][[comp]][[ct]], 
                  file = paste0(subdir, "/GOTermBP_", comp, "_", ct, "_", g, ".csv"), 
                  col.names = T, row.names = F, quote = F, sep = "\t")
    }
  }
}
```

Exploring transcription factors and GO Terms

```{r}
tfdb = read.table("../../gene_refs/human/Homo_sapiens_TF_AnimalTFDB3.txt", 
                  sep = "\t", header = T, stringsAsFactors = F)

genes_to_plot = list("Hepatocytes" = c("HAMP", "SAA1", "FGA", "G0S2", "TNFRSF1A",
                                       "CYP3A7", "MT1A", "DDX21", "IL1RAP", "MT1X",
                                       "MGLL", "MFSD2A", "APOA4", "IRF7", "CEBPA"), 
                      "Immune" = c("S100A9", "CD81", "FCGR1A", "GBP5", "RUNX3", 
                                   "CD300E", "MNDA", "IL18BP", "IL4I1", "WDFY4",
                                   "TRDC", "TRBC1", "IFNG", "CD160", "IFITM1"), 
                      "Cholangiocytes" = c("CCL2", "IRF1", "KRT19", "CCL20", "ICAM1",
                                           "MAP3K12", "EPSTI1", "P4HA1", "PEAK1", "SEC24A",
                                           "LCN2", "RAMP1", "ITGB4", "BIK", "CCL28"), 
                      "Mesenchymal" = c(), 
                      "Endothelial" = c("SOX18", "PLCG2", "TFPI2", "KLF2", "STC1",
                                        "CXXC5", "NPR3", "CDK6", "ADGRG6", "PDGFB",
                                        "RFPL1", "ITGB3", "ADAMTS4", "KLF13", "GPR182"))

plt_list_cond = list()
df_list_cond = list()
for(ct in unique(filt_comps$major_ct$healthy_v_embolised$celltype)){
  g = if(ct=="Hepatocytes") "major_ct_hep" else "major_ct"
  
  plt_reg = filt_comps[[g]]$healthy_v_regenerating[,c(1,3,4,5)]
  plt_reg = unique(plt_reg[plt_reg$celltype==ct,])
  plt_emb = filt_comps[[g]]$healthy_v_embolised[,c(1,3,4,5)]
  plt_emb = unique(plt_emb[plt_emb$celltype==ct,])
  rownames(plt_reg) = plt_reg$gene
  rownames(plt_emb) = plt_emb$gene
  
  plt_df = merge(plt_emb, plt_reg, by = 0, all = T)
  rownames(plt_df) = plt_df[,1]
  plt_df = plt_df[,-c(1,3,6,7)]
  colnames(plt_df) = c("gene", "FC_emb", "pval_emb", "FC_reg", "pval_reg")
  plt_df$gene = rownames(plt_df)
  plt_df$FC_emb[is.na(plt_df$FC_emb)] = 0
  plt_df$FC_reg[is.na(plt_df$FC_reg)] = 0
  plt_df$pval_emb[is.na(plt_df$pval_emb)] = 1
  plt_df$pval_reg[is.na(plt_df$pval_reg)] = 1
  
  # condition labels
  plt_df$cond = ifelse(plt_df$FC_emb<=(-0.2) & plt_df$pval_emb<=0.05 & plt_df$FC_reg>=0, "embolised",
                       ifelse(plt_df$FC_reg<=(-0.2) & plt_df$pval_reg<=0.05 & plt_df$FC_emb>=0, "regenerating",
                              ifelse(plt_df$FC_reg<=(-0.2) & plt_df$pval_reg<=0.05 & plt_df$FC_emb<=(-0.2) &
                                       plt_df$pval_emb<=0.05, "both","other")))
  plt_df$cond = factor(plt_df$cond, levels = c("both", "embolised", "regenerating", "other"))
  plt_df = plt_df[order(plt_df$cond, decreasing = T),]
  
  # genes from GO
  if(!is.null(go_enr_list[[g]]$healthy_v_embolised[[ct]])){
    t_emb = go_enr_list[[g]]$healthy_v_embolised[[ct]]
    t_emb = t_emb[rpfilter(t_emb) & t_emb$cond=="embolised" & t_emb$DB=="GO Term",]
    t_emb_genes = strsplit(as.character(t_emb$geneID), "/")
    names(t_emb_genes) = t_emb$Description
    t_emb_genes = reshape2::melt(t_emb_genes)
    emb_ug = rownames(plt_df)[plt_df$FC_emb<=(-0.2) & plt_df$pval_emb<=0.05 & plt_df$FC_reg>=0]
    emb_genes_go = unique(t_emb_genes[t_emb_genes$value %in% emb_ug,1])
    
    t_reg = go_enr_list[[g]]$healthy_v_regenerating[[ct]]
    t_reg = t_reg[rpfilter(t_reg) & t_reg$cond=="regenerating" & t_reg$DB=="GO Term",]
    t_reg_genes = strsplit(as.character(t_reg$geneID), "/")
    names(t_reg_genes) = t_reg$Description
    t_reg_genes = reshape2::melt(t_reg_genes)
    reg_ug = rownames(plt_df)[plt_df$FC_reg<=(-0.2) & plt_df$pval_reg<=0.05 & plt_df$FC_emb>=0]
    reg_genes_go = unique(t_reg_genes[t_reg_genes$value %in% reg_ug,1])
    
    t_both = rbind(t_emb, t_reg)
    t_both_genes = strsplit(as.character(t_both$geneID), "/")
    names(t_both_genes) = t_both$Description
    t_both_genes = reshape2::melt(t_both_genes)
    both_ug = rownames(plt_df)[plt_df$FC_reg<0 & plt_df$pval_reg<=0.05 & 
                                 plt_df$FC_emb<0 & plt_df$pval_emb<=0.05]
    both_genes_go = unique(t_both_genes[t_both_genes$value %in% both_ug,1])
    
      plt_df$cond_go = ifelse(plt_df$gene %in% emb_genes_go, "embolised", 
                           ifelse(plt_df$gene %in% reg_genes_go, "regenerating", 
                                  ifelse(plt_df$gene %in% both_genes_go, "both", "other")))
      plt_df$cond_go = factor(plt_df$cond_go, levels = c("both", "embolised", "regenerating", "other"))
  }
  
  plt_df$isTF = plt_df$gene %in% tfdb$Symbol
  
  emb_g_plt =  plt_df$gene[order(plt_df$FC_emb, decreasing = F)][plt_df$cond[order(plt_df$FC_emb, decreasing = F)]=="embolised"][1:5]
  reg_g_plt = plt_df$gene[order(plt_df$FC_reg, decreasing = F)][plt_df$cond[order(plt_df$FC_reg, decreasing = F)]=="regenerating"][1:5]
  b_g_plt = plt_df$gene[order(plt_df$FC_reg+plt_df$FC_emb, decreasing = F)][plt_df$cond[order(plt_df$FC_reg+plt_df$FC_emb, decreasing = F)]=="both"][1:5]
  
  cols = c("both" = "#C54635", "regenerating" = "salmon", "embolised" = "darkred", "other" = "grey85")
  plt = ggplot(plt_df, aes(x = FC_emb, y = FC_reg, colour = cond))+
    geom_vline(xintercept = 0)+
    geom_hline(yintercept = 0)+
    geom_point()+
    geom_label_repel(data = plt_df[plt_df$gene %in% c(emb_g_plt, reg_g_plt, b_g_plt),], 
                     mapping = aes(label = gene), show.legend = F)+
    scale_colour_manual(values = cols)+
    labs(title = ct, x = "logFC(healthy vs embolised)", 
         y = "logFC(healthy vs regenerating)")+
    theme_bw()
  
  plt_list_cond[[ct]] = plt
  df_list_cond[[ct]] = plt_df
  
  write.csv(plt_df, file = paste0("to_send/FC_condVShealthy_", ct, ".csv"), 
            col.names = T, row.names = F, quote = F)
}

terms_list = c("liver regeneration" = "GO:0097421", "regeneration" = "GO:0031099", 
               "tissue regeneration" = "GO:0042246", "liver development" = "GO:0001889",
               "cell migration" = "GO:0016477", "lipid metabolic process" = "GO:0006629", 
               "monocarboxylic acid metabolic process" = "GO:0032787", 
               "xenobiotic metabolic process" = "GO:0006805", "complement activation" = "GO:0006956")
g_path = mget(terms_list,org.Hs.egGO2ALLEGS)
refsymbol = org.Hs.egSYMBOL
g_path = lapply(g_path, function(x) unlist(as.list(refsymbol[x])))
names(g_path) = names(terms_list)

g_path_df = data.frame("gene" = unlist(g_path[-c(1,3)]),
                       "terms" = unlist(lapply(names(g_path[-c(1,3)]), function(x) rep(x, length(g_path[[x]])))),
                       stringsAsFactors = F)
g_path_df = g_path_df[!duplicated(g_path_df$gene),]

plt_df = merge(plt_df, g_path_df, by = "gene", all.x = T)
plt_df$terms[is.na(plt_df$terms)] = "other"
plt_df$terms = factor(plt_df$terms, levels = c(unique(g_path_df$terms), "other"))
plt_df = plt_df[order(plt_df$terms, decreasing = T),]
```






Cell types affected by condtions

```{r}
de_df = filt_comps$major_ct$healthy_v_embolised
de_df = de_df[de_df$p_val_adj<=0.05 & abs(de_df$avg_logFC)>=0.2,]
genes_up = tapply(de_df$gene[de_df$avg_logFC>0], de_df$celltype[de_df$avg_logFC>0], list)
genes_dn = tapply(de_df$gene[de_df$avg_logFC<0], de_df$celltype[de_df$avg_logFC<0], list)
de_hep = filt_comps$major_ct_hep$healthy_v_embolised
genes_up$Hepatocytes = de_hep$gene[de_hep$avg_logFC>=0.2 & de_hep$p_val_adj<=0.05]
genes_dn$Hepatocytes = de_hep$gene[de_hep$avg_logFC<=(-0.2) & de_hep$p_val_adj<=0.05]
de_df_g = rbind(reshape2::melt(genes_up), reshape2::melt(genes_dn))
de_df_g$dedir = c(rep("higher in healthy", length(unlist(genes_up))), 
                  rep("higher in embolised", length(unlist(genes_dn))))

de_df_r = filt_comps$major_ct$healthy_v_regenerating
de_df_r = de_df_r[de_df_r$p_val_adj<=0.05 & abs(de_df_r$avg_logFC)>=0.2,]
genes_up = tapply(de_df_r$gene[de_df_r$avg_logFC>0], de_df_r$celltype[de_df_r$avg_logFC>0], list)
genes_dn = tapply(de_df_r$gene[de_df_r$avg_logFC<0], de_df_r$celltype[de_df_r$avg_logFC<0], list)
de_hep = filt_comps$major_ct_hep$healthy_v_regenerating
genes_up$Hepatocytes = de_hep$gene[de_hep$avg_logFC>=0.2 & de_hep$p_val_adj<=0.05]
genes_dn$Hepatocytes = de_hep$gene[de_hep$avg_logFC<=(-0.2) & de_hep$p_val_adj<=0.05]
de_df_r_g = rbind(reshape2::melt(genes_up), reshape2::melt(genes_dn))
de_df_r_g$dedir = c(rep("higher in healthy", length(unlist(genes_up))), 
                    rep("higher in regenerating", length(unlist(genes_dn))))


counts_de = rbind(data.frame(table(de_df_g$L1, de_df_g$dedir)), data.frame(table(de_df_r_g$L1, de_df_r_g$dedir)))
counts_de$comp = rep(c("healthy vs embolised", "healthy vs regenerating"), each = 10)
counts_de$Var2 = factor(counts_de$Var2, levels = rev(c("higher in healthy", "higher in regenerating", "higher in embolised")))

ggplot(counts_de, aes(x = Var1, y = Freq, fill = Var2))+
  facet_wrap(~comp)+
  scale_fill_manual(values = c("higher in healthy" = "orange", "higher in regenerating" = "salmon", "higher in embolised" = "darkred"))+
  scale_y_continuous(expand = c(0,0), limits = c(0, 4000))+
  labs(fill = "DE genes", x = "Cell types", y = "# DE genes")+
  geom_bar(stat = "identity")+
  theme_classic()+
  theme(legend.position = c(0.12,0.87))



de_df = filt_comps$cell_type_simp$healthy_v_embolised
de_df = de_df[de_df$p_val_adj<=0.05 & abs(de_df$avg_logFC)>=0.2,]
genes_up = tapply(de_df$gene[de_df$avg_logFC>0], de_df$celltype[de_df$avg_logFC>0], list)
genes_dn = tapply(de_df$gene[de_df$avg_logFC<0], de_df$celltype[de_df$avg_logFC<0], list)
de_hep = filt_comps$major_ct_hep$healthy_v_embolised
genes_up$Hepatocytes = de_hep$gene[de_hep$avg_logFC>=0.2 & de_hep$p_val_adj<=0.05]
genes_dn$Hepatocytes = de_hep$gene[de_hep$avg_logFC<=(-0.2) & de_hep$p_val_adj<=0.05]
de_df_g = rbind(reshape2::melt(genes_up), reshape2::melt(genes_dn))
de_df_g$dedir = c(rep("higher in healthy", length(unlist(genes_up))), 
                  rep("higher in embolised", length(unlist(genes_dn))))

de_df_r = filt_comps$cell_type_simp$healthy_v_regenerating
de_df_r = de_df_r[de_df_r$p_val_adj<=0.05 & abs(de_df_r$avg_logFC)>=0.2,]
genes_up = tapply(de_df_r$gene[de_df_r$avg_logFC>0], de_df_r$celltype[de_df_r$avg_logFC>0], list)
genes_dn = tapply(de_df_r$gene[de_df_r$avg_logFC<0], de_df_r$celltype[de_df_r$avg_logFC<0], list)
de_hep = filt_comps$major_ct_hep$healthy_v_regenerating
genes_up$Hepatocytes = de_hep$gene[de_hep$avg_logFC>=0.2 & de_hep$p_val_adj<=0.05]
genes_dn$Hepatocytes = de_hep$gene[de_hep$avg_logFC<=(-0.2) & de_hep$p_val_adj<=0.05]
de_df_r_g = rbind(reshape2::melt(genes_up), reshape2::melt(genes_dn))
de_df_r_g$dedir = c(rep("higher in healthy", length(unlist(genes_up))), 
                    rep("higher in regenerating", length(unlist(genes_dn))))


counts_de = rbind(data.frame(table(de_df_g$L1, de_df_g$dedir)), 
                  data.frame(table(de_df_r_g$L1, de_df_r_g$dedir)))
counts_de$comp = rep(c("healthy vs embolised", "healthy vs regenerating"), each = 14)
counts_de$Var2 = factor(counts_de$Var2, 
                        levels = rev(c("higher in healthy", "higher in regenerating", 
                                       "higher in embolised")))

ggplot(counts_de, aes(x = Var1, y = Freq, fill = Var2))+
  facet_wrap(~comp)+
  scale_fill_manual(values = c("higher in healthy" = "orange", 
                               "higher in regenerating" = "salmon", "higher in embolised" = "darkred"))+
  scale_y_continuous(expand = c(0,0), limits = c(0, 3500))+
  labs(fill = "DE genes", x = "Cell types", y = "# DE genes")+
  geom_bar(stat = "identity")+
  theme_classic()+
  theme(legend.justification=c(0,1), legend.position=c(0,1),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```



```{r}
reg_hep = go_enr_list$major_ct_hep$healthy_v_regenerating$Hepatocytes
reg_hep = reg_hep[rpfilter(reg_hep) & reg_hep$DB=="GO Term",]
reg_genes_terms = apply(reg_hep, 1, function(x) unlist(strsplit(x[8], "/", fixed = T)))
names(reg_genes_terms) = reg_hep$Description
reg_genes_terms = reshape2::melt(reg_genes_terms)
reg_genes_terms = merge(reg_genes_terms, reg_hep[,c(2,11)], by.x = 2, by.y = 1)

emb_hep = go_enr_list$major_ct_hep$healthy_v_embolised$Hepatocytes
emb_hep = emb_hep[rpfilter(emb_hep) & emb_hep$DB=="GO Term",]
emb_genes_terms = apply(emb_hep, 1, function(x) unlist(strsplit(x[8], "/", fixed = T)))
names(emb_genes_terms) = emb_hep$Description
emb_genes_terms = reshape2::melt(emb_genes_terms)
emb_genes_terms = merge(emb_genes_terms, emb_hep[,c(2,11)], by.x = 2, by.y = 1)

emb_genes = setdiff(unique(emb_genes_terms$value[emb_genes_terms$cond=="embolised"]), unique(reg_genes_terms$value[reg_genes_terms$cond=="regenerating"]))
reg_genes = setdiff(unique(reg_genes_terms$value[reg_genes_terms$cond=="regenerating"]), unique(emb_genes_terms$value[emb_genes_terms$cond=="embolised"]))
```





```{r}
#setup matrix with all genes
all_g = rownames(hep_cells_ds)
mat = matrix(0, nrow = length(all_g), ncol = 3)
colnames(mat) = unique(hep_cells_ds@meta.data$Condition)
rownames(mat) = all_g

# downsample counts
set.seed(1)
hep_cells_ds = allcells_css[,allcells_css@meta.data$major_ct=="Hepatocytes" &
                              !is.na(allcells_css@meta.data$major_ct)]

fact = hep_cells_ds@meta.data$nCount_SCT/min(tapply(hep_cells_ds@meta.data$nCount_SCT,
                                                    hep_cells_ds@meta.data$Condition, median))
fact = ifelse(fact<=1, 0, log2(fact))
thinned_counts = seqgendiff::thin_lib(as.matrix(hep_cells_ds@assays$SCT@counts), 
                                      thinlog2 = fact, relative = T)
rownames(thinned_counts$mat) = rownames(hep_cells_ds@assays$SCT@data)
colnames(thinned_counts$mat) = colnames(hep_cells_ds@assays$SCT@data)
hep_cells_ds@assays$SCT@counts = Matrix::Matrix(thinned_counts$mat, sparse = T)
hep_cells_ds@assays$SCT@data = log1p(hep_cells_ds@assays$SCT@counts)

# calculate score
for(n in unique(hep_cells_ds@meta.data$Condition)){
  cells = hep_cells_ds@meta.data$Condition==n
  sub_dat = hep_cells_ds@assays$SCT@data[,cells]
  gene_exp = apply(sub_dat, 1, function(x){
    sx = sum(x>0)
    return(ifelse(sx>1, sx*sum(colSums(sub_dat[,x>0])), 
                  ifelse(sx>0, sx*sum(sub_dat[,x>0]), 0)))
  })
  sc = gene_exp/(sum(cells)*sum(colSums(sub_dat)))
  mat[names(sc),n] = sc
}
mat[is.na(mat)] = 0
maxmat = apply(mat, 2, function(x) x/max(x))
```




