---
title: "R Notebook"
output: html_notebook
---


# General Setup
Setup chunk

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_knit$get("root.dir")
```

Setup reticulate

```{r}
library(reticulate)
knitr::knit_engines$set(python = reticulate::eng_python)
py_available(initialize = FALSE)
use_python(Sys.which("python"))
py_config()
```

Load libraries

```{r}
library(Seurat)
library(ggplot2)
library(ggridges)
```



# Load data
Load gene lists

```{r}
rpl_genes = read.csv("../../gene_refs/human/RPL_genes.csv", stringsAsFactors = F)
rps_genes = read.csv("../../gene_refs/human/RPS_genes.csv", stringsAsFactors = F)
```

Load all prepared matrices, create lists of Seurat objects  
For each matrix, we'll be running scrublet before deleting it  
Start scrublet

```{r}
scr = import("scrublet", as = "scr")
runScrublet = function(mat, ncomp = 30L){
  scrub = scr$Scrublet(Matrix::t(mat), n_neighbors=round(sqrt(ncol(mat))))
  doub_df = data.frame(row.names = colnames(mat), 
                       scrub$scrub_doublets(verbose=F,n_prin_comps=ncomp))
  colnames(doub_df) = c("scrublet_score", "scrublet_isDoublet")
  return(doub_df)
}
```

Healthy

```{r}
load("data/Healthy/1d_prepared.rdata")
load("data/Healthy/2d_prepared.rdata")
load("data/Healthy/3d_prepared.rdata")
healthy_srat = list()
for(obj in ls()[grepl(".data", ls())]){
  doub_df = tryCatch({runScrublet(get(obj))}, 
                     error = function(e){runScrublet(get(obj), ncomp = 20L)}, 
                    finally = {print(obj)})
  healthy_srat[[obj]] = CreateSeuratObject(get(obj), project = obj, meta.data = doub_df)
  rm(list = obj)
}
```

Conditions

```{r}
load("data/Conditions/Donor_1/prepared_liv_reg1.rdata")
load("data/Conditions/Donor_1/prepared_seurat_Reg1Hep_N_R_reseq.rdata")
load("data/Conditions/Donor_2/prepared_liv_hepRandN_2d.rdata")
load("data/Conditions/Donor_2/prepared_liv_NPC_RandN_2d.rdata")
load("data/Conditions/Donor_3/prepared_liv_hepE_3d.rdata")
load("data/Conditions/Donor_3/prepared_liv_hepR_3d.rdata")
load("data/Conditions/Donor_3/prepared_liv_NPC_E_3d.rdata")
load("data/Conditions/Donor_3/prepared_liv_NPC_r_3d.rdata")
load("data/Conditions/Donor_5/Data.Reg5_HepE.rdata")
load("data/Conditions/Donor_5/Data.Reg5_HepR.rdata")
load("data/Conditions/Donor_5/Data.Reg5_npcE.rdata")
load("data/Conditions/Donor_5/Data.Reg5_npcR.rdata")
load("data/Conditions/Donor_6/Data.Reg6_HepE.rdata")
load("data/Conditions/Donor_6/Data.Reg6_HepR.rdata")
load("data/Conditions/Donor_6/Data.Reg6_npcE.rdata")
load("data/Conditions/Donor_6/Data.Reg6_npcR.rdata")
load("data/Conditions/Donor_7/Data.Reg7_HepE.rdata")
load("data/Conditions/Donor_7/Data.Reg7_HepR.rdata")
load("data/Conditions/Donor_7/Data.Reg7_npcE.rdata")
load("data/Conditions/Donor_7/Data.Reg7_npcR.rdata")
cond_srat = list()
for(obj in ls()[grepl("ata", ls())]){
  doub_df = tryCatch({runScrublet(get(obj))}, 
                     error = function(e){runScrublet(get(obj), ncomp = 20L)}, 
                    finally = {print(obj)})
  cond_srat[[obj]] = CreateSeuratObject(get(obj), project = obj, meta.data = doub_df)
  rm(list = obj)
}
```

Frozen (nuclei)

```{r}
frozen_srat = list()
for(obj in c("H1", "H3", "H4")){
  load(paste0("data/Frozen_sn_Healthy/sn_", obj, ".rdata"))
  doub_df = tryCatch({runScrublet(mat)}, 
                     error = function(e){runScrublet(mat, ncomp = 20L)}, 
                    finally = {print(obj)})
  frozen_srat[[obj]] = CreateSeuratObject(mat, project = paste0("sn_", obj), 
                                          meta.data = doub_df)
  rm(mat)
}
gc()
```

Adding metadata and running a "quick" UMAP to visualise scrublet scores

```{r}
basicSeurat = function(obj, npcs = 10){
  # scores
  obj = PercentageFeatureSet(obj, pattern = "^MT-", col.name = "percent.mt", assay = "RNA")
  obj = PercentageFeatureSet(obj, col.name = "percent.rb", assay = "RNA", 
                             features = c(rpl_genes$Approved.symbol,
                                          rps_genes$Approved.symbol))
  
  # cell cycle scoring
  # might have trouble with bins so running iteratively until it works
  bins = 24
  while(T){
    objwsc = tryCatch({CellCycleScoring(obj, s.features = cc.genes$s.genes, nbin = bins, 
                                        assay = "RNA", g2m.features = cc.genes$g2m.genes,
                                        set.ident = F)},
                   error = function(e){print(bins)})
    
    if(is.numeric(objwsc)){
      bins = bins-2
    } else{
      obj = objwsc
      break
    }
  }
  
  # normalise and PCA-UMAP
  obj = suppressWarnings(SCTransform(obj, do.correct.umi = T, vars.to.regress = "nCount_RNA",
                                     verbose = F, variable.features.rv.th = 1,
                                     variable.features.n = NULL, seed.use = 1))
  obj = RunPCA(obj, verbose = F)
  obj = RunUMAP(obj, dims = 1:npcs, verbose = F)
  
  return(obj)
}

for(obj in names(healthy_srat)){healthy_srat[[obj]] = basicSeurat(healthy_srat[[obj]])}
for(obj in names(cond_srat)){cond_srat[[obj]] = basicSeurat(cond_srat[[obj]])}
for(obj in names(frozen_srat)){frozen_srat[[obj]] = basicSeurat(frozen_srat[[obj]])}
```

Save Seurat objects

```{r}
save(healthy_srat, file = "data/processed/healthy_srat_beforeDoublet.RData")
save(cond_srat, file = "data/processed/cond_srat_beforeDoublet.RData")
save(frozen_srat, file = "data/processed/frozen_srat_beforeDoublet.RData")
```

Define matches between matrices and metadata

```{r}
dat_matches = c("EKS.data", "sc_H1", "HD1", "EKS","cells","healthy", "Exp1","No",
                "NPC.data", "sc_H1", "HD1", "NPC", "cells", "healthy", "Exp1","No",
                "HEP.data", "sc_H1", "HD1", "HEP", "cells", "healthy", "Exp1","No",
                "Hep.2d.data", "sc_H2", "HD2", "HEP", "cells", "healthy", "Exp1","No",
                "NPC.2d.data", "sc_H2", "HD2", "NPC", "cells", "healthy", "Exp1","No",
                "Hep.3d.data", "sc_H3", "HD3", "HEP", "cells", "healthy", "Exp1","No",
                "NPC.3d.data", "sc_H3", "HD3", "NPC", "cells", "healthy", "Exp1","No",
                
                "data.Reg1.NHep", "sc_E1", "DD1", "HEP", "cells", "embolised", "Exp1","No",
                "data.Reg1.RHep","sc_R1", "DD1", "HEP", "cells", "regenerating", "Exp1","No",
                "Hep1N.data", "sc_E1", "DD1", "HEP", "cells", "embolised", "Exp1","No",
                "Hep1R.data", "sc_R1", "DD1", "HEP", "cells", "regenerating", "Exp1","No",
                "Hep2N.data", "sc_E1", "DD1", "HEP", "cells", "embolised", "Exp2","No",
                "Hep2R.data", "sc_R1", "DD1", "HEP", "cells", "regenerating", "Exp2","No",
                "HepRn.2d.data", "sc_E2", "DD2", "HEP", "cells", "embolised", "Exp1","No",
                "HepRr.2d.data", "sc_R2", "DD2", "HEP", "cells", "regenerating", "Exp1","No",
                "NPC.Rn.2d.data", "sc_E2", "DD2", "NPC", "cells", "embolised", "Exp1","No",
                "NPC.Rr.2d.data","sc_R2", "DD2", "NPC", "cells", "regenerating", "Exp1","No",
                "NPCn.data", "sc_E1", "DD1", "NPC", "cells", "embolised", "Exp1","No",
                "NPCr.data", "sc_R1", "DD1", "NPC", "cells", "regenerating", "Exp1","No",
                "HepRe.3d.data", "sc_E3", "DD3", "HEP", "cells", "embolised", "Exp1","No",
                "HepRr.3d.data", "sc_R3", "DD3", "HEP", "cells", "regenerating", "Exp1","No",
                "NPC.Re.3d.data", "sc_E3", "DD3", "NPC", "cells", "embolised", "Exp1","No",
                "NPC.Rr.3d.data","sc_R3", "DD3", "NPC", "cells", "regenerating", "Exp1","No",
                "Data.Reg5_HepE", "sc_E5", "DD5", "HEP", "cells", "embolised", "Exp1","No",
                "Data.Reg5_HepR","sc_R5", "DD5", "HEP", "cells", "regenerating", "Exp1","No",
                "Data.Reg5_npcE", "sc_E5", "DD5", "NPC", "cells", "embolised", "Exp1","No",
                "Data.Reg5_npcR","sc_R5", "DD5", "NPC", "cells", "regenerating", "Exp1","No",
                "Data.Reg6_HepE", "sc_E6", "DD6", "HEP", "cells", "embolised", "Exp1","No",
                "Data.Reg6_HepR","sc_R6", "DD6", "HEP", "cells", "regenerating", "Exp1","No",
                "Data.Reg6_npcE","sc_E6", "DD6", "NPC", "cells", "embolised", "Exp1","No",
                "Data.Reg6_npcR","sc_R6", "DD6", "NPC", "cells", "regenerating", "Exp1","No",
                "Data.Reg7_HepE", "sc_E7", "DD7", "HEP", "cells", "embolised", "Exp1","No",
                "Data.Reg7_HepR","sc_R7", "DD7", "HEP", "cells", "regenerating", "Exp1","No",
                "Data.Reg7_npcE","sc_E7","DD7","NPC","cells","embolised","Exp1","No",
                "Data.Reg7_npcR","sc_R7", "DD7", "NPC", "cells", "regenerating", "Exp1","No",
                
                "sn_H1", "sn_H1", "HD1", "all", "nuclei", "healthy", "Exp1","No",
                "sn_H3", "sn_H1", "HD3", "all", "nuclei", "healthy", "Exp1","No",
                "sn_H4", "sn_H1", "HD4", "all", "nuclei", "healthy", "Exp1","No")

dat_matches = matrix(dat_matches, ncol = 8, byrow = T)
colnames(dat_matches) = c("mat_id", "Name", "Donor", "Fraction", "Isolation", 
                          "Condition", "Experiment", "Reseq")
dat_matches = data.frame(dat_matches, stringsAsFactors = F)
write.csv(dat_matches, "data/processed/matrix_metadata_matches.csv", 
          col.names = T, row.names = T, quote = F)
```

Fix/simplify metadata

```{r}
col_dat_matches = c("Name", "Donor", "Fraction", "Isolation",
                    "Condition", "Experiment", "Reseq")
for(n in names(healthy_srat)){
  dat = merge(healthy_srat[[n]]@meta.data, dat_matches, by = 1)[,col_dat_matches]
  rownames(dat) = rownames(healthy_srat[[n]]@meta.data)
  healthy_srat[[n]] = AddMetaData(healthy_srat[[n]], dat)
}
for(n in names(cond_srat)){
  dat = merge(cond_srat[[n]]@meta.data, dat_matches, by = 1)[,col_dat_matches]
  rownames(dat) = rownames(cond_srat[[n]]@meta.data)
  cond_srat[[n]] = AddMetaData(cond_srat[[n]], dat)
}
for(n in names(frozen_srat)){
  dat = merge(frozen_srat[[n]]@meta.data, dat_matches, by = 1)[,col_dat_matches]
  rownames(dat) = rownames(frozen_srat[[n]]@meta.data)
  frozen_srat[[n]] = AddMetaData(frozen_srat[[n]], dat)
}
```

Plot doublet score distributions

```{r}
plotDoubScore = function(l){
  plot_df = data.frame()
  thr = c()
  for(n in names(l)){
    plot_df = rbind(plot_df, l[[n]]@meta.data[,c("scrublet_score","orig.ident",
                                                 "Fraction","Name")])
    thr = c(thr, tapply(l[[n]]@meta.data$scrublet_score,
                        l[[n]]@meta.data$scrublet_isDoublet,
                        max)["FALSE"])
    
  }
  line_df = data.frame("orig.ident" = names(l),
                       "threshold" = thr)
  plt = ggplot(plot_df, aes(x = scrublet_score, y = orig.ident, 
                            group = orig.ident))+
    ggridges::geom_density_ridges(size = 0.25, mapping = aes(fill = Fraction))+
    #geom_vline(data = line_df, mapping = aes(xintercept = threshold, y = orig.ident))+
    geom_segment(data = line_df, aes(x = threshold, xend = threshold, 
                                     y = as.numeric(orig.ident),
                                     yend = as.numeric(orig.ident) + 1))+
    scale_x_continuous(expand = c(0,0), limits = c(-.01,1.01))+
    theme_bw()

  return(plt)
}

pdf("results/initial_stats/doublet_Score_distributions.pdf", 
    height = 7, width = 8, useDingbats = F)
plotDoubScore(healthy_srat)
plotDoubScore(cond_srat)
plotDoubScore(frozen_srat)
dev.off()
```

Plot doublet score UMAP

```{r}
plotUMAPFeat = function(l ,f){
  return(lapply(names(l), 
                function(x) FeaturePlot(object = l[[x]], reduction = "umap",
                                        features = f)+
                  labs(title = x)+
                  theme(aspect.ratio = 1)))
}

pdf("results/initial_stats/healthy_doublet_Score_UMAP.pdf", 
    height = 10, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(healthy_srat, "scrublet_score") 
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_doublet_Score_UMAP.pdf", 
    height = 16, width = 13, useDingbats = F)
plt_list = plotUMAPFeat(cond_srat, "scrublet_score") 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_doublet_Score_UMAP.pdf", 
    height = 6, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(frozen_srat, "scrublet_score") 
cowplot::plot_grid(plotlist = plt_list, nrow = 2, ncol = 2)
dev.off()
```

Plot cell cycle classification UMAP

```{r}
plotUMAPDim = function(l ,f){
  return(lapply(names(l), 
                function(x) DimPlot(object = l[[x]], reduction = "umap",
                                        group.by = f)+
                  labs(title = x)+
                  theme(aspect.ratio = 1)))
}

pdf("results/initial_stats/healthy_ccPhase_UMAP.pdf", 
    height = 10, width = 6, useDingbats = F)
plt_list = plotUMAPDim(healthy_srat, "Phase") 
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_ccPhase_UMAP.pdf", 
    height = 16, width = 13, useDingbats = F)
plt_list = plotUMAPDim(cond_srat, "Phase") 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_ccPhase_UMAP.pdf", 
    height = 6, width = 6, useDingbats = F)
plt_list = plotUMAPDim(frozen_srat, "Phase") 
cowplot::plot_grid(plotlist = plt_list, nrow = 2, ncol = 2)
dev.off()
```

Plot MT% UMAP

```{r}
pdf("results/initial_stats/healthy_percmt_UMAP.pdf", 
    height = 10, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(healthy_srat, "percent.mt") 
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_percmt_UMAP.pdf", 
    height = 16, width = 13, useDingbats = F)
plt_list = plotUMAPFeat(cond_srat, "percent.mt") 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_percmt_UMAP.pdf", 
    height = 6, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(frozen_srat, "percent.mt") 
cowplot::plot_grid(plotlist = plt_list, nrow = 2, ncol = 2)
dev.off()
```

Plot MT% violin

```{r}
plotViolinFeat = function(l ,f){
  return(lapply(names(l), 
                function(x) VlnPlot(object = l[[x]], features = f, pt.size = 0.15)+
                  labs(title = x)+
                  theme(legend.position = "none",
                        axis.text.x = element_blank())))
}

pdf("results/initial_stats/healthy_percmt_vio.pdf", 
    height = 10, width = 4, useDingbats = F)
plt_list = plotViolinFeat(healthy_srat, "percent.mt") 
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_percmt_vio.pdf", 
    height = 24, width = 9, useDingbats = F)
plt_list = plotViolinFeat(cond_srat, "percent.mt") 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_percmt_vio.pdf", 
    height = 3, width = 6, useDingbats = F)
plt_list = plotViolinFeat(frozen_srat, "percent.mt") 
cowplot::plot_grid(plotlist = plt_list, ncol = 3)
dev.off()
```

Plot ribosomal% UMAP

```{r}
pdf("results/initial_stats/healthy_percrb_UMAP.pdf", 
    height = 10, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(healthy_srat, "percent.rb") 
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_percrb_UMAP.pdf", 
    height = 16, width = 13, useDingbats = F)
plt_list = plotUMAPFeat(cond_srat, "percent.rb") 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_percrb_UMAP.pdf", 
    height = 6, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(frozen_srat, "percent.rb") 
cowplot::plot_grid(plotlist = plt_list, nrow = 2, ncol = 2)
dev.off()
```

Plot ribosomal% violin

```{r}
pdf("results/initial_stats/healthy_percrb_vio.pdf", 
    height = 10, width = 4, useDingbats = F)
plt_list = plotViolinFeat(healthy_srat, "percent.rb") 
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_percrb_vio.pdf", 
    height = 24, width = 9, useDingbats = F)
plt_list = plotViolinFeat(cond_srat, "percent.rb") 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_percrb_vio.pdf", 
    height = 3, width = 6, useDingbats = F)
plt_list = plotViolinFeat(frozen_srat, "percent.rb") 
cowplot::plot_grid(plotlist = plt_list, ncol = 3)
dev.off()
```

Plot total counts UMAP

```{r}
pdf("results/initial_stats/healthy_counts_UMAP.pdf", 
    height = 10, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(healthy_srat, "nCount_RNA") 
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_counts_UMAP.pdf", 
    height = 16, width = 13, useDingbats = F)
plt_list = plotUMAPFeat(cond_srat, "nCount_RNA") 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_counts_UMAP.pdf", 
    height = 6, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(frozen_srat, "nCount_RNA") 
cowplot::plot_grid(plotlist = plt_list, nrow = 2, ncol = 2)
dev.off()
```

Number of genes vs number of UMI

```{r}
plotCountsVsGenes = function(l){
  plt_l = list()
  for(x in names(l)){
      plot_df = data.frame("nUMI" = l[[x]]$nCount_RNA,
                           "nGene" = l[[x]]$nFeature_RNA)
      
      plt_l[[x]] = ggplot(plot_df, aes(x = nGene, y = nUMI))+
        geom_hex(bins = 65)+
        scale_y_log10()+
        labs(title = x)+
        theme_bw()+
        theme(aspect.ratio = 1)
  }
  
  return(plt_l)
}

pdf("results/initial_stats/healthy_counts_genes.pdf", 
    height = 10, width = 6, useDingbats = F)
plt_list = plotCountsVsGenes(healthy_srat) 
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_counts_genes.pdf", 
    height = 16, width = 13, useDingbats = F)
plt_list = plotCountsVsGenes(cond_srat) 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_counts_genes.pdf", 
    height = 6, width = 6, useDingbats = F)
plt_list = plotCountsVsGenes(frozen_srat) 
cowplot::plot_grid(plotlist = plt_list, nrow = 2, ncol = 2)
dev.off()
```

Filter cells  
We'll be doing a very fine clustering to identify groups of cells with excessive doublet scores/MT%/high or low counts

```{r}
getClusters = function(obj){
  obj <- FindNeighbors(obj, dims = 1:20, verbose = FALSE)
  obj <- FindClusters(obj, algorithm = 2, verbose = F, resolution = 10)
  return(obj)
}

healthy_srat = lapply(healthy_srat, getClusters)
cond_srat = lapply(cond_srat, getClusters)
frozen_srat = lapply(frozen_srat, getClusters)
#DimPlot(xxx, reduction = "umap", group.by = "SCT_snn_res.10")
```

Now we calculate those mean stats per cluster and define the thresholds

```{r}
calcMeanClust = function(l, v){
  a = lapply(l, function(x) tapply(x@meta.data[,v], x$SCT_snn_res.10, mean))
  res = reshape2::melt(a)
  a = lapply(l, function(x) table(x$SCT_snn_res.10))
  res2 = reshape2::melt(a)
  res = cbind(res, res2)[,-c(4,6)]
  colnames(res) = c("cluster", "mean_value", "sample", "n_cells")
  return(res)
}

var_list = list()
for(v in c("scrublet_score", "percent.mt", "percent.rb", "nCount_RNA")){
  var_list[[v]] = list("healthy" = calcMeanClust(healthy_srat, v),
                       "cond" = calcMeanClust(cond_srat, v),
                       "frozen" = calcMeanClust(frozen_srat, v))
}

plotMeanClust = function(var_l, l){
  plt_list = list()
  for(n in names(var_l)){
    plt_list[[n]] = ggplot(var_l[[n]][[l]], aes(x = n_cells, y = mean_value, 
                                                colour = sample))+
      geom_point()+
      theme_bw()+
      labs(title = n)+
      theme(aspect.ratio = 1,
            legend.position = "bottom",
            legend.key.size = unit(0.5, "cm"))
    
    if(n=="nCount_RNA") plt_list[[n]] = plt_list[[n]] + scale_y_log10()
    
    leg = cowplot::get_legend(plt_list[[n]])
    plt_list[[n]] = plt_list[[n]]+theme(legend.position = "none")
  }
  
  plt_list$legend = leg
  
  all_plt = cowplot::plot_grid(cowplot::plot_grid(plotlist = plt_list[1:4], 
                                                  ncol = 2, nrow = 2),
                               plt_list[[5]], nrow = 2, rel_heights = c(0.85, 0.2))
  return(all_plt)
}

pdf("results/initial_stats/mean_qc_microclusters.pdf", 
    height = 7, width = 6, useDingbats = F)
plotMeanClust(var_list, "healthy")
plotMeanClust(var_list, "cond")
plotMeanClust(var_list, "frozen")
dev.off()
```

Remove cells (doublets, low counts)

```{r}

```

Re-run scTransform, PCA and UMAP on filtered data

```{r}

```

Save objects

```{r}

```

