---
title: "Individual Processing"
output: html_notebook
---



# General Setup
Setup chunk

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_knit$get("root.dir")
```

Setup reticulate

```{r}
library(reticulate)
knitr::knit_engines$set(python = reticulate::eng_python)
py_available(initialize = FALSE)
use_python(Sys.which("python"))
py_config()
```

Load libraries

```{r}
library(Seurat)
library(ggplot2)
library(ggridges)
library(mixtools)
library(dplyr)
library(SoupX)
```



# Load data
Load gene lists

```{r}
rpl_genes = read.csv("../../gene_refs/human/RPL_genes.csv", stringsAsFactors = F)
rps_genes = read.csv("../../gene_refs/human/RPS_genes.csv", stringsAsFactors = F)
```

Load all prepared matrices, create lists of Seurat objects  
For each matrix, we'll be running scrublet before deleting it  
Start scrublet

```{r}
scr = import("scrublet", as = "scr")
runScrublet = function(mat, ncomp = 30L){
  scrub = scr$Scrublet(Matrix::t(mat), n_neighbors=round(sqrt(ncol(mat))))
  doub_df = data.frame(row.names = colnames(mat), 
                       scrub$scrub_doublets(verbose=F,n_prin_comps=ncomp))
  colnames(doub_df) = c("scrublet_score", "scrublet_isDoublet")
  return(doub_df)
}
```

Healthy

```{r}
load("data/Healthy/1d_prepared.rdata")
load("data/Healthy/2d_prepared.rdata")
load("data/Healthy/3d_prepared.rdata")
healthy_srat = list()
for(obj in ls()[grepl(".data", ls())]){
  doub_df = tryCatch({runScrublet(get(obj))}, 
                     error = function(e){runScrublet(get(obj), ncomp = 20L)}, 
                    finally = {print(obj)})
  healthy_srat[[obj]] = CreateSeuratObject(get(obj), project = obj, meta.data = doub_df)
  rm(list = obj)
}
```

Conditions

```{r}
load("data/Conditions/Donor_1/prepared_liv_reg1.rdata")
rm(Hep2N.data, Hep2R.data)
load("data/Conditions/Donor_1/sc_E1_mergedHep2.rdata")
load("data/Conditions/Donor_1/sc_R1_mergedHep2.rdata")
load("data/Conditions/Donor_2/prepared_liv_hepRandN_2d.rdata")
load("data/Conditions/Donor_2/prepared_liv_NPC_RandN_2d.rdata")
load("data/Conditions/Donor_3/prepared_liv_hepE_3d.rdata")
load("data/Conditions/Donor_3/prepared_liv_hepR_3d.rdata")
load("data/Conditions/Donor_3/prepared_liv_NPC_E_3d.rdata")
load("data/Conditions/Donor_3/prepared_liv_NPC_r_3d.rdata")
load("data/Conditions/Donor_5/Data.Reg5_HepE.rdata")
load("data/Conditions/Donor_5/Data.Reg5_HepR.rdata")
load("data/Conditions/Donor_5/Data.Reg5_npcE.rdata")
load("data/Conditions/Donor_5/Data.Reg5_npcR.rdata")
load("data/Conditions/Donor_6/Data.Reg6_HepE.rdata")
load("data/Conditions/Donor_6/Data.Reg6_HepR.rdata")
load("data/Conditions/Donor_6/Data.Reg6_npcE.rdata")
load("data/Conditions/Donor_6/Data.Reg6_npcR.rdata")
load("data/Conditions/Donor_7/Data.Reg7_HepE.rdata")
load("data/Conditions/Donor_7/Data.Reg7_HepR.rdata")
load("data/Conditions/Donor_7/Data.Reg7_npcE.rdata")
load("data/Conditions/Donor_7/Data.Reg7_npcR.rdata")
cond_srat = list()
for(obj in ls()[grepl("ata", ls())]){
  doub_df = tryCatch({runScrublet(get(obj))}, 
                     error = function(e){runScrublet(get(obj), ncomp = 20L)}, 
                    finally = {print(obj)})
  cond_srat[[obj]] = CreateSeuratObject(get(obj), project = obj, meta.data = doub_df)
  rm(list = obj)
}
```

Frozen (nuclei)

```{r}
frozen_srat = list()
for(obj in c("H1", "H3", "H4")){
  load(paste0("data/Frozen_sn_Healthy/sn_", obj, ".rdata"))
  doub_df = tryCatch({runScrublet(mat)}, 
                     error = function(e){runScrublet(mat, ncomp = 20L)}, 
                    finally = {print(obj)})
  frozen_srat[[obj]] = CreateSeuratObject(mat, project = paste0("sn_", obj), 
                                          meta.data = doub_df)
  rm(mat)
}
gc()
```

Adding metadata and running a "quick" UMAP to visualise scrublet scores

```{r}
basicSeurat = function(obj, npcs = 10){
  # scores
  obj = PercentageFeatureSet(obj, pattern = "^MT-", col.name = "percent.mt", assay = "RNA")
  obj = PercentageFeatureSet(obj, col.name = "percent.rb", assay = "RNA", 
                             features = c(rpl_genes$Approved.symbol,
                                          rps_genes$Approved.symbol))
  
  # cell cycle scoring
  # might have trouble with bins so running iteratively until it works
  bins = 24
  while(T){
    objwsc = tryCatch({CellCycleScoring(obj, s.features = cc.genes$s.genes, nbin = bins, 
                                        assay = "RNA", g2m.features = cc.genes$g2m.genes,
                                        set.ident = F)},
                   error = function(e){print(bins)})
    
    if(is.numeric(objwsc)){
      bins = bins-2
    } else{
      obj = objwsc
      break
    }
  }
  
  # normalise and PCA-UMAP
  obj = suppressWarnings(SCTransform(obj, do.correct.umi = T, vars.to.regress = "nCount_RNA",
                                     verbose = F, variable.features.rv.th = 1,
                                     variable.features.n = NULL, seed.use = 1))
  obj = RunPCA(obj, verbose = F)
  obj = RunUMAP(obj, dims = 1:npcs, verbose = F)
  
  return(obj)
}

for(obj in names(healthy_srat)){healthy_srat[[obj]] = basicSeurat(healthy_srat[[obj]])}
for(obj in names(cond_srat)){cond_srat[[obj]] = basicSeurat(cond_srat[[obj]])}
for(obj in names(frozen_srat)){frozen_srat[[obj]] = basicSeurat(frozen_srat[[obj]])}
```

Save Seurat objects

```{r}
save(healthy_srat, file = "data/processed/healthy_srat_beforeDoublet.RData")
save(cond_srat, file = "data/processed/cond_srat_beforeDoublet.RData")
save(frozen_srat, file = "data/processed/frozen_srat_beforeDoublet.RData")
```



# Filtering
## Deciding on filtering thresholds
Define matches between matrices and metadata

```{r}
dat_matches = c("EKS.data", "sc_H1", "HD1", "EKS","cells","healthy", "Exp1","V2",
                "NPC.data", "sc_H1", "HD1", "NPC", "cells", "healthy", "Exp1","V2",
                "HEP.data", "sc_H1", "HD1", "HEP", "cells", "healthy", "Exp1","V2",
                "Hep.2d.data", "sc_H2", "HD2", "HEP", "cells", "healthy", "Exp1","V2",
                "NPC.2d.data", "sc_H2", "HD2", "NPC", "cells", "healthy", "Exp1","V2",
                "Hep.3d.data", "sc_H3", "HD3", "HEP", "cells", "healthy", "Exp1","V2",
                "NPC.3d.data", "sc_H3", "HD3", "NPC", "cells", "healthy", "Exp1","V2",
                
                "Hep1N.data", "sc_E1", "DD1", "HEP", "cells", "embolised", "Exp1","V2",
                "Hep1R.data", "sc_R1", "DD1", "HEP", "cells", "regenerating", "Exp1","V2",
                "sc_E1_mergedHep2.data", "sc_E1", "DD1", "HEP", "cells", "embolised", "Exp2","V2",
                "sc_R1_mergedHep2.data", "sc_R1", "DD1", "HEP", "cells", "regenerating", "Exp2","V2",
                "NPCn.data", "sc_E1", "DD1", "NPC", "cells", "embolised", "Exp1","V2",
                "NPCr.data", "sc_R1", "DD1", "NPC", "cells", "regenerating", "Exp1","V2",
                "HepRn.2d.data", "sc_E2", "DD2", "HEP", "cells", "embolised", "Exp1","V2",
                "HepRr.2d.data", "sc_R2", "DD2", "HEP", "cells", "regenerating", "Exp1","V2",
                "NPC.Rn.2d.data", "sc_E2", "DD2", "NPC", "cells", "embolised", "Exp1","V2",
                "NPC.Rr.2d.data","sc_R2", "DD2", "NPC", "cells", "regenerating", "Exp1","V2",
                "HepRe.3d.data", "sc_E3", "DD3", "HEP", "cells", "embolised", "Exp1","V2",
                "HepRr.3d.data", "sc_R3", "DD3", "HEP", "cells", "regenerating", "Exp1","V2",
                "NPC.Re.3d.data", "sc_E3", "DD3", "NPC", "cells", "embolised", "Exp1","V2",
                "NPC.Rr.3d.data","sc_R3", "DD3", "NPC", "cells", "regenerating", "Exp1","V2",
                "Data.Reg5_HepE", "sc_E5", "DD5", "HEP", "cells", "embolised", "Exp1","V3",
                "Data.Reg5_HepR","sc_R5", "DD5", "HEP", "cells", "regenerating", "Exp1","V3",
                "Data.Reg5_npcE", "sc_E5", "DD5", "NPC", "cells", "embolised", "Exp1","V3",
                "Data.Reg5_npcR","sc_R5", "DD5", "NPC", "cells", "regenerating", "Exp1","V3",
                "Data.Reg6_HepE", "sc_E6", "DD6", "HEP", "cells", "embolised", "Exp1","V3",
                "Data.Reg6_HepR","sc_R6", "DD6", "HEP", "cells", "regenerating", "Exp1","V3",
                "Data.Reg6_npcE","sc_E6", "DD6", "NPC", "cells", "embolised", "Exp1","V3",
                "Data.Reg6_npcR","sc_R6", "DD6", "NPC", "cells", "regenerating", "Exp1","V3",
                "Data.Reg7_HepE", "sc_E7", "DD7", "HEP", "cells", "embolised", "Exp1","V3",
                "Data.Reg7_HepR","sc_R7", "DD7", "HEP", "cells", "regenerating", "Exp1","V3",
                "Data.Reg7_npcE","sc_E7","DD7","NPC","cells","embolised","Exp1","V3",
                "Data.Reg7_npcR","sc_R7", "DD7", "NPC", "cells", "regenerating", "Exp1","V3",
                
                "sn_H1", "sn_H1", "HD1", "all", "nuclei", "healthy", "Exp1","V3",
                "sn_H3", "sn_H3", "HD3", "all", "nuclei", "healthy", "Exp1","V3",
                "sn_H4", "sn_H4", "HD4", "all", "nuclei", "healthy", "Exp1","V3")

dat_matches = matrix(dat_matches, ncol = 8, byrow = T)
colnames(dat_matches) = c("mat_id", "Name", "Donor", "Fraction", "Isolation", 
                          "Condition", "Experiment","ProtocolVersion")
dat_matches = data.frame(dat_matches, stringsAsFactors = F)
write.csv(dat_matches, "data/processed/matrix_metadata_matches.csv", 
          col.names = T, row.names = T, quote = F)
```

Fix/simplify metadata

```{r}
col_dat_matches = c("Name", "Donor", "Fraction", "Isolation",
                    "Condition", "Experiment", "ProtocolVersion")
for(n in names(healthy_srat)){
  dat = merge(healthy_srat[[n]]@meta.data, dat_matches, by = 1)[,col_dat_matches]
  rownames(dat) = rownames(healthy_srat[[n]]@meta.data)
  healthy_srat[[n]] = AddMetaData(healthy_srat[[n]], dat)
}
for(n in names(cond_srat)){
  dat = merge(cond_srat[[n]]@meta.data, dat_matches, by = 1)[,col_dat_matches]
  rownames(dat) = rownames(cond_srat[[n]]@meta.data)
  cond_srat[[n]] = AddMetaData(cond_srat[[n]], dat)
}
for(n in names(frozen_srat)){
  dat = merge(frozen_srat[[n]]@meta.data, dat_matches, by = 1)[,col_dat_matches]
  rownames(dat) = rownames(frozen_srat[[n]]@meta.data)
  frozen_srat[[n]] = AddMetaData(frozen_srat[[n]], dat)
}
```

Simplify sample names

```{r}
for(n in names(healthy_srat)){
  newname=paste(healthy_srat[[n]]@meta.data$Name[1],healthy_srat[[n]]@meta.data$Fraction[1], 
                  healthy_srat[[n]]@meta.data$Experiment[1], sep = ".")
  names(healthy_srat)[names(healthy_srat)==n] = newname
  n_df = data.frame(row.names = colnames(healthy_srat[[n]]),
                    "unique_name" = rep(newname, ncol(healthy_srat[[n]])))
  healthy_srat[[n]] = AddMetaData(healthy_srat[[n]], n_df)
}
for(n in names(cond_srat)){
  newname = paste(cond_srat[[n]]@meta.data$Name[1], cond_srat[[n]]@meta.data$Fraction[1], 
                  cond_srat[[n]]@meta.data$Experiment[1], sep = ".")
  names(cond_srat)[names(cond_srat)==n] = newname
  n_df = data.frame(row.names = colnames(cond_srat[[n]]),
                    "unique_name" = rep(newname, ncol(cond_srat[[n]])))
  cond_srat[[n]] = AddMetaData(cond_srat[[n]], n_df)
}
for(n in names(frozen_srat)){
  newname = paste(frozen_srat[[n]]@meta.data$Name[1],frozen_srat[[n]]@meta.data$Fraction[1], 
                  frozen_srat[[n]]@meta.data$Experiment[1], sep = ".")
  names(frozen_srat)[names(frozen_srat)==n] = newname
  n_df = data.frame(row.names = colnames(frozen_srat[[n]]),
                    "unique_name" = rep(newname, ncol(frozen_srat[[n]])))
  frozen_srat[[n]] = AddMetaData(frozen_srat[[n]], n_df)
}
```

Plot doublet score distributions

```{r}
plotDoubScore = function(l){
  plot_df = data.frame()
  thr = c()
  for(n in names(l)){
    plot_df = rbind(plot_df, l[[n]]@meta.data[,c("scrublet_score","unique_name",
                                                 "Fraction","Name")])
    thr = c(thr, tapply(l[[n]]@meta.data$scrublet_score,
                        l[[n]]@meta.data$scrublet_isDoublet,
                        min)["TRUE"])
    
  }
  line_df = data.frame("unique_name" = names(l),
                       "threshold" = thr)
  plt = ggplot(plot_df, aes(x = scrublet_score, y = unique_name, 
                            group = unique_name))+
    ggridges::geom_density_ridges(size = 0.25, mapping = aes(fill = Fraction))+
    #geom_vline(data = line_df, mapping = aes(xintercept = threshold, y = orig.ident))+
    geom_segment(data = line_df, aes(x = threshold, xend = threshold, 
                                     y = as.numeric(unique_name),
                                     yend = as.numeric(unique_name) + 1))+
    scale_x_continuous(expand = c(0,0), limits = c(-.01,1.01))+
    theme_bw()

  return(plt)
}

pdf("results/initial_stats/doublet_Score_distributions.pdf", 
    height = 7, width = 8, useDingbats = F)
plotDoubScore(healthy_srat)
plotDoubScore(cond_srat)
plotDoubScore(frozen_srat)
dev.off()
```

Plot doublet score UMAP

```{r}
plotUMAPFeat = function(l, f, log_f = F){
  if(log_f){
  return(lapply(names(l), 
                  function(x) FeaturePlot(object = l[[x]], reduction = "umap",
                                          features = f,pt.size = 0.8)+
                  scale_color_gradientn(trans = "log10", guide = "colourbar",
                                        colours = viridis::cividis(150))+
                    labs(title = x)+
                    theme(aspect.ratio = 1)))
  } else{
    return(lapply(names(l), 
                  function(x) FeaturePlot(object = l[[x]], reduction = "umap",
                                          features = f, pt.size = 0.8)+
                    labs(title = x)+
                    theme(aspect.ratio = 1)))
  }
}

pdf("results/initial_stats/healthy_doublet_Score_UMAP.pdf", 
    height = 10, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(healthy_srat, "scrublet_score") 
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_doublet_Score_UMAP.pdf", 
    height = 16, width = 13, useDingbats = F)
plt_list = plotUMAPFeat(cond_srat, "scrublet_score") 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_doublet_Score_UMAP.pdf", 
    height = 6, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(frozen_srat, "scrublet_score") 
cowplot::plot_grid(plotlist = plt_list, nrow = 2, ncol = 2)
dev.off()
```

Testing GMMs to clean doublets  
The scrublet score threshold is based on the assumption that there is a bimodal score distribution, which is dependent on cell diversity

```{r}
plotDoubletGMM = function(l){
  plot_df = data.frame()
  for(n in names(l)){
    plot_df = rbind(plot_df, l[[n]]@meta.data[,c("scrublet_score","unique_name",
                                                 "Fraction","Name")])
  }
  
  gmm_cond = tapply(plot_df$scrublet_score, plot_df$unique_name, 
                    normalmixEM, k = 2)
  
  plot_mix_comps <- function(x, mu, sigma, lam) {
    lam * dnorm(x, mu, sigma)
  }
  
  plt_list = list()
  for(n in names(gmm_cond)){
    mixmdl = gmm_cond[[n]]
    plt_list[[n]] = data.frame(x = mixmdl$x) %>%
      ggplot() +
      geom_histogram(aes(x, ..density..), binwidth = 0.01, colour = "black", 
                     fill = "white") +
      stat_function(geom = "line", fun = plot_mix_comps,
                    args = list(mixmdl$mu[1], mixmdl$sigma[1], lam = mixmdl$lambda[1]),
                    colour = "red", lwd = 0.9) +
      stat_function(geom = "line", fun = plot_mix_comps,
                    args = list(mixmdl$mu[2], mixmdl$sigma[2], lam = mixmdl$lambda[2]),
                    colour = "blue", lwd = 0.9) +
      ylab("Density")+
      labs(title = n)+
      theme_bw()
  }
  
  return(plt_list)
}

healthy_scores = plotDoubletGMM(healthy_srat)
cond_scores = plotDoubletGMM(cond_srat)
frozen_scores = plotDoubletGMM(frozen_srat)

pdf("results/initial_stats/healthy_doublet_scores_gmm.pdf", 
    height = 4, width = 8, useDingbats = F)
cowplot::plot_grid(plotlist = healthy_scores, nrow = 2, ncol = 4)
dev.off()

pdf("results/initial_stats/cond_doublet_scores_gmm.pdf", 
    height = 6, width = 12, useDingbats = F)
cowplot::plot_grid(plotlist = cond_scores, nrow = 4, ncol = 7)
dev.off()

pdf("results/initial_stats/frozen_doublet_scores_gmm.pdf", 
    height = 2, width = 6, useDingbats = F)
cowplot::plot_grid(plotlist = frozen_scores, ncol = 3)
dev.off()
```

Plot cell cycle classification UMAP

```{r}
plotUMAPDim = function(l ,f){
  return(lapply(names(l), 
                function(x) DimPlot(object = l[[x]], reduction = "umap",
                                        group.by = f)+
                  labs(title = x)+
                  theme(aspect.ratio = 1)))
}

pdf("results/initial_stats/healthy_ccPhase_UMAP.pdf", 
    height = 10, width = 6, useDingbats = F)
plt_list = plotUMAPDim(healthy_srat, "Phase") 
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_ccPhase_UMAP.pdf", 
    height = 16, width = 13, useDingbats = F)
plt_list = plotUMAPDim(cond_srat, "Phase") 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_ccPhase_UMAP.pdf", 
    height = 6, width = 6, useDingbats = F)
plt_list = plotUMAPDim(frozen_srat, "Phase") 
cowplot::plot_grid(plotlist = plt_list, nrow = 2, ncol = 2)
dev.off()
```

Plot MT% UMAP

```{r}
pdf("results/initial_stats/healthy_percmt_UMAP.pdf", 
    height = 10, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(healthy_srat, "percent.mt") 
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_percmt_UMAP.pdf", 
    height = 16, width = 13, useDingbats = F)
plt_list = plotUMAPFeat(cond_srat, "percent.mt") 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_percmt_UMAP.pdf", 
    height = 6, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(frozen_srat, "percent.mt") 
cowplot::plot_grid(plotlist = plt_list, nrow = 2, ncol = 2)
dev.off()
```

Plot MT% violin

```{r}
plotViolinFeat = function(l ,f){
  return(lapply(names(l), 
                function(x) VlnPlot(object = l[[x]], features = f, pt.size = 0.15)+
                  labs(title = x)+
                  theme(legend.position = "none",
                        axis.text.x = element_blank())))
}

pdf("results/initial_stats/healthy_percmt_vio.pdf", 
    height = 10, width = 4, useDingbats = F)
plt_list = plotViolinFeat(healthy_srat, "percent.mt") 
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_percmt_vio.pdf", 
    height = 24, width = 9, useDingbats = F)
plt_list = plotViolinFeat(cond_srat, "percent.mt") 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_percmt_vio.pdf", 
    height = 3, width = 6, useDingbats = F)
plt_list = plotViolinFeat(frozen_srat, "percent.mt") 
cowplot::plot_grid(plotlist = plt_list, ncol = 3)
dev.off()
```

Plot ribosomal% UMAP

```{r}
pdf("results/initial_stats/healthy_percrb_UMAP.pdf", 
    height = 10, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(healthy_srat, "percent.rb") 
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_percrb_UMAP.pdf", 
    height = 16, width = 13, useDingbats = F)
plt_list = plotUMAPFeat(cond_srat, "percent.rb") 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_percrb_UMAP.pdf", 
    height = 6, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(frozen_srat, "percent.rb") 
cowplot::plot_grid(plotlist = plt_list, nrow = 2, ncol = 2)
dev.off()
```

Plot ribosomal% violin

```{r}
pdf("results/initial_stats/healthy_percrb_vio.pdf", 
    height = 10, width = 4, useDingbats = F)
plt_list = plotViolinFeat(healthy_srat, "percent.rb") 
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_percrb_vio.pdf", 
    height = 24, width = 9, useDingbats = F)
plt_list = plotViolinFeat(cond_srat, "percent.rb") 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_percrb_vio.pdf", 
    height = 3, width = 6, useDingbats = F)
plt_list = plotViolinFeat(frozen_srat, "percent.rb") 
cowplot::plot_grid(plotlist = plt_list, ncol = 3)
dev.off()
```

Plot total counts UMAP

```{r}
pdf("results/initial_stats/healthy_counts_UMAP.pdf", 
    height = 10, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(healthy_srat, "nCount_RNA", log_f = T)
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_counts_UMAP.pdf", 
    height = 16, width = 13, useDingbats = F)
plt_list = plotUMAPFeat(cond_srat, "nCount_RNA", log_f = T) 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_counts_UMAP.pdf", 
    height = 6, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(frozen_srat, "nCount_RNA", log_f = T) 
cowplot::plot_grid(plotlist = plt_list, nrow = 2, ncol = 2)
dev.off()
```

Plot total genes UMAP

```{r}
pdf("results/initial_stats/healthy_genes_UMAP.pdf", 
    height = 10, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(healthy_srat, "nFeature_RNA", log_f = T)
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_genes_UMAP.pdf", 
    height = 16, width = 13, useDingbats = F)
plt_list = plotUMAPFeat(cond_srat, "nFeature_RNA", log_f = T) 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_genes_UMAP.pdf", 
    height = 6, width = 6, useDingbats = F)
plt_list = plotUMAPFeat(frozen_srat, "nFeature_RNA", log_f = T) 
cowplot::plot_grid(plotlist = plt_list, nrow = 2, ncol = 2)
dev.off()
```

Number of genes vs number of UMI

```{r}
plotCountsVsGenes = function(l){
  plt_l = list()
  for(x in names(l)){
      plot_df = data.frame("nUMI" = l[[x]]$nCount_RNA,
                           "nGene" = l[[x]]$nFeature_RNA)
      
      plt_l[[x]] = ggplot(plot_df, aes(x = nGene, y = nUMI))+
        geom_hex(bins = 65)+
        scale_y_log10()+
        labs(title = x)+
        theme_bw()+
        theme(aspect.ratio = 1)
  }
  
  return(plt_l)
}

pdf("results/initial_stats/healthy_counts_genes.pdf", 
    height = 10, width = 6, useDingbats = F)
plt_list = plotCountsVsGenes(healthy_srat) 
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 2)
dev.off()

pdf("results/initial_stats/cond_counts_genes.pdf", 
    height = 16, width = 13, useDingbats = F)
plt_list = plotCountsVsGenes(cond_srat) 
cowplot::plot_grid(plotlist = plt_list, nrow = 7, ncol = 4)
dev.off()

pdf("results/initial_stats/frozen_counts_genes.pdf", 
    height = 6, width = 6, useDingbats = F)
plt_list = plotCountsVsGenes(frozen_srat) 
cowplot::plot_grid(plotlist = plt_list, nrow = 2, ncol = 2)
dev.off()
```

Plot number of read distribution in the condition donors

```{r}
all_meta = Reduce(rbind, lapply(cond_srat, function(x) x@meta.data))
all_meta$frac_exp_tec = paste(all_meta$Fraction, all_meta$Experiment, sep = "_")

pdf("results/initial_stats/violin_number_reads.pdf", 
    width = 7, height = 6, useDingbats = F)
ggplot(all_meta, aes(x = Name, y = nCount_RNA, fill = frac_exp_tec))+
  geom_violin()+
  scale_y_log10()+
  labs(x = "Donor and Condition", y = "log10(UMI)", fill = "Fraction_Exp_Reseq")+
  theme_bw()+
  theme(legend.position = "bottom")
dev.off()
```

## Filter cells  
We'll be doing a very fine clustering to identify groups of cells with excessive doublet scores/MT%/high or low counts

```{r}
getClusters = function(obj){
  obj <- FindNeighbors(obj, dims = 1:20, verbose = FALSE)
  obj <- FindClusters(obj, algorithm = 2, verbose = F, resolution = 10)
  return(obj)
}

healthy_srat = lapply(healthy_srat, getClusters)
cond_srat = lapply(cond_srat, getClusters)
frozen_srat = lapply(frozen_srat, getClusters)
```

Compare some samples to understand the low count cells

```{r}
sub_cond = cond_srat[c("sc_E1.HEP.Exp1","sc_E6.HEP.Exp1","sc_E1.HEP.Exp2", 
                       "sc_E5.HEP.Exp1","sc_R7.HEP.Exp1","sc_E6.NPC.Exp1")]

markers_sub = list()
for(n in names(sub_cond)){
  sub_cond[[n]] = FindClusters(sub_cond[[n]], algorithm = 2, verbose = F, resolution = 0.6)
  markers_sub[[n]] = quickMarkers(sub_cond[[n]]@assays$RNA@counts,
                                  sub_cond[[n]]@meta.data$SCT_snn_res.0.6, 
                                  N = 20, FDR = 0.05)
}
DimPlot(sub_cond$sc_E6.NPC.Exp1, reduction = "umap", group.by = "SCT_snn_res.0.6")
FeaturePlot(sub_cond$sc_E6.NPC.Exp1, features = c("FGA", "FGB"))
```

Now we calculate those mean stats per cluster and define the thresholds

```{r}
calcMeanClust = function(l, v){
  a = lapply(l, function(x) tapply(x@meta.data[,v], x$SCT_snn_res.10, mean))
  res = reshape2::melt(a)
  a = lapply(l, function(x) table(x$SCT_snn_res.10))
  res2 = reshape2::melt(a)
  res = cbind(res, res2)[,-c(4,6)]
  colnames(res) = c("cluster", "mean_value", "sample", "n_cells")
  return(res)
}

var_list = list()
for(v in c("scrublet_score", "percent.mt", "percent.rb", "nCount_RNA")){
  var_list[[v]] = list("healthy" = calcMeanClust(healthy_srat, v),
                       "cond" = calcMeanClust(cond_srat, v),
                       "frozen" = calcMeanClust(frozen_srat, v))
}

plotMeanClust = function(var_l, l){
  plt_list = list()
  for(n in names(var_l)){
    plt_list[[n]] = ggplot(var_l[[n]][[l]], aes(x = n_cells, y = mean_value, 
                                                colour = sample))+
      geom_point()+
      theme_bw()+
      labs(title = n)+
      theme(aspect.ratio = 1,
            legend.position = "bottom",
            legend.key.size = unit(0.5, "cm"))
    
    if(n=="nCount_RNA") plt_list[[n]] = plt_list[[n]] + scale_y_log10()
    
    leg = cowplot::get_legend(plt_list[[n]])
    plt_list[[n]] = plt_list[[n]]+theme(legend.position = "none")
  }
  
  plt_list$legend = leg
  
  all_plt = cowplot::plot_grid(cowplot::plot_grid(plotlist = plt_list[1:4], 
                                                  ncol = 2, nrow = 2),
                               plt_list[[5]], nrow = 2, rel_heights = c(0.85, 0.2))
  return(all_plt)
}

pdf("results/initial_stats/mean_qc_microclusters.pdf", 
    height = 7, width = 6, useDingbats = F)
plotMeanClust(var_list, "healthy")
plotMeanClust(var_list, "cond")
plotMeanClust(var_list, "frozen")
dev.off()
```

Since  different samples can have very different qualities, we have chosen to define individual filtering thresholds for each sample, which are summarised in the follwing table

```{r}
filt_names = c("scrublet_score_all", "scrublet_score_clusters", 
               "mt_percentage", "min_nUMI", "min_nGenes", "max_nUMI", "max_nGenes")

dat_filt = list("sc_H1.EKS.Exp1" = c(0.25, 0.3, 15, 1000, 800, 40000, 6000),
                "sc_H1.NPC.Exp1" = c(0.25, 0.3, 15, 1000, 800, 40000, 6000),
                "sc_H1.HEP.Exp1" = c(0.3, 0.3, 40, 1000, 800, 40000, 6000),
                "sc_H2.HEP.Exp1" = c(0.3, 0.3, 40, 1000, 800, 40000, 6000),
                "sc_H2.NPC.Exp1" = c(0.25, 0.3, 15, 1000, 800, 40000, 6000),
                "sc_H3.HEP.Exp1" = c(0.3, 0.3, 40, 1000, 800, 40000, 6000),
                "sc_H3.NPC.Exp1" = c(0.25, 0.3, 15, 1000, 800, 40000, 6000),
                
               "sc_E1.HEP.Exp2" = c(0.35, 0.35, 25, 1000, 800, 40000,6000),
               "sc_R1.HEP.Exp2" = c(0.35, 0.35, 25, 1000, 800, 50000,8000),
               "sc_E1.HEP.Exp1" = c(0.3, 0.35, 25, 1000, 800, 40000,6000),
               "sc_R1.HEP.Exp1" = c(0.4, 0.35, 25, 1000, 800, 40000,6000),
               "sc_E2.HEP.Exp1" = c(0.3, 0.35, 20, 1000, 800, 40000,6000),
               "sc_R2.HEP.Exp1" = c(0.5, 0.35, 20, 1000, 800, 40000,6000),
               "sc_E2.NPC.Exp1" = c(0.25, 0.35, 20, 1000, 800, 40000,6000),
               "sc_R2.NPC.Exp1" = c(0.25, 0.35, 20, 1000, 800, 40000,6000),
               "sc_E1.NPC.Exp1" = c(0.25, 0.35, 20, 1000, 800, 40000,6000),
               "sc_R1.NPC.Exp1" = c(0.25, 0.35, 20, 1000, 800, 40000,6000),
               "sc_E3.HEP.Exp1" = c(0.3, 0.35, 30, 1000, 800, 40000,6000),
               "sc_R3.HEP.Exp1" = c(0.3, 0.35, 30, 1000, 800, 40000,6000),
               "sc_E3.NPC.Exp1" = c(0.25, 0.35, 20, 1000, 800, 40000,6000),
               "sc_R3.NPC.Exp1" = c(0.25, 0.35, 20, 1000, 800, 40000,6000),
               "sc_E5.HEP.Exp1" = c(0.4, 0.35, 40, 1000, 450, 40000,6000),
               "sc_R5.HEP.Exp1" = c(0.4, 0.35, 40, 1000, 750, 40000,6000),
               "sc_E5.NPC.Exp1" = c(0.25, 0.35, 25, 1000, 800, 40000,6000),
               "sc_R5.NPC.Exp1" = c(0.25, 0.35, 25, 1000, 800, 40000,6000),
               "sc_E6.HEP.Exp1" = c(0.4, 0.35, 25, 1000, 750, 40000,6000),
               "sc_R6.HEP.Exp1" = c(0.4, 0.35, 25, 1000, 750, 40000,6000),
               "sc_E6.NPC.Exp1" = c(0.25, 0.35, 25, 1000, 800, 40000,6000),
               "sc_R6.NPC.Exp1" = c(0.25, 0.35, 25, 1000, 800, 40000,6000),
               "sc_E7.HEP.Exp1" = c(0.4, 0.35, 25, 1000, 750, 40000,6000),
               "sc_R7.HEP.Exp1" = c(0.4, 0.35, 25, 1000, 700, 40000,6000),
               "sc_E7.NPC.Exp1" = c(0.25, 0.35, 25,1000,800,40000,6000),
               "sc_R7.NPC.Exp1" = c(0.25, 0.35, 25, 1000, 800, 40000,6000),
                  
               "sn_H1.all.Exp1" = c(0.3, 0.3, 5, 400, 300, 40000, 6000),
               "sn_H3.all.Exp1" = c(0.3, 0.3, 5, 400, 300, 40000, 6000),
               "sn_H4.all.Exp1" = c(0.3, 0.3, 5, 400, 300, 40000, 6000))

dat_filt = t(data.frame(dat_filt, stringsAsFactors = F))
colnames(dat_filt) = filt_names
write.csv(dat_filt, "./data/processed/filtering_thresholds_data.csv", 
          col.names = T, row.names = T, quote = F)
```

Apply filters

```{r}
clust_list = data.frame()
for(n in names(var_list$scrublet_score)){
  f = var_list$scrublet_score[[n]]$mean_value<dat_filt[var_list$scrublet_score[[n]]$sample,2]
  clust_list = rbind(clust_list, var_list$scrublet_score[[n]][f,c("cluster", "sample")])
}

filt_perc = data.frame("sample" = "A", "filter" = "A", "perc" = 0)

filt_list = list()
for(n in names(healthy_srat)){
  sub_filts = list(
    cl = healthy_srat[[n]]$SCT_snn_res.10 %in% clust_list$cluster[clust_list$sample==n],
    dou = healthy_srat[[n]]$scrublet_score<dat_filt[n,"scrublet_score_all"],
    mt = healthy_srat[[n]]$percent.mt<dat_filt[n,"mt_percentage"],
    min_umi = healthy_srat[[n]]$nCount_RNA>=dat_filt[n,"min_nUMI"],
    min_gene = healthy_srat[[n]]$nFeature_RNA>=dat_filt[n,"min_nGenes"],
    max_umi = healthy_srat[[n]]$nCount_RNA<=dat_filt[n,"max_nUMI"],
    max_gene = healthy_srat[[n]]$nFeature_RNA<=dat_filt[n,"max_nGenes"]
  )
  
  percs = unlist(lapply(sub_filts, function(x) sum(x)/length(x)*100))
  tot = sum(Reduce("&", sub_filts))/length(Reduce("&", sub_filts))*100
  filt_perc = rbind(filt_perc, data.frame("sample" = n, "filter" = c(names(percs), "total"),
                                          "perc" = c(percs, tot)))
  
  filt_list[[n]] = colnames(healthy_srat[[n]])[Reduce("&", sub_filts)]
}
for(n in names(cond_srat)){
  sub_filts = list(
    cl = cond_srat[[n]]$SCT_snn_res.10 %in% clust_list$cluster[clust_list$sample==n],
    dou = cond_srat[[n]]$scrublet_score<dat_filt[n,"scrublet_score_all"],
    mt = cond_srat[[n]]$percent.mt<dat_filt[n,"mt_percentage"],
    min_umi = cond_srat[[n]]$nCount_RNA>=dat_filt[n,"min_nUMI"],
    min_gene = cond_srat[[n]]$nFeature_RNA>=dat_filt[n,"min_nGenes"],
    max_umi = cond_srat[[n]]$nCount_RNA<=dat_filt[n,"max_nUMI"],
    max_gene = cond_srat[[n]]$nFeature_RNA<=dat_filt[n,"max_nGenes"]
  )
  
  percs = unlist(lapply(sub_filts, function(x) sum(x)/length(x)*100))
  tot = sum(Reduce("&", sub_filts))/length(Reduce("&", sub_filts))*100
  filt_perc = rbind(filt_perc, data.frame("sample" = n, "filter" = c(names(percs), "total"),
                                          "perc" = c(percs, tot)))
  
  filt_list[[n]] = colnames(cond_srat[[n]])[Reduce("&", sub_filts)]
}
for(n in names(frozen_srat)){
  sub_filts = list(
    cl = frozen_srat[[n]]$SCT_snn_res.10 %in% clust_list$cluster[clust_list$sample==n],
    dou = frozen_srat[[n]]$scrublet_score<dat_filt[n,"scrublet_score_all"],
    mt = frozen_srat[[n]]$percent.mt<dat_filt[n,"mt_percentage"],
    min_umi = frozen_srat[[n]]$nCount_RNA>=dat_filt[n,"min_nUMI"],
    min_gene = frozen_srat[[n]]$nFeature_RNA>=dat_filt[n,"min_nGenes"],
    max_umi = frozen_srat[[n]]$nCount_RNA<=dat_filt[n,"max_nUMI"],
    max_gene = frozen_srat[[n]]$nFeature_RNA<=dat_filt[n,"max_nGenes"]
  )
  
  percs = unlist(lapply(sub_filts, function(x) sum(x)/length(x)*100))
  tot = sum(Reduce("&", sub_filts))/length(Reduce("&", sub_filts))*100
  filt_perc = rbind(filt_perc, data.frame("sample" = n, "filter" = c(names(percs), "total"),
                                          "perc" = c(percs, tot)))
  
  filt_list[[n]] = colnames(frozen_srat[[n]])[Reduce("&", sub_filts)]
}
filt_perc = filt_perc[-1,]

# add filter column to meta data
for(n in names(healthy_srat)){
  filt_df = data.frame(row.names = colnames(healthy_srat[[n]]),
                       "keep_allfilt" = colnames(healthy_srat[[n]]) %in% filt_list[[n]])
  healthy_srat[[n]] = AddMetaData(healthy_srat[[n]], filt_df)
}
for(n in names(cond_srat)){
  filt_df = data.frame(row.names = colnames(cond_srat[[n]]),
                       "keep_allfilt" = colnames(cond_srat[[n]]) %in% filt_list[[n]])
  cond_srat[[n]] = AddMetaData(cond_srat[[n]], filt_df)
}
for(n in names(frozen_srat)){
  filt_df = data.frame(row.names = colnames(frozen_srat[[n]]),
                       "keep_allfilt" = colnames(frozen_srat[[n]]) %in% filt_list[[n]])
  frozen_srat[[n]] = AddMetaData(frozen_srat[[n]], filt_df)
}
```

Plot percentage of removed cells

```{r}
filt_perc$filter = factor(filt_perc$filter, 
                          levels = c("cl", "dou", "mt", "min_umi", "min_gene", 
                                     "max_umi", "max_gene", "total"))

pdf("results/initial_stats/percent_cells_filtered.pdf", 
    width = 12, height = 10, useDingbats = F)
ggplot(filt_perc, aes(x = sample, y = perc, fill = filter))+
  facet_wrap(~sample, scales = "free_x")+
  geom_bar(stat="identity", position = "dodge")+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        strip.text = element_text(size = 6.8))
dev.off()
```

Plot number of cells (removed or not)

```{r}
plotFilteredCells = function(l){
  all_meta = Reduce(rbind, lapply(l, function(x) x@meta.data))
  all_meta$frac_exp = paste(all_meta$Fraction, all_meta$Experiment, sep = "_")
  
  plt = ggplot(all_meta, aes(x = frac_exp, fill = keep_allfilt))+
    geom_bar(position = "dodge")+
    facet_wrap(~Name, scales = "free_y")+
    theme_bw()+
    theme(legend.position = "bottom",
          axis.text.x = element_text(angle = 40, hjust = 1, vjust = 1))
  return(plt)
}

pdf("results/initial_stats/healthy_number_cells_filtered.pdf", 
    width = 7, height = 5, useDingbats = F)
plotFilteredCells(healthy_srat)
dev.off()
pdf("results/initial_stats/cond_number_cells_filtered.pdf", 
    width = 7, height = 5, useDingbats = F)
plotFilteredCells(cond_srat)
dev.off()
pdf("results/initial_stats/frozen_number_cells_filtered.pdf", 
    width = 7, height = 5, useDingbats = F)
plotFilteredCells(frozen_srat)
dev.off()
```

Save Seurat object lists before filtering

```{r}
save(healthy_srat, file = "data/processed/healthy_srat_beforeFiltering.RData")
save(cond_srat, file = "data/processed/cond_srat_beforeFiltering.RData")
save(frozen_srat, file = "data/processed/frozen_srat_beforeFiltering.RData")
```

Remove cells

```{r}
filtSeurat = function(srat, filt_col){
  cells = rownames(srat@meta.data)[srat@meta.data$keep_allfilt]
  res = SubsetData(srat, cells = cells)
  
  return(res)
}

healthy_srat = lapply(healthy_srat, filtSeurat)
cond_srat = lapply(cond_srat, filtSeurat)
frozen_srat = lapply(frozen_srat, filtSeurat)
```

## Saving final objects
Save objects

```{r}
save(healthy_srat, file = "data/processed/healthy_srat.RData")
save(cond_srat, file = "data/processed/cond_srat.RData")
save(frozen_srat, file = "data/processed/frozen_srat.RData")
```



# Retired code
Compare technical replicates

```{r, fig.height=5, fig.width=6}
pdf("results/initial_stats/tech_replicates_comparison.pdf",
    useDingbats = F, width = 6, height = 7)
# global correlation
cor_E = cor(as.matrix(cond_srat$Hep2N.data@assays$RNA@counts),
            as.matrix(cond_srat$data.Reg1.NHep@assays$RNA@counts))
common_cells = intersect(rownames(cor_E), colnames(cor_E))
plot_df = data.frame("cor" = diag(cor_E[common_cells, common_cells]))
plt1 = ggplot(plot_df, aes(y = cor, x = "cells"))+
  ggbeeswarm::geom_quasirandom()+
  labs(y = "Pearson correlation", title = "Embolised")+
  theme_bw()

cor_R = cor(as.matrix(cond_srat$Hep2R.data@assays$RNA@counts),
            as.matrix(cond_srat$data.Reg1.RHep@assays$RNA@counts))
common_cells = intersect(rownames(cor_R), colnames(cor_R))
plot_df = data.frame("cor" = diag(cor_R[common_cells, common_cells]))
plt2 = ggplot(plot_df, aes(y = cor, x = "cells"))+
  ggbeeswarm::geom_quasirandom()+
  labs(y = "Pearson correlation", title = "Regenerating")+
  theme_bw()

cowplot::plot_grid(plt1, plt2, ncol = 2)

# compare number of genes
par(mfrow = c(2,1))
common_cells = intersect(rownames(cor_E), colnames(cor_E))
plot(cond_srat$Hep2N.data@meta.data[common_cells,"nFeature_RNA"],
     cond_srat$data.Reg1.NHep@meta.data[common_cells,"nFeature_RNA"],
     xlab = "# genes (rep1)", ylab = "# genes (rep2)",
     main = "# genes = Regenerating")
common_cells = intersect(rownames(cor_R), colnames(cor_R))
plot(cond_srat$Hep2R.data@meta.data[common_cells,"nFeature_RNA"],
     cond_srat$data.Reg1.RHep@meta.data[common_cells,"nFeature_RNA"],
     xlab = "# genes (rep1)", ylab = "# genes (rep2)", 
     main = "# genes = Regenerating")

# compare shared expressed genes
par(mfrow = c(2,1))
common_cells = intersect(rownames(cor_E), colnames(cor_E))
giE_l = list()
for(cc in common_cells){
  g1 = rownames(cond_srat$Hep2N.data@assays$RNA@counts)[cond_srat$Hep2N.data@assays$RNA@counts[,cc]>0]
  g2 = rownames(cond_srat$data.Reg1.NHep@assays$RNA@counts)[cond_srat$data.Reg1.NHep@assays$RNA@counts[,cc]>0]
  
  giE_l[[cc]] = length(intersect(g1, g2))/length(union(g1,g2))*100
}
plot(unlist(giE_l)[order(unlist(giE_l))], ylab = "% shared genes", xlab = "cells",
     main = "Replicate shared genes - Embolised")

common_cells = intersect(rownames(cor_R), colnames(cor_R))
giR_l = list()
for(cc in common_cells){
  g1 = rownames(cond_srat$Hep2R.data@assays$RNA@counts)[cond_srat$Hep2R.data@assays$RNA@counts[,cc]>0]
  g2 = rownames(cond_srat$data.Reg1.RHep@assays$RNA@counts)[cond_srat$data.Reg1.RHep@assays$RNA@counts[,cc]>0]
  
  giR_l[[cc]] = length(intersect(g1, g2))/length(union(g1,g2))*100
}
plot(unlist(giR_l)[order(unlist(giR_l))], ylab = "% shared genes", xlab = "cells",
     main = "Replicate shared genes - Regenerating")

dev.off()
```
