---
title: "R Notebook"
output: html_notebook
---




Load libraries

```{r}
library(Seurat)
library(ggplot2)
library(ggh4x)
library(dplyr)
library(RColorBrewer)

source("https://raw.githubusercontent.com/tomasgomes/personal_toolbox/main/R/analysis_Seurat.R")
```

Load data

```{r}
load("../data/frozen_tissue_dataset.rdata")
fresh_dat = readRDS("../data/processed/allcells_css_reannot.RDS")

cols_use = c("Hepatocyte\ndoublets" = "lightpink2", "Hep/Chol\ndoublets" = "lightpink2",
             "Fibroblasts\n(activated)" = "salmon4", "Fibroblasts" = "salmon3",
             "Fibroblasts 2" = "salmon3", "Fibroblasts 1" = "salmon3", 
             "Fibroblasts 3" = "salmon3", "Stellate cells" = "peru", 
             "LSEC\ndoublets" = "cadetblue2", "VSMC" = "aquamarine4")
colcond = c("healthy" = "orange", "regenerating" = "salmon", 
            "embolised" = "darkred", "embolized" = "darkred")
```



# Fresh data
Subset mesenchymal cells

```{r}
fresh_mes = fresh_dat[,fresh_dat$allcells_major=="Stellate cells"]
fresh_mes = basicSeurat(fresh_mes)

fresh_mes = harmony::RunHarmony(fresh_mes, group.by.vars = "Donor", max.iter.harmony = 50)
fresh_mes = runSeuratClust(fresh_mes, red = "harmony", ncomp = 20)
```

Plot clusters

```{r}
DimPlot(fresh_mes, reduction = "umap", group.by = "Donor")+
  DimPlot(fresh_mes, reduction = "umap", group.by = "Condition")+
  DimPlot(fresh_mes, reduction = "umap", group.by = "harmony20_res.0.6")+
  DimPlot(fresh_mes, reduction = "umap", group.by = "harmony20_res.1")
```

Plot some features

```{r}
FeaturePlot(fresh_mes, reduction = "umap", 
            features = c("COLEC11", "CXCL12", "PDGFRB", "PDGFB", "PDGFRA", "PDGFA", "ACTA2",
                         "DCN", "PPARG", "TGFBI", "TGFB1",  "DES", "VIM", 
                         "LUM", "TNC", "NCAM1", "LRAT", "S100A6", "CCL2"), order = T, pt.size = 0.8)
Seurat::VlnPlot(fresh_mes, c("COLEC11", "CXCL12", "PDGFRB", "PDGFB", "PDGFRA", "PDGFA", "ACTA2",
                           "DCN", "PPARG", "S100A6", "TGFBI", "TGFB1", "DES", "CCL2", "VIM",
                           "LUM", "GFAP", "TNC", "NCAM1", "LRAT", "RBP1"), 
                group.by = "harmony20_res.0.6", ncol = 7)

FeaturePlot(fresh_mes, reduction = "umap", 
            features = c("PDGFRB", "PDGFRA", "LUM", "ACTA2", "SAA1", "PECAM1", "MYL9", "PTPRC", "LYZ"), order = T, pt.size = 0.8)
```

Calculate DE genes

```{r}
fresh_deg = presto::wilcoxauc(fresh_mes, group_by = "harmony20_res.0.6", seurat_assay = "SCT")
fresh_deg = fresh_deg[fresh_deg$padj<=0.05 & fresh_deg$logFC>0.2,]

table(fresh_mes$harmony20_res.0.6, fresh_mes$Condition)
```

Top genes

```{r}
n = 5
topmk = fresh_deg |>
  group_by(group) |> # for each cluster
  filter(padj<=0.05) |> # only p-value below 0.05
  filter(pct_in>(pct_out+10)) |> 
  top_n(n = n, wt = logFC)
```

Make prompt for annotation guess

```{r}
annotationPrompt(fresh_deg, n = 25, pct_diff = 10, 
                 add_info = list("species" = "human", "tissue" = "liver"))
```

DE within each cluster

```{r}
de_cond = list()
for(cl in unique(fresh_mes$harmony20_res.0.6)){
  subdat = fresh_mes[,fresh_mes$harmony20_res.0.6==cl]
  de_cond[[paste0("cl",cl)]] = presto::wilcoxauc(subdat, group_by = "Condition", 
                                                 seurat_assay = "SCT")
  de_cond[[paste0("cl",cl)]] = de_cond[[paste0("cl",cl)]][de_cond[[paste0("cl",cl)]]$padj<=0.05 &
                                                            de_cond[[paste0("cl",cl)]]$logFC>0.2,]
}
de_pve = list()
for(cl in unique(fresh_mes$harmony20_res.0.6)){
  subdat = fresh_mes[,fresh_mes$harmony20_res.0.6==cl]
  subdat$pve = subdat$Condition=="healthy"
  de_pve[[paste0("cl",cl)]] = presto::wilcoxauc(subdat, group_by = "pve", seurat_assay = "SCT")
  de_pve[[paste0("cl",cl)]] = de_pve[[paste0("cl",cl)]][de_pve[[paste0("cl",cl)]]$padj<=0.05 &
                                                          de_pve[[paste0("cl",cl)]]$logFC>0.2,]
}
```



```{r}
fresh_deg_23 = presto::wilcoxauc(fresh_mes[,fresh_mes$harmony20_res.0.6 %in% c(2,3)], 
                                 group_by = "harmony20_res.0.6", seurat_assay = "SCT")
fresh_deg_23 = fresh_deg_23[fresh_deg_23$padj<=0.05 & fresh_deg_23$logFC>0.2,]
fresh_deg_13 = presto::wilcoxauc(fresh_mes[,fresh_mes$harmony20_res.0.6 %in% c(1,3)], 
                                 group_by = "harmony20_res.0.6", seurat_assay = "SCT")
fresh_deg_13 = fresh_deg_13[fresh_deg_13$padj<=0.05 & fresh_deg_13$logFC>0.2,]
fresh_deg_12 = presto::wilcoxauc(fresh_mes[,fresh_mes$harmony20_res.0.6 %in% c(1,2)], 
                                 group_by = "harmony20_res.0.6", seurat_assay = "SCT")
fresh_deg_12 = fresh_deg_12[fresh_deg_12$padj<=0.05 & fresh_deg_12$logFC>0.2,]
```





# Frozen data
Subset and process

```{r}
frozen_mes = frozen_tissue_dataset[,frozen_tissue_dataset$cell.type=="Mesenchyme"]
frozen_mes = basicSeurat(frozen_mes)

frozen_mes = harmony::RunHarmony(frozen_mes, group.by.vars = "donor", max.iter.harmony = 50)

frozen_mes = runSeuratClust(frozen_mes, red = "harmony", ncomp = 20)
```

Plot clusters

```{r}
DimPlot(frozen_mes, reduction = "umap", group.by = "donor")+
DimPlot(frozen_mes, reduction = "umap", group.by = "harmony20_res.0.7", label = T)
```

Plot some markers

```{r}
FeaturePlot(frozen_mes, reduction = "umap", 
            features = c("COLEC11", "CXCL12", "PDGFRB", "PDGFRA", "PDGFA", "ACTA2",
                         "DCN", "PPARG", "S100A6", "TGFBI", "TGFB1", "CLEC4M", "CCL2", "VIM", 
                         "LUM", "ALB", "TNC", "NCAM1", "LRAT", "RBP1"), order = T, pt.size = 0.6)
Seurat::VlnPlot(frozen_mes, c("COLEC11", "CXCL12", "PDGFRB", "PDGFB", "PDGFRA", "PDGFA", "ACTA2",
                           "DCN", "PPARG", "S100A6", "TGFBI", "TGFB1", "DES", "CCL2", "VIM",
                           "LUM", "GFAP", "TNC", "NCAM1", "LRAT", "RBP1"), 
                group.by = "harmony20_res.0.7", ncol = 7)
```

Calculate DE genes

```{r}
frozen_deg = presto::wilcoxauc(frozen_mes, group_by = "harmony20_res.0.7", seurat_assay = "SCT")
frozen_deg = frozen_deg[frozen_deg$padj<=0.05 & frozen_deg$logFC>0.2,]
```

DE pairwise

```{r}
frozen_deg_24 = presto::wilcoxauc(frozen_mes[,frozen_mes$harmony20_res.0.7 %in% c(2,4)], 
                                 group_by = "harmony20_res.0.7", seurat_assay = "SCT")
frozen_deg_24 = frozen_deg_24[frozen_deg_24$padj<=0.05 & frozen_deg_24$logFC>0.2,]
```


Make prompt for annotation guess

```{r}
annotationPrompt(frozen_deg, n = 20, add_info = list("species" = "human", "tissue" = "liver"))
```



# Make figure
The figure should have:
 - table with cluster proportions per condition
 - signature enrichments

UMAPs

```{r}
fresh_mes$annotation = as.character(fresh_mes$harmony20_res.0.6)
fresh_mes$annotation[fresh_mes$annotation=="0"] = "Stellate cells"
fresh_mes$annotation[fresh_mes$annotation=="1"] = "Fibroblasts 1" # not much difference with 2
fresh_mes$annotation[fresh_mes$annotation=="2"] = "Fibroblasts 2"
fresh_mes$annotation[fresh_mes$annotation=="3"] = "Fibroblasts 3"
fresh_mes$annotation[fresh_mes$annotation=="4"] = "Hep/Chol\ndoublets"
fresh_mes$annotation[fresh_mes$annotation=="5"] = "VSMC"
fresh_umap = DimPlot(fresh_mes, reduction = "umap", group.by = "annotation", 
                     label = T, pt.size = 0.4, label.size = 2.5)+
  scale_colour_manual(values = cols_use)+
  theme_classic()+
  theme_void()+
  theme(title = element_blank(),
        aspect.ratio = 1,
        legend.position = "none", 
        plot.background = element_rect(colour = "transparent"))

fresh_mes$Cond2 = fresh_mes$Condition
fresh_mes$Cond2[fresh_mes$Cond2=="embolised"] = "embolized"
fresh_mes$Cond2 = factor(fresh_mes$Cond2, levels = c("healthy", "regenerating", "embolized"))
fresh_umap_cond = DimPlot(fresh_mes, reduction = "umap", group.by = "Cond2", pt.size = 0.4)+
  scale_colour_manual(values = colcond)+
  theme_classic()+
  theme_void()+
  theme(title = element_blank(),
        aspect.ratio = 1,
        legend.key.size = unit(0.4, "cm"),
        legend.text = element_text(size = 6.5),
        legend.position = c(0,1),legend.justification = c(0,1),
        plot.background = element_rect(colour = "transparent"))

frozen_mes$annotation = as.character(frozen_mes$harmony20_res.0.7)
frozen_mes$annotation[frozen_mes$annotation=="0"] = "Stellate cells"
frozen_mes$annotation[frozen_mes$annotation=="1"] = "Stellate cells"
frozen_mes$annotation[frozen_mes$annotation=="2"] = "VSMC"
frozen_mes$annotation[frozen_mes$annotation=="3"] = "Hepatocyte\ndoublets"
frozen_mes$annotation[frozen_mes$annotation=="4"] = "LSEC\ndoublets"
frozen_mes$annotation[frozen_mes$annotation=="5"] = "Fibroblasts"
frozen_umap = DimPlot(frozen_mes, reduction = "umap", group.by = "annotation", 
                      label = T, pt.size = 0.4, label.size = 2.5)+
  scale_colour_manual(values = cols_use)+
  theme_classic()+
  theme_void()+
  theme(title = element_blank(),
        aspect.ratio = 1,
        legend.position = "none", 
        plot.background = element_rect(colour = "transparent"))

pdf("../figure_panels_rev2/fresh_umap.pdf", height = 2.9, width = 2.9)
print(fresh_umap)
dev.off()
pdf("../figure_panels_rev2/fresh_umap_cond.pdf", height = 1.8, width = 1.8)
print(fresh_umap_cond)
dev.off()
pdf("../figure_panels_rev2/frozen_umap.pdf", height = 2.9, width = 2.9)
print(frozen_umap)
dev.off()
```

Marker heatmaps

```{r}
n = 7
top_fresh_mk = fresh_deg |>
  group_by(group) |> # for each cluster
  filter(padj<=0.05) |> # only p-value below 0.05
  filter(pct_in>(pct_out+10)) |> 
  top_n(n = n, wt = logFC)
top_frozen_mk = frozen_deg |>
  group_by(group) |> # for each cluster
  filter(padj<=0.05) |> # only p-value below 0.05
  filter(pct_in>(pct_out+10)) |> 
  top_n(n = n, wt = logFC)
genes_use = union(top_fresh_mk$feature, top_frozen_mk$feature)

avg_fresh = AverageExpression(fresh_mes, assays = "SCT", features = genes_use,
                              group.by = "annotation", slot = "data")$SCT
avg_fresh = t(scale(t(avg_fresh)))
avg_frozen = AverageExpression(frozen_mes, assays = "SCT", features = genes_use,
                              group.by = "annotation", slot = "data")$SCT
avg_frozen = t(scale(t(avg_frozen)))

pdf("../figure_panels_rev2/heatmap_mes_fresh.pdf", height = 6.5, width = 2.2)
annot_col = data.frame(row.names = unique(fresh_mes$annotation), 
                       "cell.type" = unique(fresh_mes$annotation))
pheatmap::pheatmap(avg_fresh, clustering_method = "ward.D2", treeheight_row = 0, 
                   treeheight_col = 10, fontsize_row = 5.5, fontsize_col = 5.5,fontsize = 5.5,
                   annotation_col = annot_col,
                   annotation_colors = list("cell.type" = cols_use),annotation_legend = F,
                   color = colorRampPalette(rev(brewer.pal(n = 9, name = "RdBu")))(100))
dev.off()

pdf("../figure_panels_rev2/heatmap_mes_frozen.pdf", height = 6.5, width = 2.2)
annot_col = data.frame(row.names = unique(frozen_mes$annotation), 
                       "cell.type" = unique(frozen_mes$annotation))
pheatmap::pheatmap(avg_frozen, clustering_method = "ward.D2", treeheight_row = 0, 
                   treeheight_col = 10, fontsize_row = 5.5, fontsize_col = 5.5,fontsize = 5.5,
                   annotation_col = annot_col,
                   annotation_colors = list("cell.type" = cols_use),annotation_legend = F,
                   color = colorRampPalette(rev(brewer.pal(n = 9, name = "RdBu")))(100))
dev.off()
```

Proportions heatmap

```{r}
fresh_mes$annotation2 = fresh_mes$annotation
fresh_mes$annotation2[grepl("Fibrob", fresh_mes$annotation2)] = "Fibroblasts"
tab_cnt = table(fresh_mes$annotation, fresh_mes$Cond2)
tab_ctCond = prop.table(table(fresh_mes$annotation, fresh_mes$Cond2), margin = 2)
hcr = hclust(dist(tab_ctCond), method = "ward.D2")
tab_ctCond = reshape2::melt(tab_ctCond)
tab_ctCond$Var1 = factor(tab_ctCond$Var1, levels = hcr$labels[hcr$order])
tab_cnt = reshape2::melt(tab_cnt)

plt_prop = ggplot()+
  geom_tile(data = tab_ctCond, mapping = aes(x = Var1, y = Var2, fill = value))+
  geom_text(data = tab_cnt, mapping = aes(x = Var1, y = Var2, label = value), 
             size = 2.5, colour = "gray20")+
  labs(x = "cell type", y = "condition", fill = "% (condition)")+
  scale_x_discrete(expand = c(0,0))+
  scale_y_discrete(expand = c(0,0))+
  theme_classic()+
  theme(axis.text = element_text(colour = "black", size = 6),
        axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
        axis.title = element_text(size = 6.5),
        axis.line = element_blank(),
        legend.text = element_text(size = 6, colour = "black"),
        legend.title = element_text(size = 6.5),
        legend.margin = margin(0,0,0,0),
        legend.box.spacing = unit(0.3, "cm"),
        legend.key.width = unit(0.35, "cm"))

pdf("../figure_panels_rev2/prop_fresh_ct.pdf", height = 2, width = 3.25)
print(plt_prop)
dev.off()
```

Signatures

```{r}
mk_hend = readxl::read_xlsx("../data/other/Mesenchyme_markers_fromCell_Henderson.xlsx")
mk_per_ct = tapply(mk_hend$gene, mk_hend$cluster, function(x) toupper(x))
```

Module signature

```{r}
#fresh_mes@meta.data = fresh_mes@meta.data[,-which(colnames(fresh_mes@meta.data) %in% names(mk_per_ct))]
fresh_mes = AddModuleScore(fresh_mes, features = mk_per_ct, assay = "SCT")
colnames(fresh_mes@meta.data)[grepl("Cluster", colnames(fresh_mes@meta.data))] = names(mk_per_ct)
VlnPlot(fresh_mes, features = names(mk_per_ct), group.by = "annotation")

#frozen_mes@meta.data = frozen_mes@meta.data[,-which(colnames(frozen_mes@meta.data) %in% names(mk_per_ct))]
frozen_mes = AddModuleScore(frozen_mes, features = mk_per_ct, nbin = 22, ctrl = 50, assay = "SCT")
colnames(frozen_mes@meta.data)[grepl("Cluster", colnames(frozen_mes@meta.data))] = names(mk_per_ct)
VlnPlot(frozen_mes, features = names(mk_per_ct), group.by = "annotation")
```

DE gene enrichment

```{r}
hypergeometricTest <- function(geneSet, backgroundSet, universeSize) {
  # Calculate the number of genes in the intersection of geneSet and backgroundSet
  intersectionSize <- length(intersect(geneSet, backgroundSet))
  
  # Calculate the number of genes in geneSet
  geneSetSize <- length(geneSet)
  
  # Calculate the number of genes in backgroundSet
  backgroundSetSize <- length(backgroundSet)
  
  # Calculate the number of genes not in either geneSet or backgroundSet
  notInSetSize <- universeSize - (geneSetSize + backgroundSetSize - intersectionSize)
  
  # Perform the hypergeometric test
  p_value <- phyper(intersectionSize - 1, geneSetSize, notInSetSize, backgroundSetSize, 
                    lower.tail = FALSE)
  
  # Create a result object with relevant information
  result <- list(
    geneSetSize = geneSetSize,
    backgroundSetSize = backgroundSetSize,
    intersectionSize = intersectionSize,
    notInSetSize = notInSetSize,
    p_value = p_value
  )
  
  return(result)
}


fresh_mk_l = tapply(fresh_deg$feature, fresh_deg$group, function(x) x)
fresh_enr_res = list()
for(cl in names(fresh_mk_l)){
  for(n in names(mk_per_ct)){
    nn = paste0(cl, "_", n)
    fresh_enr_res[[nn]] = hypergeometricTest(mk_per_ct[[n]], backgroundSet = fresh_mk_l[[cl]],
                                             universeSize = nrow(fresh_mes))
  }
}
fresh_enr_res = reshape2::melt(fresh_enr_res)
fresh_enr_res = fresh_enr_res[fresh_enr_res$L2=="p_value",]
fresh_enr_res$group = sapply(strsplit(fresh_enr_res$L1, "_"), function(x) x[1])
fresh_enr_res$geneset = sapply(strsplit(fresh_enr_res$L1, "_"), function(x) x[2])
fresh_enr_res = reshape2::dcast(fresh_enr_res, group ~ geneset, value.var = "value")
rownames(fresh_enr_res) = fresh_enr_res[,1]
fresh_enr_res = fresh_enr_res[,-1]


frozen_mk_l = tapply(frozen_deg$feature, frozen_deg$group, function(x) x)
frozen_enr_res = list()
for(cl in names(frozen_mk_l)){
  for(n in names(mk_per_ct)){
    nn = paste0(cl, "_", n)
    frozen_enr_res[[nn]] = hypergeometricTest(mk_per_ct[[n]], backgroundSet = frozen_mk_l[[cl]],
                                             universeSize = nrow(frozen_mes))
  }
}
frozen_enr_res = reshape2::melt(frozen_enr_res)
frozen_enr_res = frozen_enr_res[frozen_enr_res$L2=="p_value",]
frozen_enr_res$group = sapply(strsplit(frozen_enr_res$L1, "_"), function(x) x[1])
frozen_enr_res$geneset = sapply(strsplit(frozen_enr_res$L1, "_"), function(x) x[2])
frozen_enr_res = reshape2::dcast(frozen_enr_res, group ~ geneset, value.var = "value")
rownames(frozen_enr_res) = frozen_enr_res[,1]
frozen_enr_res = frozen_enr_res[-2,-1] # removing the second HSC cluster


annot_fresh = unique(fresh_mes$annotation)
names(annot_fresh) = unique(as.character(fresh_mes$harmony20_res.0.6))
annot_frozen = c("LSEC\ndoublets", "VSMC", "Stellate cells", "Stellate cells", 
                 "Fibroblasts", "Hepatocyte\ndoublets")
names(annot_frozen) = c("4", "2", "0", "1", "5", "3")

rownames(fresh_enr_res) = annot_fresh[rownames(fresh_enr_res)]
rownames(frozen_enr_res) = annot_frozen[rownames(frozen_enr_res)]

fresh_pval_df = reshape2::melt(t(fresh_enr_res))
fresh_pval_df$Var2 = factor(fresh_pval_df$Var2, levels = sort(unique(fresh_mes$annotation)))
frozen_pval_df = reshape2::melt(t(frozen_enr_res))
frozen_pval_df$Var2 = factor(frozen_pval_df$Var2, levels = sort(unique(frozen_mes$annotation)))
```

Combined plot

```{r}
plot_scores = rbind(fresh_mes@meta.data[,c("annotation", "FB", "HSC", "VSMC")], 
                    frozen_mes@meta.data[,c("annotation", "FB", "HSC", "VSMC")])
plot_scores$protocol = c(rep("fresh", ncol(fresh_mes)), rep("frozen", ncol(frozen_mes)))
plot_scores = reshape2::melt(plot_scores)

scores_plt = ggplot(plot_scores, aes(x = annotation, y = value, colour = annotation))+
  ggh4x::facet_grid2(variable ~ protocol, independent = "y", scales = "free")+
  ggbeeswarm::geom_quasirandom(size = 0.4)+
  scale_colour_manual(values = cols_use)+
  theme_bw()+
  theme(axis.text = element_text(colour = "black", size = 6),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.title = element_text(size = 6.5),
        axis.line = element_blank(),
        axis.title.x = element_blank(),
        strip.text = element_text(size = 7),
        legend.position = "none")
fresh_heat = ggplot(fresh_pval_df, aes(y = Var1, x = Var2, fill = -log10(value)))+
  geom_tile()+
  labs(fill = "-log10(p-value)", y = "gene signatures")+
  scale_x_discrete(expand = c(0,0))+
  scale_y_discrete(expand = c(0,0))+
  guides(fill = guide_colourbar(title.position="top", title.hjust = 0))+
  theme_classic()+
  theme(axis.text.y = element_text(colour = "black", size = 6),
        axis.title.y = element_text(colour = "black", size = 7),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.line = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "bottom",
        legend.text = element_text(size = 6, colour = "black"),
        legend.title = element_text(size = 6.5, vjust = 1),
        legend.margin = margin(0,0,0,0),
        legend.box.margin = margin(0,0,0,0),
        legend.box.spacing = unit(0.25, "cm"),
        legend.key.height = unit(0.35, "cm"),
        legend.key.width = unit(0.45, "cm"))
frozen_heat = ggplot(frozen_pval_df, aes(y = Var1, x = Var2, fill = -log10(value)))+
  geom_tile()+
  labs(fill = "-log10(p-value)", y = "gene signatures")+
  scale_x_discrete(expand = c(0,0))+
  scale_y_discrete(expand = c(0,0))+
  guides(fill = guide_colourbar(title.position="top", title.hjust = 0))+
  theme_classic()+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.line = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "bottom",
        legend.text = element_text(size = 6, colour = "black"),
        legend.title = element_text(size = 6.5, vjust = 1),
        legend.margin = margin(0,0,0,0),
        legend.box.margin = margin(0,0,0,0),
        legend.box.spacing = unit(0.25, "cm"),
        legend.key.height = unit(0.35, "cm"),
        legend.key.width = unit(0.45, "cm"))

sc_plt_all = scores_plt+(fresh_heat+frozen_heat)+patchwork::plot_layout(heights = c(5, 1), 
                                                                        nrow = 2)
pdf("../figure_panels_rev2/scores_signatures.pdf", height = 5, width = 3.25)
print(sc_plt_all)
dev.off()
```



