---
title: "Zonation analysis V2"
output: html_notebook
---

This is the second version of the Zonation Analysis notebook. Decided to branch to start anew and have the notebook more organised.



# General Setup
Setup chunk

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_knit$get("root.dir")
```

Setup reticulate

```{r}
library(reticulate)
knitr::knit_engines$set(python = reticulate::eng_python)
py_available(initialize = FALSE)
use_python(Sys.which("python"))
py_config()
```

Load libraries

```{r}
library(Seurat)
library(ggplot2)
library(ggridges)
library(dplyr)
library(destiny)
library(fdrtool)
library(gam)
library(dtw)
library(data.table)
```



# Load data
Load data (from all cells)

```{r}
allcells_css = readRDS(file = "data/processed/allcells_css.RDS")
```



# Hepatocytes
Subset and process each condition

```{r}
hep_cells = list()
for(cond in unique(allcells_css@meta.data$Condition)){
  cat(cond)
  hep_cells[[cond]] = allcells_css[,allcells_css@meta.data$allcells_major=="Hepatocytes" &
                                     allcells_css@meta.data$Condition==cond]
  print(dim(hep_cells[[cond]]))
  hep_cells[[cond]] = suppressWarnings(SCTransform(hep_cells[[cond]], do.correct.umi = T, 
                                                   verbose = F, 
                                                   vars.to.regress=c("unique_name","nCount_RNA"),
                                                   variable.features.rv.th = 1, seed.use = 1,
                                                   return.only.var.genes = F, 
                                                   variable.features.n = NULL))
  hep_cells[[cond]] = RunPCA(hep_cells[[cond]], verbose = F)
  hep_cells[[cond]] = RunUMAP(hep_cells[[cond]], dims = 1:20, verbose = F)
}
```


## Healthy analysis
For healthy cells (which will serve as reference), do additional filtering and reprocessing

```{r}
DimPlot(hep_cells$healthy, reduction = "umap", group.by = "unique_name")
outcells = (hep_cells$healthy@reductions$umap@cell.embeddings[,1]>(1) &
              hep_cells$healthy@reductions$umap@cell.embeddings[,2]>5)
hep_cells$healthy = hep_cells$healthy[,!outcells]

hep_cells$healthy = suppressWarnings(SCTransform(hep_cells$healthy, do.correct.umi = T, 
                                                 verbose = F,
                                                 vars.to.regress = c("unique_name","nCount_RNA"),
                                                 variable.features.rv.th = 1, seed.use = 1,
                                                 return.only.var.genes = F, 
                                                 variable.features.n = NULL))
hep_cells$healthy = RunPCA(hep_cells$healthy, verbose = F)
hep_cells$healthy = RunUMAP(hep_cells$healthy, dims = 1:15, verbose = F)
DimPlot(hep_cells$healthy, reduction = "umap", group.by = "unique_name")

hep_cells$healthy = FindNeighbors(hep_cells$healthy, reduction = "pca", dims = 1:15,
                                  force.recalc = T, graph.name = "hep")
hep_cells$healthy = FindClusters(hep_cells$healthy, graph.name = "hep", algorithm = 4, 
                                 resolution = seq(0.1,0.8,0.1), verbose = F)
DimPlot(hep_cells$healthy, reduction = "umap", group.by = "hep_res.0.3")
hep_cells$healthy = SetIdent(hep_cells$healthy, value = "hep_res.0.6")
mk_hep_h = FindAllMarkers(hep_cells$healthy, assay = "SCT")

saveRDS(hep_cells$healthy, file = "results/zonation_healthy/hep_healthy_srat.RDS")
```

Save hepatocytes

```{r}
saveRDS(hep_cells, file="results/zonation_cond/hep_cells.RDS")
```

Plot expression of zonation markers

```{r}
gfresh = c("CYP1A2", "CYP2E1", "GLUL", "CYP3A4", "G0S2", "APOC1", "UGT2B17", "BCHE", # central 
           "SAA1", "HAMP", "SAA2", "CPS1", "C3", "C1R", "MT-ND4", "MT-CO2") # portal

png("results/zonation_healthy/healthy_hepatocyte_zonation_markers_violin.png", 
    height = 1400, width = 1400)
VlnPlot(hep_cells$healthy, features = gfresh, group.by = "names_clusters", 
        pt.size = 0, sort = "increasing")
dev.off()
png("results/zonation_healthy/healthy_hepatocyte_zonation_markers_umap.png", 
    height = 1400, width = 1400)
FeaturePlot(hep_cells$healthy, features = gfresh, pt.size = 0.6)
dev.off()
```

Obtain Diffusion maps projection

```{r}
dm = DiffusionMap(hep_cells$healthy@reductions$pca@cell.embeddings[,1:15], 
                  rotate = T, n_eigs = 5)
dpt = DPT(dm)
plot(dpt)
```

Check expression of some genes

```{r}
df_dc = cbind(hep_cells$healthy@meta.data, 
              data.frame("DC1" = dpt@dm$DC1, "DC2" = dpt@dm$DC2),
              Matrix::t(hep_cells$healthy@assays$SCT@data[gfresh,]))
colnames(df_dc) = gsub("-", ".", colnames(df_dc), fixed = T)

plt_list = list()
for(n in c(gsub("-", ".", gfresh, fixed = T), "names_clusters", "unique_name")){
  plt_list[[n]] = ggplot(df_dc, aes_string(x = "DC1", y = "DC2", colour = n))+
    geom_point()+
    theme_classic()
}

png("results/zonation_healthy/healthy_hepatocyte_zonation_markers_dc.png", 
    height = 900, width = 2800)
cowplot::plot_grid(plotlist = plt_list, nrow = 3, ncol = 6, align = "hv")
dev.off()
```

We'll be using DC1 as the pseudotime. This was the best approximation possible to the zonation gradient based on diffusion maps

```{r}
pdf("results/zonation_healthy/hep_distributions_dc1.pdf", useDingbats = F, 
    height = 4, width = 9)
plt1 = ggplot(df_dc, aes(x = DC1, y = names_clusters, 
                  group = names_clusters, fill = names_clusters))+
  geom_density_ridges2(alpha = 0.4)+
  theme_bw()+
  theme(legend.position = "none")

plt2 = ggplot(df_dc, aes(x = DC1, y = unique_name, 
                  group = unique_name, fill = unique_name))+
  geom_density_ridges2(alpha = 0.4)+
  theme_bw()+
  theme(legend.position = "none")

cowplot::plot_grid(plt1, plt2, ncol = 2, align = "h")
dev.off()
```

Find genes varying along the pseudotime

```{r}
# will only use variable genes
hvg = hep_cells$healthy@assays$SCT@var.features

# Fit GAM for each gene using pseudotime as independent variable.
t = df_dc$DC1
gene_fit_p = c()
gene_fit_vals = data.frame(row.names = 1:100)
for(i in 1:length(hvg)){
  g = hvg[i]
  z = hep_cells$healthy@assays$SCT@data[g,]
  
  d = data.frame(z=z, t=t)
  tmp = suppressMessages(gam(z ~ ns(t, df = 4), data=d))
  
  # bins for model fitting
  bb = seq(min(t), max(t), length.out = 100)
  gene_fit_vals[,g] = suppressMessages(predict(tmp,
                                               newdata = data.frame(t = bb)))
  p = summary(tmp)$parametric.anova$`Pr(>F)`[1]
  gene_fit_p = c(gene_fit_p, p)
  names(gene_fit_p)[length(gene_fit_p)] = g
}

gene_fit_p = fdrtool::fdrtool(gene_fit_p, statistic="pvalue", plot=F, 
                              verbose=F, cutoff.method="pct0", pct0=0.9)$qval
```


## Analysis of other conditions
Now use the top 500 genes to learn the pseudozonation coordinates and project the embolised and regenerating data

```{r}
sig_genes = gene_fit_p[gene_fit_p<=0.05]
top_sig_genes = names(sig_genes[order(sig_genes, decreasing = F)][1:500]) 
# using all genes was identical

data_gam_h = cbind(data.frame("DC1" = df_dc$DC1),
                   Matrix::t(hep_cells$health@assays$SCT@data[top_sig_genes,]))

gam_healthy = gam::gam(DC1 ~ ., data = data_gam_h)

hea_pred = predict(gam_healthy,
                   Matrix::t(hep_cells$health@assays$SCT@data[top_sig_genes,]))
reg_pred = predict(gam_healthy,
                   Matrix::t(hep_cells$regenerating@assays$SCT@data[top_sig_genes,]))
emb_pred = predict(gam_healthy,
                   Matrix::t(hep_cells$embolised@assays$SCT@data[top_sig_genes,]))

plot(density(reg_pred), col = "blue")
lines(density(emb_pred), col = "red")
lines(density(data_gam_h$DC1), col = "green")
lines(density(hea_pred), col = "grey50")
legend(-0.073,32, legend = c("healthy (DC1)", "healthy (pred)", 
                         "embolised (pred)", "regenerating (pred)"),
       fill = c("green", "grey50", "red", "blue"))

traj_list = list("healthy" = df_dc$DC1, "embolised" = emb_pred, "regenerating" = reg_pred)
```

Calculate correlations

```{r}
cor_h = t(cor(traj_list$healthy, as.matrix(Matrix::t(hep_cells$healthy@assays$SCT@data)),
              method = "sp"))
cor_emb = t(cor(traj_list$embolised, as.matrix(Matrix::t(hep_cells$embolised@assays$SCT@data)),
                method = "sp"))
cor_reg = t(cor(traj_list$regenerating, as.matrix(Matrix::t(hep_cells$regenerating@assays$SCT@data)),
                method = "sp"))

l = list(cor_h, cor_emb, cor_reg)
all_corr = Reduce(function(x, y) merge(x,y, by = "rn", all = T), 
                  lapply(l, function(x) data.frame(x, rn = row.names(x))))
colnames(all_corr) = c("genes", "healthy", "embolised", "regenerating")

write.csv(all_corr, file = paste0("results/zonation_cond/hep_correlations_zonation.csv"),
          col.names = T, row.names = F, quote = F)
```

Find varying genes

```{r}
qvals_list = list()
fits_list = list()

# get HVG from all conditions
hvg = unique(c(hep_cells[["healthy"]]@assays$SCT@var.features,
               hep_cells[["embolised"]]@assays$SCT@var.features,
               hep_cells[["regenerating"]]@assays$SCT@var.features))
hvg = hvg[hvg %in% rownames(hep_cells[["healthy"]]@assays$SCT@data) &
            hvg %in% rownames(hep_cells[["embolised"]]@assays$SCT@data) &
            hvg %in% rownames(hep_cells[["regenerating"]]@assays$SCT@data)]

for(cond in names(hep_cells)){
  print(cond)
  # Fit GAM for each gene using pseudotime as independent variable.
  t = traj_list[[cond]]
  gene_fit_p = c()
  gene_fit_vals = data.frame(row.names = 1:100)
  for(i in 1:length(hvg)){
    g = hvg[i]
    z = hep_cells[[cond]]@assays$SCT@data[g,]
    
    d = data.frame(z=z, t=t)
    tmp = suppressMessages(gam(z ~ ns(t, df = 4), data=d))
    
    # bins for model fitting
    bb = seq(min(t), max(t), length.out = 100)
    gene_fit_vals[,g] = suppressMessages(predict(tmp,
                                                 newdata = data.frame(t = bb)))
    p = summary(tmp)$parametric.anova$`Pr(>F)`[1]
    gene_fit_p = c(gene_fit_p, p)
    names(gene_fit_p)[length(gene_fit_p)] = g
  }
  
  if(sum(is.na(gene_fit_p))>0){
    gene_fit_p[is.na(gene_fit_p)] = 1
  }
  gene_fit_p = fdrtool::fdrtool(gene_fit_p, statistic="pvalue", plot=F, 
                                verbose=F, cutoff.method="pct0", pct0=0.9)$qval
  
  qvals_list[[cond]] = gene_fit_p
  fits_list[[cond]] = gene_fit_vals
}
```

Sabe trajectories and genes

```{r}
save(fits_list, qvals_list, traj_list, 
     file = "results/zonation_cond/hep_traj_qval_fits.RData")
```

Calculate distances for all significant genes

```{r}
for(cond in names(qvals_list)){
  gene_fit_p = qvals_list[[cond]]
  gene_fit_vals = fits_list[[cond]]
  gene_fit_vals_filt = t(gene_fit_vals[,names(gene_fit_p)[gene_fit_p<=0.05]])
  scale_fit = t(apply(gene_fit_vals_filt, 1, scale))
  
  # dist
  if(file.exists(paste0("results/zonation_cond/dtw_distance_", cond, "_sigGenes.RDS"))){
    next
  } else{
    dd = dtw::dtwDist(scale_fit) # TAKES 4.5 HOURS
    saveRDS(dd, paste0("results/zonation_cond/dtw_distance_", cond, "_sigGenes.RDS"))
  }
}
```

Cluster genes (per condition)

```{r}
top_n = 1000
for(cond in names(qvals_list)){
  print(cond)
  gene_fit_p = qvals_list[[cond]]
  gene_fit_vals = fits_list[[cond]]
  
  dd = readRDS(paste0("results/zonation_cond/dtw_distance_", cond, "_sigGenes.RDS"))
  
  gene_fit_vals_filt = t(gene_fit_vals[,names(gene_fit_p)[gene_fit_p<=0.05]])
  scale_fit = t(apply(gene_fit_vals_filt, 1, scale))
  
  # select top genes (0 if using all)
  topgenes = if(top_n==0) names(gene_fit_p) else names(head(gene_fit_p[order(gene_fit_p,
                                                                             decreasing = F)],
                                                            top_n))
  
  clust_genes = hclust(as.dist(dd[topgenes,topgenes]), method = "ward.D2")
  
  cls = cutree(clust_genes, 4)
  #ape::plot.phylo(ape::as.phylo(clust_genes), tip.color = rainbow(10)[cls])
  plt_list = list()
  for(i in unique(cls)){
    sub_df = scale_fit[names(cls)[cls==i],]
    plot_df = reshape2::melt(sub_df)
    plt_list[[paste0("gene_",i)]] = ggplot(plot_df, aes(x = Var2, y = value, group = Var1))+
      geom_hline(yintercept = 0, linetype = "dashed")+
      geom_line(alpha = 0.1, colour = "grey69")+
      stat_summary(aes(y = value, group=1), fun=mean, colour="red", geom="line",group=1)+
      labs(title = paste0("gene_",i))+
      theme_bw()
  }
  
  pdf(paste0("results/zonation_cond/hep_gene_expression_profile_", cond, ".pdf"),
      height = 4.5, width = 4.4)
  print(cowplot::plot_grid(plotlist = plt_list, nrow = 2, ncol = 2, align = "hv"))
  dev.off()
  
  df_pval_cl = merge(data.frame(gene_fit_p), data.frame(cls), by = 0, all = T)
  colnames(df_pval_cl) = c("gene", "pval", "cluster")
  
  write.csv(df_pval_cl, file = paste0("results/zonation_cond/hep_pval_clust_", cond, ".csv"),
            col.names = T, row.names = F, quote = F)

  write.csv(gene_fit_vals, file = paste0("results/zonation_cond/hep_gene_fits_", cond, ".csv"),
            col.names = T, row.names = F, quote = F)
}
```

Plot cell distributions in pseudotime

```{r}
plot_df = data.frame("pseudotime" = c(traj_list$healthy, traj_list$embolised,
                                      traj_list$regenerating),
           "donor" = c(hep_cells$healthy@meta.data[,"Donor"],
                       hep_cells$embolised@meta.data[names(traj_list$embolised),"Donor"],
                       hep_cells$regenerating@meta.data[names(traj_list$regenerating),"Donor"]),
           "cond" = c(rep("healthy", length(traj_list$healthy)), 
                      rep("embolised", length(traj_list$embolised)),
                      rep("regenerating", length(traj_list$regenerating))))

pdf("results/zonation_cond/hep_cell_distributions_zonation.pdf", useDingbats = F, 
    width = 6, height = 5)
ggplot(plot_df, aes(x = donor, y = pseudotime, fill = cond))+
  geom_boxplot(notch = T)+
  theme_bw()

ggplot(plot_df, aes(x = pseudotime, group = donor, colour = donor))+
  facet_wrap(~cond)+
  geom_density()+
  theme_bw()

ggplot(plot_df, aes(x = pseudotime, colour = cond))+
  geom_density()+
  theme_bw()
dev.off()
```

Do cross validation on the healthy trajectory

```{r}
sig_genes = qvals_list$healthy[qvals_list$healthy<=0.05]
top_sig_genes = names(sig_genes[order(sig_genes, decreasing = F)][1:500]) 

data_gam_h = cbind(data.frame("DC1" = traj_list$healthy),
                   Matrix::t(hep_cells$health@assays$SCT@data[top_sig_genes,]))
colnames(data_gam_h) = gsub("-", "_", colnames(data_gam_h), fixed = T)
colnames(data_gam_h) = gsub(".", "_", colnames(data_gam_h), fixed = T)

CVgam = function (formula, data, nfold = 10, debug.level = 0, method = "glm.fit", 
    printit = TRUE, cvparts = NULL, gamma = 1, seed = 29){
  # modified from the `gamclass` package to use gam::gam
    if (is.null(cvparts)) {
        set.seed(seed)
        cvparts <- sample(1:nfold, nrow(data), replace = TRUE)
    }
    folds <- unique(cvparts)
    khat <- hat <- numeric(nrow(data))
    scale.gam <- summary(gam::gam(formula, data = data, method = method))$scale
    for (i in folds) {
        trainrows <- cvparts != i
        testrows <- cvparts == i
        elev.gam <- gam::gam(formula, data = data[trainrows, ], method = method, 
            gamma = gamma)
        hat[testrows] <- predict(elev.gam, newdata = data[testrows, 
            ], select = TRUE)
        res <- residuals(elev.gam)
    }
    y <- eval(formula[[2]], envir = as.data.frame(data))
    res <- y - hat
    cvscale <- sum(res^2)/length(res)
    prntvec <- c(GAMscale = scale.gam, `CV-mse-GAM ` = cvscale)
    if (printit) 
        print(round(prntvec, 4))
    invisible(list(fitted = hat, resid = res, cvscale = cvscale, 
        scale.gam = scale.gam))
}

cv_gam = CVgam(formula = DC1 ~ ., data = data_gam_h, nfold = 10, seed = 1, method = "glm.fit")

pdf("results/zonation_healthy/hep_original_fitted_pseudotime.pdf", useDingbats = F, 
    width = 6, height = 5)
plot(data_gam_h$DC1, cv_gam$fitted, xlab = "Healthy pseudotime (DC1)",ylab = "Predicted Healthy",
     main = paste0("Healthy original vs fitted pseudotime\nMSE: ", round(cv_gam$cvscale, 6)),
     pch = 19, cex = 0.5)
abline(0,1, col = "blue", lwd = 3)
dev.off()
```

Save zonation results

```{r}
# define the intervals based on healthy
df_h = data.table("zonation_pt" = traj_list$healthy)
df_h$zonation_int = cut(df_h$zonation_pt, 3)

df_e = data.table("zonation_pt" = traj_list$embolised)
df_e = df_h[df_e, on="zonation_pt", roll=Inf, rollends = T]

df_r = data.table("zonation_pt" = traj_list$regenerating)
df_r = df_h[df_r, on="zonation_pt", roll=Inf, rollends = T]

list_df = list("healthy" = df_h, "embolised" = df_e, "regenerating" = df_r)

for(n in names(traj_list)){
  ptdf = list_df[[n]]
  rownames(ptdf) = colnames(hep_cells[[n]])
  hep_cells[[n]] = AddMetaData(hep_cells[[n]], metadata = ptdf)
  write.csv(hep_cells[[n]]@meta.data[,c("unique_name", "Condition", 
                                        "zonation_pt", "zonation_int")],
            file = paste0("../results/zonation_cond/hep_zonation_", n, ".csv"),
            col.names = T, row.names = T, quote = F)
}

saveRDS(hep_cells, file = "results/zonation_cond/hep_cells_zonation.RDS")
```



# Endothelial cells
Subset and process each condition (only LSEC)

```{r}
end_cells = list()
for(cond in unique(allcells_css@meta.data$Condition)){
  cat(cond)
  end_cells[[cond]] = allcells_css[,grepl("LSEC ", allcells_css@meta.data$allcells_clusters) &
                                     allcells_css@meta.data$Condition==cond]
  if(cond=="healthy"){
    end_cells[[cond]] = end_cells[[cond]][,grepl("LSEC ", end_cells[[cond]]@meta.data$names_clusters)]
  }
  print(dim(end_cells[[cond]]))
  end_cells[[cond]] = suppressWarnings(SCTransform(end_cells[[cond]], do.correct.umi = T, 
                                                   verbose = F, 
                                                   vars.to.regress=c("unique_name","nCount_RNA"),
                                                   variable.features.rv.th = 1, seed.use = 1,
                                                   return.only.var.genes = F, 
                                                   variable.features.n = NULL))
  end_cells[[cond]] = RunPCA(end_cells[[cond]], verbose = F)
  end_cells[[cond]] = RunUMAP(end_cells[[cond]], dims = 1:20, verbose = F)
}
```


## Healthy analysis
For healthy cells (which will serve as reference), do additional filtering and reprocessing

```{r}
DimPlot(end_cells$healthy, reduction = "umap", group.by = "allcells_clusters")
gfresh = c("CLEC1B", "CLEC4G", "LYVE1", "CD14",  # central 
           "PECAM1", "AQP1", "VWF", "CD34", # portal
           "ALB", "MKI67", "TRAC", "DCN") 
FeaturePlot(end_cells$healthy, features = gfresh, reduction = "umap")

outcells = (end_cells$healthy@reductions$umap@cell.embeddings[,1]<(-4)) # this removes what seem to be central hepatocytes
end_cells$healthy = end_cells$healthy[,!outcells]

end_cells$healthy = suppressWarnings(SCTransform(end_cells$healthy, do.correct.umi = T, 
                                                 verbose = F,
                                                 vars.to.regress = c("unique_name","nCount_RNA"),
                                                 variable.features.rv.th = 1, seed.use = 1,
                                                 return.only.var.genes = F, 
                                                 variable.features.n = NULL))
end_cells$healthy = RunPCA(end_cells$healthy, verbose = F)
end_cells$healthy = RunUMAP(end_cells$healthy, dims = 1:15, verbose = F)
DimPlot(end_cells$healthy, reduction = "umap", group.by = "unique_name")

end_cells$healthy = FindNeighbors(end_cells$healthy, reduction = "pca", dims = 1:15,
                                  force.recalc = T, graph.name = "end")
end_cells$healthy = FindClusters(end_cells$healthy, graph.name = "end", algorithm = 2, 
                                 resolution = seq(0.1,0.8,0.1), verbose = F)
DimPlot(end_cells$healthy, reduction = "umap", group.by = "end_res.0.2")
end_cells$healthy = SetIdent(end_cells$healthy, value = "end_res.0.2")
mk_end_h = FindAllMarkers(end_cells$healthy, assay = "SCT")

saveRDS(end_cells$healthy, file = "results/zonation_healthy/end_healthy_srat.RDS")
```

Save endothelial cells

```{r}
saveRDS(end_cells, file="results/zonation_cond/end_cells.RDS")
```

Plot expression of zonation markers

```{r}
gfresh = c("CLEC1B", "CLEC4G", "LYVE1", "CD14",  # central 
           "PECAM1", "AQP1", "VWF", "CD34") # portal

png("results/zonation_healthy/healthy_hepatocyte_zonation_markers_violin.png", 
    height = 1400, width = 1400)
VlnPlot(end_cells$healthy, features = gfresh, group.by = "names_clusters", 
        pt.size = 0, sort = "increasing")
dev.off()
png("results/zonation_healthy/healthy_hepatocyte_zonation_markers_umap.png", 
    height = 1400, width = 1400)
FeaturePlot(end_cells$healthy, features = gfresh, pt.size = 0.6)
dev.off()
```

Obtain Diffusion maps projection

```{r}
set.seed(1)
dm = DiffusionMap(end_cells$healthy@reductions$pca@cell.embeddings[,1:4], 
                  rotate = T, n_eigs = 5)
dpt = DPT(dm, tips = c(2382, 1379, 424))
plot(dpt, col_by = "DPT424")
plot(dpt, col_by = "branch")
```

Check expression of some genes

```{r}
df_dc = cbind(end_cells$healthy@meta.data, 
              data.frame("DC1" = dpt@dm$DC1, "DC2" = dpt@dm$DC2, "DPT" = dpt$DPT424),
              Matrix::t(end_cells$healthy@assays$SCT@data[gfresh,]))
colnames(df_dc) = gsub("-", ".", colnames(df_dc), fixed = T)
df_dc = cbind(df_dc, end_cells$healthy@reductions$umap@cell.embeddings)

plt_list = list()
for(n in c(gsub("-", ".", gfresh, fixed = T), "names_clusters", "unique_name")){
  plt_list[[n]] = ggplot(df_dc, aes_string(x = "DC1", y = "DC2", colour = n))+
    geom_point()+
    theme_classic()
}

png("results/zonation_healthy/healthy_endothelial_zonation_markers_dc.png", 
    height = 900, width = 2800)
cowplot::plot_grid(plotlist = plt_list, nrow = 4, ncol = 3, align = "hv")
dev.off()
```

We'll be using DPT424 as the pseudotime. This was the best approximation possible to the zonation gradient based on diffusion maps

```{r}
pdf("results/zonation_healthy/end_distributions_dpt424.pdf", useDingbats = F, 
    height = 4, width = 9)
plt1 = ggplot(df_dc, aes(x = DPT, y = names_clusters, 
                  group = names_clusters, fill = names_clusters))+
  geom_density_ridges2(alpha = 0.4)+
  theme_bw()+
  theme(legend.position = "none")

plt2 = ggplot(df_dc, aes(x = DPT, y = unique_name, 
                  group = unique_name, fill = unique_name))+
  geom_density_ridges2(alpha = 0.4)+
  theme_bw()+
  theme(legend.position = "none")

cowplot::plot_grid(plt1, plt2, ncol = 2, align = "h")
dev.off()
```

Find genes varying along the pseudotime

```{r}
# will only use variable genes
hvg = end_cells$healthy@assays$SCT@var.features

# Fit GAM for each gene using pseudotime as independent variable.
t = df_dc$DPT
gene_fit_p = c()
gene_fit_vals = data.frame(row.names = 1:100)
for(i in 1:length(hvg)){
  g = hvg[i]
  z = end_cells$healthy@assays$SCT@data[g,]
  
  d = data.frame(z=z, t=t)
  tmp = suppressMessages(gam(z ~ ns(t, df = 4), data=d))
  
  # bins for model fitting
  bb = seq(min(t), max(t), length.out = 100)
  gene_fit_vals[,g] = suppressMessages(predict(tmp,
                                               newdata = data.frame(t = bb)))
  p = summary(tmp)$parametric.anova$`Pr(>F)`[1]
  gene_fit_p = c(gene_fit_p, p)
  names(gene_fit_p)[length(gene_fit_p)] = g
}

gene_fit_p = fdrtool::fdrtool(gene_fit_p, statistic="pvalue", plot=F, 
                              verbose=F, cutoff.method="pct0", pct0=0.9)$qval
```


## Analysis of other conditions
First, clean the data from the other two conditions

```{r}
end_cells$embolised = FindNeighbors(end_cells$embolised, reduction = "pca", dims = 1:20,
                                    force.recalc = T, graph.name = "endo")
end_cells$embolised = FindClusters(end_cells$embolised, graph.name = "endo", algorithm = 2, 
                                   resolution = 0.8, verbose = F)
DimPlot(end_cells$embolised, reduction = "umap", group.by = "endo_res.0.8")
end_cells$embolised = SetIdent(end_cells$embolised, value = "endo_res.0.8")
mk_emb = FindAllMarkers(end_cells$embolised, assay = "SCT")

outcells = (end_cells$embolised@reductions$umap@cell.embeddings[,2]>2.3 & 
              end_cells$embolised@reductions$umap@cell.embeddings[,1]<1) # this removes what seem to be central hepatocytes
end_cells$embolised = end_cells$embolised[,!outcells]
end_cells$embolised = end_cells$embolised[,end_cells$embolised@meta.data[,"endo_res.0.8"]!=12] # remove these extra stellate cells

end_cells$embolised = suppressWarnings(SCTransform(end_cells$embolised, do.correct.umi = T, 
                                                 verbose = F,
                                                 vars.to.regress = c("unique_name","nCount_RNA"),
                                                 variable.features.rv.th = 1, seed.use = 1,
                                                 return.only.var.genes = F, 
                                                 variable.features.n = NULL))
end_cells$embolised = RunPCA(end_cells$embolised, verbose = F)
end_cells$embolised = RunUMAP(end_cells$embolised, dims = 1:15, verbose = F)
DimPlot(end_cells$embolised, reduction = "umap", group.by = "unique_name")


outcells = (end_cells$regenerating@reductions$umap@cell.embeddings[,2]>2.7 & 
              end_cells$regenerating@reductions$umap@cell.embeddings[,1]<(-1)) # this removes what seem to be central hepatocytes
end_cells$regenerating = end_cells$regenerating[,!outcells]

end_cells$regenerating = suppressWarnings(SCTransform(end_cells$regenerating, do.correct.umi = T, 
                                                 verbose = F,
                                                 vars.to.regress = c("unique_name","nCount_RNA"),
                                                 variable.features.rv.th = 1, seed.use = 1,
                                                 return.only.var.genes = F, 
                                                 variable.features.n = NULL))
end_cells$regenerating = RunPCA(end_cells$regenerating, verbose = F)
end_cells$regenerating = RunUMAP(end_cells$regenerating, dims = 1:15, verbose = F)
DimPlot(end_cells$regenerating, reduction = "umap", group.by = "unique_name")
```

Now use the top 500 genes to learn the pseudozonation coordinates and project the embolised and regenerating data

```{r}
sig_genes = gene_fit_p[gene_fit_p<=0.05]
top_sig_genes = names(sig_genes[order(sig_genes, decreasing = F)][1:500]) 

data_gam_h = cbind(data.frame("DPT" = df_dc$DPT),
                   Matrix::t(end_cells$healthy@assays$SCT@data[top_sig_genes,]))

gam_healthy = gam::gam(DPT ~ ., data = data_gam_h)

hea_pred = predict(gam_healthy,
                   Matrix::t(end_cells$health@assays$SCT@data[top_sig_genes,]))
reg_pred = predict(gam_healthy,
                   Matrix::t(end_cells$regenerating@assays$SCT@data[top_sig_genes,]))
emb_pred = predict(gam_healthy,
                   Matrix::t(end_cells$embolised@assays$SCT@data[top_sig_genes,]))

plot(density(data_gam_h$DPT), col = "green", xlim = c(-8, 21))
lines(density(reg_pred), col = "blue")
lines(density(emb_pred), col = "red")
lines(density(hea_pred), col = "grey50")
legend(-0.073,32, legend = c("healthy (DC1)", "healthy (pred)", 
                         "embolised (pred)", "regenerating (pred)"),
       fill = c("green", "grey50", "red", "blue"))
```

Calculate correlations

```{r}
cor_h = t(cor(df_dc$DPT, as.matrix(Matrix::t(end_cells$healthy@assays$SCT@data)),
              method = "sp"))
cor_emb = t(cor(emb_pred, as.matrix(Matrix::t(end_cells$embolised@assays$SCT@data)),
                method = "sp"))
cor_reg = t(cor(reg_pred, as.matrix(Matrix::t(end_cells$regenerating@assays$SCT@data)),
                method = "sp"))

l = list(cor_h, cor_emb, cor_reg)
all_corr = Reduce(function(x, y) merge(x,y, by = "rn", all = T), 
                  lapply(l, function(x) data.frame(x, rn = row.names(x))))
colnames(all_corr) = c("genes", "healthy", "embolised", "regenerating")

write.csv(all_corr, file = paste0("results/zonation_cond/end_correlations_zonation.csv"),
          col.names = T, row.names = F, quote = F)
```

Find varying genes

```{r}
traj_list = list("healthy" = df_dc$DPT, "embolised" = emb_pred, "regenerating" = reg_pred)

qvals_list = list()
fits_list = list()

# get HVG from all conditions
hvg = unique(c(end_cells[["healthy"]]@assays$SCT@var.features,
               end_cells[["embolised"]]@assays$SCT@var.features,
               end_cells[["regenerating"]]@assays$SCT@var.features))
hvg = hvg[hvg %in% rownames(end_cells[["healthy"]]@assays$SCT@data) &
            hvg %in% rownames(end_cells[["embolised"]]@assays$SCT@data) &
            hvg %in% rownames(end_cells[["regenerating"]]@assays$SCT@data)]

for(cond in names(end_cells)){
  print(cond)
  # Fit GAM for each gene using pseudotime as independent variable.
  t = traj_list[[cond]]
  gene_fit_p = c()
  gene_fit_vals = data.frame(row.names = 1:100)
  for(i in 1:length(hvg)){
    g = hvg[i]
    z = end_cells[[cond]]@assays$SCT@data[g,]
    
    d = data.frame(z=z, t=t)
    tmp = suppressMessages(gam(z ~ ns(t, df = 4), data=d))
    
    # bins for model fitting
    bb = seq(min(t), max(t), length.out = 100)
    gene_fit_vals[,g] = suppressMessages(predict(tmp,
                                                 newdata = data.frame(t = bb)))
    p = summary(tmp)$parametric.anova$`Pr(>F)`[1]
    gene_fit_p = c(gene_fit_p, p)
    names(gene_fit_p)[length(gene_fit_p)] = g
  }
  
  if(sum(is.na(gene_fit_p))>0){
    gene_fit_p[is.na(gene_fit_p)] = 1
  }
  gene_fit_p = fdrtool::fdrtool(gene_fit_p, statistic="pvalue", plot=F, 
                                verbose=F, cutoff.method="pct0", pct0=0.9)$qval
  
  qvals_list[[cond]] = gene_fit_p
  fits_list[[cond]] = gene_fit_vals
}
```

Save trajectories and genes

```{r}
save(fits_list, qvals_list, traj_list, 
     file = "results/zonation_cond/end_traj_qval_fits.RData")
```

Calculate distances for all significant genes

```{r}
for(cond in names(qvals_list)){
  gene_fit_p = qvals_list[[cond]]
  gene_fit_vals = fits_list[[cond]]
  gene_fit_vals_filt = t(gene_fit_vals[,names(gene_fit_p)[gene_fit_p<=0.05]])
  scale_fit = t(apply(gene_fit_vals_filt, 1, scale))
  
  # dist
  if(file.exists(paste0("results/zonation_cond/dtw_distance_end_", cond, "_sigGenes.RDS"))){
    next
  } else{
    dd = dtw::dtwDist(scale_fit) # TAKES 4.5 HOURS
    saveRDS(dd, paste0("results/zonation_cond/dtw_distance_end_", cond, "_sigGenes.RDS"))
  }
}
```

Cluster genes (per condition)

```{r}
top_n = 1000
for(cond in names(qvals_list)){
  print(cond)
  gene_fit_p = qvals_list[[cond]]
  gene_fit_vals = fits_list[[cond]]
  
  dd = readRDS(paste0("results/zonation_cond/dtw_distance_end_", cond, "_sigGenes.RDS"))
  
  gene_fit_vals_filt = t(gene_fit_vals[,names(gene_fit_p)[gene_fit_p<=0.05]])
  scale_fit = t(apply(gene_fit_vals_filt, 1, scale))
  
  # select top genes (0 if using all)
  topgenes = if(top_n==0) names(gene_fit_p) else names(head(gene_fit_p[order(gene_fit_p,
                                                                             decreasing = F)],
                                                            top_n))
  
  clust_genes = hclust(as.dist(dd[topgenes,topgenes]), method = "ward.D2")
  
  cls = cutree(clust_genes, 4)
  #ape::plot.phylo(ape::as.phylo(clust_genes), tip.color = rainbow(10)[cls])
  plt_list = list()
  for(i in unique(cls)){
    sub_df = scale_fit[names(cls)[cls==i],]
    plot_df = reshape2::melt(sub_df)
    plt_list[[paste0("gene_",i)]] = ggplot(plot_df, aes(x = Var2, y = value, group = Var1))+
      geom_hline(yintercept = 0, linetype = "dashed")+
      geom_line(alpha = 0.1, colour = "grey69")+
      stat_summary(aes(y = value, group=1), fun=mean, colour="red", geom="line",group=1)+
      labs(title = paste0("gene_",i))+
      theme_bw()
  }
  
  pdf(paste0("results/zonation_cond/end_gene_expression_profile_", cond, ".pdf"),
      height = 4.5, width = 4.4)
  print(cowplot::plot_grid(plotlist = plt_list, nrow = 2, ncol = 2, align = "hv"))
  dev.off()
  
  df_pval_cl = merge(data.frame(gene_fit_p), data.frame(cls), by = 0, all = T)
  colnames(df_pval_cl) = c("gene", "pval", "cluster")
  
  write.csv(df_pval_cl, file = paste0("results/zonation_cond/end_pval_clust_", cond, ".csv"),
            col.names = T, row.names = F, quote = F)

  write.csv(gene_fit_vals, 
            file = paste0("results/zonation_cond/end_gene_fits_", cond, ".csv"),
            col.names = T, row.names = F, quote = F)
}
```

Plot cell distributions in pseudotime

```{r}
plot_df = data.frame("pseudotime" = c(traj_list$healthy, traj_list$embolised,
                                      traj_list$regenerating),
           "donor" = c(end_cells$healthy@meta.data[,"Donor"],
                       end_cells$embolised@meta.data[names(traj_list$embolised),"Donor"],
                       end_cells$regenerating@meta.data[names(traj_list$regenerating),"Donor"]),
           "cond" = c(rep("healthy", length(traj_list$healthy)), 
                      rep("embolised", length(traj_list$embolised)),
                      rep("regenerating", length(traj_list$regenerating))))

pdf("results/zonation_cond/end_cell_distributions_zonation.pdf", useDingbats = F, 
    width = 6, height = 5)
ggplot(plot_df, aes(x = donor, y = pseudotime, fill = cond))+
  geom_boxplot(notch = T)+
  theme_bw()

ggplot(plot_df, aes(x = pseudotime, group = donor, colour = donor))+
  facet_wrap(~cond)+
  geom_density()+
  theme_bw()

ggplot(plot_df, aes(x = pseudotime, colour = cond))+
  geom_density()+
  theme_bw()
dev.off()
```

Do cross validation on the healthy trajectory

```{r}
sig_genes = qvals_list$healthy[qvals_list$healthy<=0.05]
top_sig_genes = names(sig_genes[order(sig_genes, decreasing = F)][1:500]) 
# 500 in this case actually seems to be one of the models with the lowest MSE

data_gam_h = cbind(data.frame("DPT" = traj_list$healthy),
                   Matrix::t(end_cells$healthy@assays$SCT@data[top_sig_genes,]))
colnames(data_gam_h) = gsub("-", "_", colnames(data_gam_h), fixed = T)
colnames(data_gam_h) = gsub(".", "_", colnames(data_gam_h), fixed = T)

cv_gam = CVgam(formula = DPT ~ ., data = data_gam_h, nfold = 10, seed = 1, method = "glm.fit")

pdf("results/zonation_healthy/end_original_fitted_pseudotime.pdf", useDingbats = F, 
    width = 6, height = 5)
plot(data_gam_h$DPT, cv_gam$fitted, xlab = "Healthy pseudotime (DPT424)",ylab = "Predicted Healthy",
     main = paste0("Healthy original vs fitted pseudotime\nMSE: ", round(cv_gam$mse, 6)),
     pch = 19, cex = 0.5)
abline(0,1, col = "blue", lwd = 3)
dev.off()
```

Save zonation results

```{r}
# define the intervals based on healthy
df_h = data.table("zonation_pt" = traj_list$healthy)
df_h$zonation_int = cut(df_h$zonation_pt, 3)

df_e = data.table("zonation_pt" = traj_list$embolised)
df_e = df_h[df_e, on="zonation_pt", roll=Inf, rollends = T]

df_r = data.table("zonation_pt" = traj_list$regenerating)
df_r = df_h[df_r, on="zonation_pt", roll=Inf, rollends = T]

list_df = list("healthy" = df_h, "embolised" = df_e, "regenerating" = df_r)

for(n in names(traj_list)){
  ptdf = list_df[[n]]
  rownames(ptdf) = colnames(end_cells[[n]])
  end_cells[[n]] = AddMetaData(end_cells[[n]], metadata = ptdf)
  write.csv(end_cells[[n]]@meta.data[,c("unique_name", "Condition", 
                                        "zonation_pt", "zonation_int")],
            file = paste0("results/zonation_cond/end_zonation_", n, ".csv"),
            col.names = T, row.names = T, quote = F)
}

saveRDS(end_cells, file = "results/zonation_cond/end_cells_zonation.RDS")
```



# Comparing genes between conditions
Load data

```{r}
end_cells = readRDS(file = "results/zonation_cond/end_cells_zonation.RDS")
hep_cells = readRDS(file = "results/zonation_cond/hep_cells_zonation.RDS")
```

Load fits

```{r}
load("results/zonation_cond/hep_traj_qval_fits.RData")
hep_fits = fits_list
hep_traj = traj_list
load("results/zonation_cond/end_traj_qval_fits.RData")
end_fits = fits_list
end_traj = traj_list
```

Calculate expression per bins

```{r}
nbins = 100

dfhep_h = data.table("zonation_pt" = hep_traj$healthy)
dfhep_h$zonation_int = cut(dfhep_h$zonation_pt, nbins)
dfhep_e = data.table("zonation_pt" = hep_traj$embolised)
dfhep_e = dfhep_h[dfhep_e, on="zonation_pt", roll=Inf, rollends = T]
dfhep_r = data.table("zonation_pt" = hep_traj$regenerating)
dfhep_r = dfhep_h[dfhep_r, on="zonation_pt", roll=Inf, rollends = T]

hep_mean_bins = list(
healthy = apply(hep_cells$healthy@assays$SCT@data, 1, 
                        function(x) tapply(x, dfhep_h$zonation_int, mean)),
embolised = apply(hep_cells$embolised@assays$SCT@data, 1, 
                        function(x) tapply(x, dfhep_e$zonation_int, mean)),
regenerating = apply(hep_cells$regenerating@assays$SCT@data, 1, 
                        function(x) tapply(x, dfhep_r$zonation_int, mean))
)
for(n in names(hep_mean_bins)){ hep_mean_bins[[n]][is.na(hep_mean_bins[[n]])] = 0 }


dfend_h = data.table("zonation_pt" = end_traj$healthy)
dfend_h$zonation_int = cut(dfend_h$zonation_pt, nbins)
dfend_e = data.table("zonation_pt" = end_traj$embolised)
dfend_e = dfend_h[dfend_e, on="zonation_pt", roll=Inf, rollends = T]
dfend_r = data.table("zonation_pt" = end_traj$regenerating)
dfend_r = dfend_h[dfend_r, on="zonation_pt", roll=Inf, rollends = T]

end_mean_bins = list(
healthy = apply(end_cells$healthy@assays$SCT@data, 1, 
                        function(x) tapply(x, dfend_h$zonation_int, mean)),
embolised = apply(end_cells$embolised@assays$SCT@data, 1, 
                        function(x) tapply(x, dfend_e$zonation_int, mean)),
regenerating = apply(end_cells$regenerating@assays$SCT@data, 1, 
                        function(x) tapply(x, dfend_r$zonation_int, mean))
)
for(n in names(end_mean_bins)){ end_mean_bins[[n]][is.na(end_mean_bins[[n]])] = 0 }
```

Correlate binned expression

```{r}
comb_cond = combn(names(end_cells), 2)
hep_cor_list = list()
for(i in 1:ncol(comb_cond)){
  c1 = comb_cond[1,i]
  c2 = comb_cond[2,i]
  
  genes = intersect(colnames(hep_mean_bins[[c1]]),colnames(hep_mean_bins[[c2]]))
  
  for(g in genes){
    hep_cor_list[[paste0(c1,"_",c2)]][[g]] = cor(cbind(hep_mean_bins[[c1]][,g],
                                                       hep_mean_bins[[c2]][,g]),
                                                 method = "sp")[1,2]
  }
}
hep_cor_list = lapply(hep_cor_list, data.frame)
hep_cor_list = lapply(hep_cor_list, function(x) cbind(rownames(x), x))

end_cor_list = list()
for(i in 1:ncol(comb_cond)){
  c1 = comb_cond[1,i]
  c2 = comb_cond[2,i]
  
  genes = intersect(colnames(end_mean_bins[[c1]]),colnames(end_mean_bins[[c2]]))
  
  for(g in genes){
    end_cor_list[[paste0(c1,"_",c2)]][[g]] = cor(cbind(end_mean_bins[[c1]][,g],
                                                       end_mean_bins[[c2]][,g]),
                                                 method = "sp")[1,2]
  }
}
end_cor_list = lapply(end_cor_list, data.frame)
end_cor_list = lapply(end_cor_list, function(x) cbind(rownames(x), x))
```

Save correlations

```{r}
save(hep_cor_list, end_cor_list, 
     file = "results/zonation_cond/zonation_cor_cond_lists.RData")
```





# LEGACY CODE
Find genes with differencial variation - not very interesting since most genes seem to be showing a variation

```{r}
cells = unlist(lapply(hep_cells, colnames))
genes = unique(unlist(lapply(hep_cells, VariableFeatures)))
exp_hep = allcells_css[genes,cells]@assays$SCT@data
genes = rownames(exp_hep)

df_meta = data.frame("pt" = c(df_dc$DC1, emb_pred, reg_pred),
                     "cond" = allcells_css[genes,cells]@meta.data$Condition)

# Fit GAM for each gene using pseudotime as independent variable.
t = df_meta$pt
cond = df_meta$cond
gene_fit_int_p = matrix(nrow = 0, ncol = 3)
colnames(gene_fit_int_p) = c("spline", "condition", "interaction")
gene_fit_int_vals = data.frame(row.names = 1:300)
predgrid = data.frame("t" = rep(seq(min(t), max(t), length.out = 100), 3),
                      "cond" = rep(names(hep_cells), each = 100))
for(i in genes){
  z = exp_hep[i,]
  
  d = data.frame(z=z, t=t, cond=cond)
  tmp = suppressMessages(gam(z ~ ns(t, df = 4)*cond, data=d))
  
  # bins for model fitting
  gene_fit_int_vals[,g] = suppressMessages(predict(tmp, newdata = predgrid))
  p = summary(tmp)$parametric.anova$`Pr(>F)`[1:3]
  gene_fit_int_p = rbind(gene_fit_int_p, p)
  rownames(gene_fit_int_p)[nrow(gene_fit_int_p)] = i
}
```

DTW condition vs condition

```{r}
cond_comb = combn(names(qvals_list), 2)
for(i in 1:ncol(cond_comb)){
  genes = intersect(names(qvals_list[[cond_comb[1,i]]])[qvals_list[[cond_comb[1,i]]]<=0.05],
                    names(qvals_list[[cond_comb[2,i]]])[qvals_list[[cond_comb[2,i]]]<=0.05])
  
  sc_list = list()
  for(cond in cond_comb[,i]){
    gene_fit_vals = fits_list[[cond]]
    gene_fit_vals_filt = t(gene_fit_vals[,genes])
    sc_list[[cond]] = t(apply(gene_fit_vals_filt, 1, scale))
  }
  
  xxx = dtw(sc_list[[1]], sc_list[[2]], window.type = "itakura", dist.method = "DTW")
  yyy = dtw(dist(t(sc_list[[1]]), t(sc_list[[2]]), method = "DTW"), window.type = "itakura")
}
```



```{r}
source("aux_scripts/primate_cerebral_organoids_pt_alignment.r")

xcv = align_pt_traj(expr_ref = hep_cells$healthy@assays$SCT@data[genes,], 
                    pt_ref = traj_list$healthy,
                    expr_query = hep_cells$embolised@assays$SCT@data[genes,], 
                    pt_query = traj_list$embolised, num_breaks_ref = 100, 
                    ref_name = "healthy", query_name = "embolised")

plot(traj_list$embolised, xcv$pt_query_aligned)
abline(0,1)

xcb = align_pt_traj(expr_ref = hep_cells$healthy@assays$SCT@data[genes,], 
                    pt_ref = traj_list$healthy,
                    expr_query = hep_cells$regenerating@assays$SCT@data[genes,], 
                    pt_query = traj_list$regenerating, num_breaks_ref = 100, 
                    ref_name = "healthy", query_name = "regenerating")

plot(traj_list$regenerating, xcb$pt_query_aligned)
abline(0,1)
```

```{r}
g = "CLEC1B"
plt_df = data.frame("traj" = c(traj_list$regenerating, traj_list$healthy),
           "exp" = c(end_cells$regenerating@assays$SCT@data[g,names(traj_list$regenerating)],
           end_cells$healthy@assays$SCT@data[g,]),
           "cond" = c(rep("regenerating", length(traj_list$regenerating)),
                      rep("healthy", length(traj_list$healthy))),
           "cl" = c(end_cells$regenerating@meta.data$allcells_simp,
                    end_cells$healthy@meta.data$allcells_simp))
plt_df = plt_df[order(plt_df$traj, decreasing = T),]

ggplot(plt_df, aes(x = traj, y = exp, group = cond, colour = cond))+
  geom_point()+
  stat_smooth()+
  labs(title = g)+
  theme_bw()

ggplot(plt_df, aes(x = traj, y = exp, group = cond, colour = cl, shape = cond))+
  geom_point()+
  stat_smooth()+
  labs(title = g)+
  theme_bw()

xxx = dtw::dtw(x = plt_df[plt_df$cond=="healthy","exp"], y = plt_df[plt_df$cond=="regenerating","exp"])
plot(xxx)
abline(0,1)

hti = cut_number(plt_df$traj[plt_df$cond=="healthy"], 100)
rti = cut_number(plt_df$traj[plt_df$cond=="regenerating"], 100)
df2 = data.frame("r" = tapply(plt_df$exp[plt_df$cond=="regenerating"], rti, mean),
                 "h" = tapply(plt_df$exp[plt_df$cond=="healthy"], hti, mean))
yyy = dtw::dtw(x = df2$h, y = df2$r)
plot(yyy)
abline(0,1)
```


